### 2020-06-06 src/bbData2020.Rmd
further updates to explore temporal features

---
title: "descriptive anaysis"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r settings, include = FALSE}
library(dplyr)
library(reshape2)
library(lubridate)
library(ggplot2)
library(openair)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(tmap)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)

# knitr::opts_chunk$set(echo = FALSE
#                       , comment = NA
#                       , warning = FALSE
#                       , error = FALSE
#                       , message = FALSE
#                       , tidy = TRUE)

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# This is the project path
path <- find_rstudio_root_file()

```

```{r echo=FALSE, results= 'asis', message=FALSE, warning=TRUE}
#upload data file
path.data <- paste(path, "./data/raw/speedtest.csv", sep = "")

#bb2020 <- read.csv("./data/raw/speedtest.csv")
bb2020 <- read.csv(path.data)

#name columns
names(bb2020)[1:6] <- c("datetext","speeddown","speedup","provider","lat", "lon")

str(bb2020)
summary(bb2020) #2030205 obs
#conversions needed to numeric? df$col <- as.numeric(df$col) 
#remove empty rows or coerced to NA? df[complete.cases(df),] or df <- df[!is.na(df$col),]

#remove outliers based on Riddlesden and Singleton 2014 for slower end
#check fastest upload and download speeds commercially available from Virgin Media (Gigaclear offers even faster!)
bb2020 <- bb2020[bb2020$speeddown>512,] 
bb2020 <- bb2020[bb2020$speeddown<362000,]
bb2020 <- bb2020[bb2020$speedup<21000,] #1839333

#remove below seconds, create columns for date and hour
bb2020$datetext <- strtrim(x = bb2020$datetext, width = 19) 
bb2020$datetext2 <- bb2020$datetext
bb2020$datetext2 <- strtrim(x = bb2020$datetext2, width = 13)
bb2020$datetext2 <- colsplit(bb2020$datetext2, pattern = " ", c("date", "hour"))
bb2020$hour <- bb2020$datetext2$hour
bb2020$date <- bb2020$datetext2$date
#delete datetext2 column(s)
bb2020 <- bb2020[ ,c(1:6,8:9)]
#convert to date form datetext and date columns
bb2020$datetext <- as.POSIXct(strptime(bb2020$datetext, format = "%Y-%m-%d %H:%M:%S", "GMT")) 
bb2020$date <- as.POSIXct(strptime(bb2020$date, format = "%Y-%m-%d", "GMT"))
```

# Temporal analysis

```{r echo=FALSE, results= 'asis', message=FALSE, warning=TRUE}
#add column for day of the week
bb2020$weekday <- wday(bb2020$date, label = TRUE)
#create 2 dataframes for 2 matching periods
Wks2019 <- selectByDate(bb2020, start = "2019-01-21", end = "2019-06-02")
names(Wks2019)[1:2] <- c("dateOnly", "date") #413788 obs
Wks2020 <- selectByDate(bb2020, start = "2020-01-20", end = "2020-05-31")
names(Wks2020)[1:2] <- c("dateOnly", "date") #521773 obs
#timeplots
TimeVar2019 <- timeVariation(Wks2019, pollutant = "speeddown", statistic = "mean", name.pol = "speedown 2019")
TimeVar2019up <- timeVariation(Wks2019, pollutant = "speedup", statistic = "mean", name.pol = "speedup 2019")
TimeVar2020 <- timeVariation(Wks2020, pollutant = "speeddown", statistic = "mean", name.pol = "speeddown 2020")
TimeVar2020up <- timeVariation(Wks2020, pollutant = "speedup", statistic = "mean", name.pol = "speedup 2020")
timePlot(Wks2019, pollutant = "speeddown", avg.time = "day", smooth = TRUE, name.pol = "speeddown 2019")
timePlot(Wks2020, pollutant = "speeddown", avg.time = "day", smooth = TRUE, name.pol = "speeddown 2020")
timePlot(Wks2019, pollutant = "speedup", avg.time = "day", smooth = TRUE, name.pol = "speedup 2019")
timePlot(Wks2020, pollutant = "speedup", avg.time = "day", smooth = TRUE, name.pol = "speedup 2020")
```

# Frequency per date

```{r echo=FALSE, results= 'asis', message=FALSE, warning=TRUE}
DateCount19 <- count(Wks2019, dateOnly)
names(DateCount19) <- c("Date19", "tests19")
HrCount19 <- count(Wks2019, dateOnly, hour)
HrSpeed19 <- summarise(group_by(Wks2019, dateOnly, hour), mean(speeddown))
HrSpeed19up <- summarise(group_by(Wks2019, dateOnly, hour), mean(speedup))
HrStats19 <- left_join(HrCount19, HrSpeed19)
HrStats19 <- left_join(HrStats19, HrSpeed19up)
names(HrStats19) <- c("dateOnly", "hour", "test19", "meanSp19", "meanSp19up")
DateCount20 <- count(Wks2020, dateOnly)
names(DateCount20) <- c("Date20", "tests20")
HrCount20 <- count(Wks2020, dateOnly, hour)
HrSpeed20 <- summarise(group_by(Wks2020, dateOnly, hour), mean(speeddown))
HrSpeed20up <- summarise(group_by(Wks2020, dateOnly, hour), mean(speedup))
HrStats20 <- left_join(HrCount20, HrSpeed20)
HrStats20 <- left_join(HrStats20, HrSpeed20up)
names(HrStats20) <- c("dateOnly", "hour", "test20", "meanSp20", "meanSp20up")
DateCount <- cbind(DateCount19, DateCount20)
HrStats <- cbind(HrStats19, HrStats20)
#add columns tests per hour, mean speeddown per hr, mean speed up per hr
Wks2019 <- left_join(Wks2019, HrStats19, by = c("dateOnly", "hour"))
Wks2020 <- left_join(Wks2020, HrStats20, by = c("dateOnly", "hour"))
#frequency plot
timePlot(Wks2019, pollutant = "test19", avg.time = "hour", smooth = TRUE, name.pol = "hourlyTests 2019")
timePlot(Wks2020, pollutant = "tests20", avg.time = "hour", smooth = TRUE, name.pol = "hourlyTests 2020")
TimeVar2019n <- timeVariation(Wks2019, pollutant = "test19", statistic = "mean", name.pol = "hourlytests 2019")
TimeVar2020n <- timeVariation(Wks2020, pollutant = "test20", statistic = "mean", name.pol = "hourlytests 2020")
Plot2019 <- ggplot(DateCount19, aes(Date19, tests19))
Plot2019 + geom_line()
Plot2020 <- ggplot(DateCount20, aes(Date20, tests20))
Plot2020 + geom_line()

```

# Spatial analysis

```{r echo=FALSE, results= 'asis', message=FALSE, warning=TRUE}
#convert to spatial object
coords_bb <- cbind(bb2020$lon, bb2020$lat)
bbNUTS3 <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2020))
proj4string(bbNUTS3) <- CRS("+init=epsg:4326") #define projection

#get NUTS3 from up a level and read in (145 NUTS3)
NUTS3 <- readOGR("../NUTS_Level_3_January_2015_Super_Generalised_Clipped_Boundaries_in_England_and_Wales.shp") 
NUTS3 <- spTransform(NUTS3, CRS("+init=epsg:4326"))                                
NUTS3@data$nuts315nm <- as.character(NUTS3@data$nuts315nm)

#spatial join to NUTS3
bbNUTS3sp <- over(bbNUTS3, NUTS3[, "nuts315nm"]) #not a spatial object
bbNUTS3$nuts315nm <- bbNUTS3sp$nuts315nm

#create dataframe object to analyse (remove lat and lon?)
bb2020sp <- bbNUTS3@data
#% obs NA for NUTS3?

#summarise by date and geography
NUTS3datecount <- count(bb2020sp, nuts315nm, date)
NUTS3dateAvg <- summarise(group_by(NUTS3datecount, nuts315nm), mean(n))

#map NUTS3 mean test frequency per date
NUTS3@data <- left_join(NUTS3@data, NUTS3dateAvg, by = "nuts315nm")
var <- NUTS3@data[ ,'mean(n)']
breaks <- classIntervals(var, n = 5, style = "quantile")
my_colours <- brewer.pal(5, "Blues")
plot(NUTS3, col = my_colours[findInterval(var, breaks$brks, all.inside = TRUE)], axes = FALSE,
     border = rgb(0.8, 0.8, 0.8, 0))
legend(x = "top", legend = leglabs(breaks$brks), fill = my_colours, bty = "n", cex = 0.5)
```