### 2020-06-19 src/ts_clusters.Rmd

---
title: "Time series clusters"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(tsbox)
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)
library(imputeTS)
library(tsbox)
library(dtwclust)

# This is the project path
path <- find_rstudio_root_file()
```

```{r include = FALSE}

# load and clean the data.
# copied from bbData2020.Rmd

#upload data file
path.data <- paste(path, "./data/raw/speedtest.csv", sep = "")

#bb2020 <- read.csv("./data/raw/speedtest.csv")
bb2020 <- read.csv(path.data)

#name columns
names(bb2020)[1:6] <- c("datetext","speeddown","speedup","provider","lat", "lon")

str(bb2020)
summary(bb2020) #2030205 obs
#conversions needed to numeric? df$col <- as.numeric(df$col) 
#remove empty rows or coerced to NA? df[complete.cases(df),] or df <- df[!is.na(df$col),]

#remove outliers based on Riddlesden and Singleton 2014 for slower end
#check fastest upload and download speeds commercially available from Virgin Media (Gigaclear offers even faster!)
bb2020 <- bb2020[bb2020$speeddown>512,] 
bb2020 <- bb2020[bb2020$speeddown<362000,]
bb2020 <- bb2020[bb2020$speedup<21000,] #1839333

#remove below seconds, create columns for date and hour
bb2020$datetext <- strtrim(x = bb2020$datetext, width = 19) 
bb2020$datetext2 <- bb2020$datetext
bb2020$datetext2 <- strtrim(x = bb2020$datetext2, width = 13)
bb2020$datetext2 <- colsplit(bb2020$datetext2, pattern = " ", c("date", "hour"))
bb2020$hour <- bb2020$datetext2$hour
bb2020$date <- bb2020$datetext2$date
#delete datetext2 column(s)
bb2020 <- bb2020[ ,c(1:6,8:9)]
#convert to date form datetext and date columns
bb2020$datetext <- as.POSIXct(strptime(bb2020$datetext, format = "%Y-%m-%d %H:%M:%S", "GMT")) 
bb2020$date <- as.POSIXct(strptime(bb2020$date, format = "%Y-%m-%d", "GMT"))
#create dataframe with number of tests, mean and standard deviation by month for adding to dataframe later
MonthFreq <- timeAverage(bb2020, avg.time = "month", statistic = "frequency")
MonthMean <- timeAverage(bb2020, avg.time = "month", statistic = "mean")
MonthSD <- timeAverage(bb2020, avg.time = "month", statistic = "sd")
MonthStats <- cbind(MonthFreq[,1:2], MonthMean[,c(3:4)], 
                    MonthSD[,c(3:4)])
names(MonthStats) <- c("month", "monthlyTests", "monthlyMeanSpDown", 
                       "monthlyMeanSpUp", "monthlySDSpDown", 
                       "monthlySDSpUp")
MonthStats$month <- month(MonthStats$month, label = TRUE, abbr = TRUE)
#add column for month and day of the week to create stats below
bb2020$month <- month(bb2020$date, label = TRUE, abbr = TRUE)
bb2020$weekday <- wday(bb2020$date, label = TRUE)
#divide MonthStats for future join with separate 2019 and 2020 dataframes
MonthStats19 <- MonthStats[1:6,]
MonthStats20 <- MonthStats[13:17,]
```

## Structured time series

```{r eval=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# convert to spatial object
coords_bb <- cbind(bb2020$lon, bb2020$lat)
bb.la <- SpatialPointsDataFrame(coords_bb, data = data.frame(bb2020))
proj4string(bb.la) <- CRS("+init=epsg:4326") #define projection

# get LA 
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
# UK BFC la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson")

# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))
la@data$LAD19NM <- as.character(la@data$LAD19NM)

# spatial join to LAD
bb.la.sp <- over(bb.la, la[, "LAD19NM"]) #not a spatial object
bb.la$LAD19NM <- bb.la.sp$LAD19NM

# create dataframe object to analyse (remove lat and lon?)
bb2020sp <- bb.la@data

# drop NAs
bb2020sp <- bb2020sp[!is.na(bb2020sp$LAD19NM),]
#13431/1839332 = .0073 NA LAD19NM, coord outside UK

# summarise by date and geography
la.date.count <- count(bb2020sp, LAD19NM, date)
la.dateAvg <- summarise(group_by(la.date.count, LAD19NM), mean(n))
range(la.dateAvg$`mean(n)`)

# map LA mean test frequency per date
la@data <- left_join(la@data, la.dateAvg, by = "LAD19NM")
var <- la@data[ ,'mean(n)']
breaks <- classIntervals(var, n = 5, style = "quantile")
my_colours <- brewer.pal(5, "Blues")
plot(la, col = my_colours[findInterval(var, breaks$brks, all.inside = TRUE)], axes = FALSE,
                 border = rgb(0.8, 0.8, 0.8, 0))

```

## Keep working hours and days

The time series are based on mean upload and download speeds as well as the total number of tests for working days and during working hours

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=600, fig.width=14}

bb2020sp <- bb2020sp %>% arrange(datetext)

# keep working hours
bb2020sp <- bb2020sp[bb2020sp$hour>=9 & bb2020sp$hour<17,]

# drop w/e
bb2020sp <- bb2020sp %>%
  filter(bb2020sp$weekday!="Sat" & bb2020sp$weekday!="Sun")

head(bb2020sp)
```

## Time series clusters

I create clusters based on upload, download and number of tests time series for tests working hours and days from 16/2/2020 onwards.

More work needs to be done to select the optimal *k* as well as the clustering methods.
The `dtwclust` package comes with quite some readings, saved in the `.\literature\`.

Also, we need to think about upload / download / differnces / period + **imputation**.

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# download
ts.down <- bb2020sp %>%
  group_by(date,LAD19NM) %>%
  summarise(mean.down = mean(speeddown))#, mean.down = mean(speeddown), n.tests = n())

# upload
ts.up <- bb2020sp %>%
  group_by(date,LAD19NM) %>%
  summarise(mean.up = mean(speedup))

# tests
ts.tests <- bb2020sp %>%
  group_by(date,LAD19NM) %>%
  summarise(n.tests = n())
```

### Upload speed

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# upload working hours and days

ts.wide <- ts.up %>%
  select(date, LAD19NM, mean.up) %>%
  filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means
library(imputeTS)
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# 5 - 20 cluster sollutions
pc_k <- tsclust(ts.wide, k = c(5,10,15,20), seed = 94L,
                type="partitional",
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

# decide k, see RJ-2019-023 for max/min
names(pc_k) <- paste0("k_", c("5","10","15", "20"))
sapply(pc_k, cvi, type = "internal")

# I need to work on the selection

pc_k <- tsclust(ts.wide, k = 20, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

plot(pc_k)

cluster_up <- data.frame(LAD=ts.wide.names,
                           cluster=pc_k@cluster)
```

### Download speed

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# downlad working hours and days

ts.wide <- ts.down %>%
  select(date, LAD19NM, mean.down) %>%
  filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means
library(imputeTS)
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# 5 - 20 cluster sollutions
pc_k <- tsclust(ts.wide, k = c(5,10,15,20), seed = 94L,
                type="partitional",
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

# decide k, see RJ-2019-023 for max/min
names(pc_k) <- paste0("k_", c("5","10","15", "20"))
sapply(pc_k, cvi, type = "internal")

# I need to work on the selection

pc_k <- tsclust(ts.wide, k = 20, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

plot(pc_k)

cluster_down <- data.frame(LAD=ts.wide.names,
                         cluster=pc_k@cluster)
```

### N. of speeds

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# tests working hours and days

ts.wide <- ts.tests %>%
  select(date, LAD19NM, n.tests) %>%
  filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means
library(imputeTS)
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# 5 - 20 cluster sollutions
pc_k <- tsclust(ts.wide, k = c(5,10,15,20), seed = 94L,
                type="partitional",
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

# decide k, see RJ-2019-023 for max/min
names(pc_k) <- paste0("k_", c("5","10","15", "20"))
sapply(pc_k, cvi, type = "internal")

# I need to work on the selection

pc_k <- tsclust(ts.wide, k = 15, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))

plot(pc_k)

cluster_tests <- data.frame(LAD=ts.wide.names,
                           cluster=pc_k@cluster)

clusters <- merge(cluster_up, cluster_down, by = "LAD")
names(clusters)[2] <- "cluster.up"
names(clusters)[3] <- "cluster.down"
clusters <- merge(clusters, cluster_tests, by = "LAD")
names(clusters)[4] <- "cluster.tests"
#sapply(clusters, function(x) sum(is.na(x)))

clusters  %>% arrange(cluster.up) %>% kable() 
```

## Mapping the clusters

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

la@data <- merge(la@data, clusters, by.x = "LAD19NM", by.y = "LAD")

var <- la@data[ ,'cluster.up']
plot(la, col = var, main = "Upload clusters",
     border = rgb(0.8, 0.8, 0.8, 0))

var <- la@data[ ,'cluster.down']
plot(la, col = var, main = "Download clusters",
     border = rgb(0.8, 0.8, 0.8, 0))

var <- la@data[ ,'cluster.tests']
plot(la, col = var, main = "Tests clusters",
     border = rgb(0.8, 0.8, 0.8, 0))

```