### 2020-06-29 src/Data_Spatial.Rmd
I just moved the .Rmd to the .\src and added some spaces to makeit more readable.

---
title: "Data_Spatial"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: html_document
df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(tsbox)
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)

# This is the project path
path <- find_rstudio_root_file()
```

```{r include = FALSE}

# load and clean the data.
# copied from bbData2020.Rmd

#upload data file
path.data <- paste(path, "./data/raw/speedtest.csv", sep = "")

#bb2020 <- read.csv("./data/raw/speedtest.csv")
bb2020 <- read.csv(path.data)

#name columns
names(bb2020)[1:6] <- c("datetext","speeddown","speedup","provider","lat", "lon")

str(bb2020)
summary(bb2020) #2030205 obs
#conversions needed to numeric? df$col <- as.numeric(df$col) 
#remove empty rows or coerced to NA? df[complete.cases(df),] or df <- df[!is.na(df$col),]

#remove outliers based on Riddlesden and Singleton 2014 for slower end
#check fastest upload and download speeds commercially available from Virgin Media (Gigaclear offers even faster!)
bb2020 <- bb2020[bb2020$speeddown>512,] 
bb2020 <- bb2020[bb2020$speeddown<362000,]
bb2020 <- bb2020[bb2020$speedup<21000,] #1839333

#remove below seconds, create columns for date and hour
bb2020$datetext <- strtrim(x = bb2020$datetext, width = 19) 
bb2020$datetext2 <- bb2020$datetext
bb2020$datetext2 <- strtrim(x = bb2020$datetext2, width = 13)
bb2020$datetext2 <- colsplit(bb2020$datetext2, pattern = " ", c("date", "hour"))
bb2020$hour <- bb2020$datetext2$hour
bb2020$date <- bb2020$datetext2$date

#delete datetext2 column(s)
bb2020 <- bb2020[ ,c(1:6,8:9)]

#convert to date form datetext and date columns
bb2020$datetext <- as.POSIXct(strptime(bb2020$datetext, format = "%Y-%m-%d %H:%M:%S", "GMT")) 
bb2020$date <- as.POSIXct(strptime(bb2020$date, format = "%Y-%m-%d", "GMT"))

#add column for month and day of the week to create stats below
bb2020$month <- month(bb2020$date, label = TRUE, abbr = TRUE)
bb2020$weekday <- wday(bb2020$date, label = TRUE)

#create 2 dataframes for 2 matching periods
Wks2019 <- selectByDate(bb2020, start = "2019-01-21", end = "2019-06-02")
names(Wks2019)[1:2] <- c("dateOnly", "date") #413788 obs
Wks2019$Year <- rep(2019)
Wks2020 <- selectByDate(bb2020, start = "2020-01-20", end = "2020-05-31")
names(Wks2020)[1:2] <- c("dateOnly", "date") #521773 obs
Wks2020$Year <- rep(2020)
TSbb19_20 <- rbind(Wks2019, Wks2020) #935561 obs

#exclude obs from Saturdays, Sundays, bank holidays and midnight-6am
TSbb19_20 <- selectByDate(TSbb19_20, day = "weekday", hour = 6:23) #641830 obs
TSbb19_20$dateOnly <- as.character(TSbb19_20$dateOnly)
TSbb19_20 <- TSbb19_20[which(TSbb19_20$dateOnly != "2019-04-19" &
                               TSbb19_20$dateOnly != "2019-04-22" &
                               TSbb19_20$dateOnly != "2019-05-06" &
                               TSbb19_20$dateOnly != "2019-05-27" &
                               TSbb19_20$dateOnly != "2020-04-10" &
                               TSbb19_20$dateOnly != "2020-04-13" &
                               TSbb19_20$dateOnly != "2020-05-08" &
                     TSbb19_20$dateOnly != "2020-05-25"),] #619821 obs
rm(bb2020)
rm(Wks2019)
rm(Wks2020)
```

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# Structured TS

# convert to spatial object
coords_bb <- cbind(TSbb19_20$lon, TSbb19_20$lat)
bb.la <- SpatialPointsDataFrame(coords_bb, data = data.frame(TSbb19_20))
proj4string(bb.la) <- CRS("+init=epsg:4326") #define projection

# get LA 
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
la <- readOGR(path.json)#, layer="OGRGeoJSON")
# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))
la@data$LAD19NM <- as.character(la@data$LAD19NM)

# spatial join to LAD
bb.la.sp <- over(bb.la, la[, "LAD19NM"]) #not a spatial object
bb.la$LAD19NM <- bb.la.sp$LAD19NM

# create dataframe object to analyse (remove lat and lon?)
TSbb19_20sp <- bb.la@data

# drop NAs
TSbb19_20sp <- TSbb19_20sp[!is.na(TSbb19_20sp$LAD19NM),]
#NA LAD19NM, coord outside UK =  615516 obs (99.3% of obs retained)
```

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=600, fig.width=14}

#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs

#5 weekdays x 18 hours x 382 LA = 34,380 obs
ts.tests <- TS2020 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(n.tests = n())
sapply(ts.tests, function(x) sum(is.na(x)))

# summarise by frequency of tests and geography
la.dateAvg <- summarise(group_by(ts.tests, LAD19NM), mean(n.tests))
range(la.dateAvg$`mean(n.tests)`)
#exclude LAs with average of less than 3 tests per hour per weekday
la.exc <- la.dateAvg[which(la.dateAvg$`mean(n.tests)` <3),]
la.dateAvg <- la.dateAvg[which(la.dateAvg$`mean(n.tests)` >=3),] #374 or 98% of LADs, so sufficient to keep 90 data points in ts
la.exc
#remove 8 LAD from TS2019 and TS2020?

# map LA mean test frequency per date
la@data <- left_join(la@data, la.dateAvg, by = "LAD19NM")
var <- la@data[ ,"mean(n.tests)"]
breaks <- classIntervals(var, n = 5, style = "quantile")
my_colours <- brewer.pal(5, "Blues")
plot(la, col = my_colours[findInterval(var, breaks$brks, all.inside = TRUE)], axes = FALSE,
                 border = rgb(0.8, 0.8, 0.8, 0))

#look at 2019?
ts.tests <- TS2019 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(n.tests = n())
sapply(ts.tests, function(x) sum(is.na(x)))

```