### 2020-07-14 descriptive_Clusters.Rmd

---
title: "descriptive_Clusters"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(tsbox)
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)

# This is the project path
path <- find_rstudio_root_file()
```

```{r include = FALSE}

# load the data created by Data_Spatial.Rmd and ts_clusters.Rmd

LAD.cluster <- paste(path, "./data/temp/clusters.csv", sep = "")

clusters <- read_csv(LAD.cluster)

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")

TSbb19_20sp <- read_csv(path.data)

#split out 2020 before join to clusters to create descriptives
TS2020 <- selectByDate(TSbb19_20sp, year = 2020)

#join clusters to TS2020
names(clusters) [1] <- "LAD19NM"
TS2020 <- left_join(TS2020, clusters)
#remove NAs? - otherwise cannot create unified dataframes for below stats

#summarise by LADs per cluster and Total tests / Mean speed per cluster
LAD.cluster.tests <- clusters %>%
  group_by(cluster.tests) %>%
  summarise(n.LADs = n())
LAD.cluster.tests$Total <- TS2020 %>%
  group_by(cluster.tests) %>%
  summarise(n.cluster = n())
LAD.cluster.up <- clusters %>%
  group_by(cluster.up) %>%
  summarise(up.LADs = n())
LAD.cluster.up$Mean <- summarise(group_by(TS2020, cluster.up),
  mean(speedup))
LAD.cluster.up$SD <- summarise(group_by(TS2020, cluster.up),
  sd(speedup))
LAD.cluster.down <- clusters %>%
  group_by(cluster.down) %>%
  summarise(down.LADs = n())
LAD.cluster.down$Mean <- summarise(group_by(TS2020, cluster.down),
  mean(speeddown))
LAD.cluster.down$SD <- summarise(group_by(TS2020, cluster.down),
  sd(speeddown))
LAD.cluster.diffUp <- clusters %>%
  group_by(cluster.up.diff) %>%
  summarise(diffUp.LADs = n())
LAD.cluster.diffDown <- clusters %>%
  group_by(cluster.down.diff) %>%
  summarise(diffDown.LADs = n())

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r }
#summarise test frequency clusters by hour of each weekday and plot
cluster.tests <- TS2020 %>%
  group_by(cluster.tests, weekday, hour) %>%
  summarise(n.tests = n())
names(cluster.tests)[1] <- "clusterID"
cluster.tests$clusterID <- as.factor(cluster.tests$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.tests <- merge(cluster.tests, help.dates, by = "weekday")
cluster.tests$hour <- paste(cluster.tests$hour, ":00", sep="")
cluster.tests$help.date.time <- with(cluster.tests, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for number tests
TestCluster <- ggplot(cluster.tests, aes(help.date.time, n.tests, 
                                         colour = clusterID))
TestCluster + geom_line() + labs(x = "Weekdays March-May") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri"))

#summarise upload speed clusters by hour of each weekday and plot
cluster.up <- TS2020 %>%
  group_by(cluster.up, weekday, hour) %>%
  summarise(mean.up = mean(speedup))
names(cluster.up)[1] <- "clusterID"
names(LAD.cluster.up)[1] <- "clusterID"
cluster.up <- left_join(cluster.up, LAD.cluster.up)
cluster.up <- cluster.up[which(cluster.up$up.LADs > 5),]
cluster.up$clusterID <- as.factor(cluster.up$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.up <- merge(cluster.up, help.dates, by = "weekday")
cluster.up$hour <- paste(cluster.up$hour, ":00", sep="")
cluster.up$help.date.time <- with(cluster.up, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for mean upload speed
UpCluster <- ggplot(cluster.up, aes(help.date.time, mean.up, 
                                         colour = clusterID))
UpCluster + geom_line() + labs(x = "Weekdays March-May") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri"))

#summarise download speed clusters by hour of each weekday and plot
cluster.down <- TS2020 %>%
  group_by(cluster.down, weekday, hour) %>%
  summarise(mean.down = mean(speeddown))
names(cluster.down)[1] <- "clusterID"
names(LAD.cluster.down)[1] <- "clusterID"
cluster.down <- left_join(cluster.down, LAD.cluster.down)
cluster.down <- cluster.down[which(cluster.down$down.LADs > 5),]
cluster.down$clusterID <- as.factor(cluster.down$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.down <- merge(cluster.down, help.dates, by = "weekday")
cluster.down$hour <- paste(cluster.down$hour, ":00", sep="")
cluster.down$help.date.time <- with(cluster.down, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for mean download speeds
TestCluster <- ggplot(cluster.down, aes(help.date.time, mean.down, 
                                         colour = clusterID))
TestCluster + geom_line() + labs(x = "Weekdays March-May") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri"))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.