### 2020-12-01 paper/v1/broadband.speed.covid.Rmd
Paper edits, esp tests descriptives in Methods section.

---
title: |
  WFH and broadband speed (title needs rework)
type: ARTICLE TEMPLATE
author:
  - name: A. N. Author
    affil: a
    email: latex.helpdesk@tandf.co.uk
  - name: John Smith
    affil: b
    email: john.smith@uni-freiburg.de
affiliation:
  - num: a
    address: |
      Taylor & Francis, 4 Park Square, Milton Park, Abingdon, UK
  - num: b
    address: |
      Institut für Informatik, Albert-Ludwigs-Universität, Freiburg, Germany
bibliography: bibliography.bib
appendix: appendix.tex
abstract: |
  TBC
keywords: |
  covid; internet; working from home; broadband speed; time-series clusters
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction

<!-- [our google doc](https://docs.google.com/document/d/1PWjkmgzWGYKR9wFogKYw7l-8mZLoORt593x-Tu-f2-M/edit#heading=h.i5om1o8wpcd9) -->

During the pandemic, working from home using digital technologies, whether partially
or exclusively, was transformed from a niche means of accessing work, albeit one that
had been on a slow, upward trend, to a widespread way of life in many countries.
The ability to work from home or telecommute meant millions retained their jobs and,
to a varying extent, maintained productivity during periods of strict
lockdown around the world.
However, this ability has not been evenly distributed socially or spatially,
creating a new type of digital divide.
On one side are those who can work from home, supported by digital technologies,
and have thus been able to enjoy both economic resilience and greater personal safety.
On the other side, previously employed individuals have been forced to accept
furlough or redundancy packages unless they are part of the cadre of essential
workers, who are potentially at high risk of infection.
Whilst the basis for this new digital divide has been viewed as mainly occupational,
here we consider whether the divide is also technological.

<!-- our approach -->
Using the UK as a case study, this paper aims to understand how the quality and
reliability of internet service, as reflected in *experienced* internet speeds,
may reinforce or redress the spatial and social dimensions of the digital division
exposed by the pandemic.
To do so, we employ volunteered geographic data on individual broadband speed tests
and state-of-the-art time-series clustering methods to create clusters of UK local
authorities with similar temporal signatures of experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore how they overlap with or diverge from
the existing economic and digital geography of the UK. 
Our analysis enables us to better understand how the spatial and social distribution
of both occupations and online accessibility intersect to enable or hinder the
practice of telecommuting at a time of extreme demand.
We will also consider what lessons can be learned from this time for a future where
telecommuting is likely to remain a common means of accessing work, and broadband
services and infrastructure must be fit for purpose.
**LET'S LEAVE IT FOR NOW, BUT I THINK WE CAN CRYSTALISE MORE THE RQ**

<!-- capability to WFH -->
The capability to work from home has previously been studied from the
perspective of whether work tasks in a given occupation both can be and are
allowed to be performed using digital technologies independently of
location or co-location with colleagues, including supervisors
[@allen2015effective; @singh2013modeling].
However, successful telecommuting also requires that the quality and
reliability of digital services, particularly home internet connection speeds,
enable the completion of work tasks with a minimum of delay or interruption.
Prior to the pandemic, the performance of broadband services with respect to
telecommuters was never tested at scale, as working from home and connecting to
colleagues and workplace resources via the internet was the purview of a small minority of workers.
Instead, leisure use in the evening, when video streaming services are at their peak,
has been used to benchmark broadband performance and service delivery by different
Internet Service Providers (ISPs), at least in the UK [@ofcom2017].
Yet the shift towards telecommuting during various stages of lockdown around the world
has been drastic and there are speculations that post-Covid, the tendency to work from
home will be much higher than pre-Covid, raising questions around whether internet
services can accommodate the increased demand.
For example, $47$% of people in employment in the UK worked soley from home in April
$2020$, whilst the same figure only reached $5$% the year before [@ons2020;
@ons2020lm2019].
A back of the envelope calculation suggests that up to 40% of the working force
could work from home on an ongoing basis [@batty2020editorial].
Similar figures have been reported for other countries [@felstead2020homeworking].
For instance, $37$% of the European workforce worked from home in April $2020$ with
countries like Finland reaching $60$% [@eurofound2020].
In the US, almost half of the working population worked from home during the same
period because of the pandemic [@brynjolfsson2020covid],
and a recent estimate indicated that $37$% of all jobs in the US can be permanently
performed entirely from home [@NBERw26948].

<!-- ICT infrastructure and digital divides-->
None of these changes could have happened in the absence of reliable information
and communication technology (ICT) infrastructure -- both in terms of software
and hardware.
But while software innovations are easily diffused across space and
society^[See for example the huge success of videoconferencing apps such as Zoom
[@marks2020zoom].],
the same does not apply for ICT hardware infrastructure such as internet
broadband connectivity.
The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest in
different geographies in the UK [@riddlesden2014broadband; @philip2017digital].
Second level digital divides consider the presence or lack of the necessary skills
to effectively utilise digital technologies and the internet
[@blank2014dimensions; @van2011internet].
The third level focuses on the heterogenous returns of internet usage among
different socioeconomic groups and, consequently, how digital technologies can
assist in bridging or further enhancing existing socioeconomic divides.
[@stern2009levels; @van2014digital; @van2015third].
The capability to telecommute is related to all three levels of digital divides,
but more importantly leads to differentiated outcomes regarding the economic
resilience of people and places to overcome a systemic shock such as the current
pandemic.
We identify this as a new digital divide, one that fundamentally alters the
potential returns of internet use for the user and wider community,
assumes skills or functions that are present in some occupations but not in others,
and relies upon access to high quality internet services.
**WE NEED TO CONNECT OUR FINDINGS WITH THIS AND JUSTIFY THAT THIS IS INDEED A NEW DIVIDE**
As the quality of internet infrastructure and services, as well as variation in
occupations are spatially dependent and clustered in space, our approach offers a
framework for understanding which types of places, are more likely to land on the
right side of this new digital divide.
By asking how resilient broadband speeds are as experienced in different parts of
the UK during a time of extreme demand, we interrogate which places benefit from
the greater economic resilience digital technologies can offer, not only during
the pandemic, but also into the future.

The structure of this paper is as follows.
First we review the literature on telecommuting and digital divides to better
understand the origins of the new digital divide revealed by the pandemic and its
impact on the economic resilience of different places.
We then describe our data and methodology.
Our results section first offers a classification of how internet services vary
across the UK local authorities and then assesses whether these clusters replicate
or repudiate other socio-economic and geographic patterns of economic resilience.

# Literature review

## From telecommuting to #WFH

<!-- Who works from home-->
In this analysis, the terms ‘telecommuting’ and ‘working from home’ are used
interchangeably, as most remote labour during the Covid-19 crisis was carried out
in the homes of individual employees rather than any other location [@eurofound2020].
However, it should be noted that previous research has explored how telecommuting
can occur in other places, including satellite offices or on public transport
[@felstead2012rapid; @siha2006telecommuting].
Previous research has also used a variety of definitions to measure the level of
telecommuting within different workforces, distinguishing, for example, between those
directly employed, indirectly employed, self-employed, full-time or part-time, and
those who use digital technologies to work remotely full-days or part-days
[@allen2015effective; @bailey2002review; @haddad2009examination].
No matter the definition, the option and capability to telecommute or work from home
has never been equally distributed spatially or socio-economically any more than
different industries and employment opportunities have.
For example, studies from the United States, the Netherlands, and the UK found
that telecommuters are most likely to hold professional, managerial, and technical
occupations where the workforce is better educated and wealthier, and that there is
suppressed demand among women and part-time workers
[@headicar2016move; @peters2004employees; @singh2013modeling].
Opportunities for working from home during the current pandemic have likewise not
been equally spread across the workforce.
@NBERw26948 indicated that in the US, managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from home,
and that occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in more spatially fixed occupations,
from farming, construction and manufacturing to hospitality and care services.
In the US, these occupations tend to be lower-income, non-white,
without a university degree, live in rental accommodation and lack health insurance
[@NBERw27085].
Similar trends can be observed for other countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, whilst only $34$% of workers with secondary education and
$14$% of those primary education did so [@eurofound2020].

## Digital divides and economic resilience
<!-- WFH and digital divides-->
Our understanding of telecommuting as a product of enabled occupations can be
described as a manifestation of the third level digital divide, as those who are able
to use digital technologies to work from home benefit from a high rate of return on
their use of the internet in terms of autonomy, flexibility, and time saved from
commuting [@peters2004employees; @siha2006telecommuting; @singh2013modeling].
These returns have been even greater during the Covid-19 crisis,
when those with the ability to telecommute also have the ability to maintain
their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the first
level digital divide, which is associated with access and quality of internet
connectivity at a time of extremely high demand.
@SALEMINK2017360 provides a systematic review of the pre-pandemic, 
first level digital divide in infrastructure quality between urban and rural areas 
in various advanced economies.
Yet whether this variation in infrastructure quality affects the spatial footprint of
telecommuting has not previously been measured.
There are indications that those who purchase high speed connections consume 
more data of all sorts and use their connections for a variety of purposes 
[@hauge2011consumer], 
and that there is a correlation between access to internet services and a
reduction in household transport spend [@bris2017ict].
Whether the implication is more internet use and less travel 
because of increased telecommuting, these studies suggest that better internet 
services enable households to make savings and efficiencies, 
an example of the first level digital divide reinforcing the third level.
Yet did the purchase of high speed connections and increased internet access also  
prepare households for long-term home-working, enforced by government restrictions?
The extreme demand during the pandemic provides a new opportunity to understand 
how infrastructure accessibility, quality, and reliability affects telecommuting, 
particularly in light of the high volumes of bandwidth-intensive video conferencing 
required in order to avoid the face-to-face contact that could increase 
the spread of infection. 
We seek to answer how internet service resilience contributes to or reduces 
economic resilience 
when the latter is dependent upon the capability to work from home, 
and thus improve our understanding of this new digital divide.
**RELATED TO YOUR POINT ABOVE, WE MAY CONCLUDE THIS IS NOT A NEW DIGITAL DIVIDE, BUT THAT IS STILL ANSWERING A USEFUL QUESTION - MIGHT END UP REWORDING AS INTERACTION BETWEEN DIGITAL DIVIDES?**

<!-- Covid-19 and working from home, cities, urban structure -->
The multi-layered digital divides intersect with material divides and the economic
geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south part of the UK regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only are all three levels of digital divides associated to a certain extent 
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome
some of the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

## other subsection?
**I THINK WE HAVE ENOUGH OF A LIT REV**
- circumstances where work can be carried out more flexibly in space and time
- how relate to digital divides
- capability theory of mobility?
- broadband tech stuff? broadband studies / divides / resilience

Some new papers google recommended to me:

- https://urbanstudies.uva.nl/binaries/content/assets/subsites/centre-for-urban-studies/working-paper-series/wps_43.pdf
- https://link.springer.com/article/10.1007/s11116-020-10136-6
- https://www.sciencedirect.com/science/article/pii/S0966692319311305
- check who cites the above and what they cite

<!-- How can we observe observe WFM? -->
<!-- removed for now -->
<!-- Directly through surveys (@felstead2020homeworking).
Add survey limitations (cost, representativeness, time lag between survey and data availability)
Data gap.
Passive data collection through internet speeds
Another option could be mobility data.
How is internet speed data a better source than mobility?-->

**MORE SOURCES:**

- https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas
- [EPB commentaries](https://journals.sagepub.com/toc/EPB/current)
- https://www.coronavirusandtheeconomy.com/question/what-has-coronavirus-taught-us-about-working-home
- https://www.coronavirusandtheeconomy.com/question/who-can-work-home-and-how-does-it-affect-their-productivity
- https://www.coronavirusandtheeconomy.com/question/how-will-economic-effects-coronavirus-vary-across-areas-uk
- https://www.coronavirusandtheeconomy.com/which-parts-uk-have-been-hit-hardest-covid-19-crisis
- https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas

# Methods and data

<!-- clustering in general -->
The starting point of our methodological framework is cluster analysis, which can
be defined within the modern machine learning framework as an unsupervised
learning task, partitioning unlabelled observations into homogeneous groups 
known as clusters [@montero2014tsclust].
The key idea is that observations within clusters tend to be more similar than
observations between clusters.
Clustering is particularly useful for exploratory studies as it identifies
structures within the data [@aghabozorgi2015time].
Therefore, cluster analysis is a widely used family of techniques in geography
[@gordon1977classification; @everitt1974cluster].
For instance, clustering methods are the basis of *geodemographics*, a discipline 
which aims to create small area indicators or typologies of neighbourhoods 
based on various and sometimes diverse variables 
[@SINGLETON2009289; @harris2005geodemographics].
Clustering techniques have also been employed to solve regionalisation problems
[@niesterowicz2016].

<!-- time-series clustering -->
Common characteristics of these studies are the cross-sectional nature of the data
they employ.
Indeed, most clustering problems in geography deal with observations that are
fixed in time.
However, for this paper we are interested in internet speeds, which vary over time. 
We aim to create clusters of local authorities in the UK with similar temporal 
signatures of experienced internet speeds.
Hence, we deviate from the established geographical clustering tools and
employ time-series clustering methods.

Time-series clustering methods have been developed in order to deal with clustering
problems linked to, for instance, stock or other financial data, economic,
governmental or medical data as well as machine monitoring
[@aggarwal2013time; @aggarwal2001surprising; @hyndman2015large; @WARRENLIAO20051857].
The main challenge -- and also the difference with cross-sectional clustering
problems -- is data dimensionality given the multiplicity of data points for every
individual object -- local authority in our case -- included in the data set.
Time-series are dynamic data as the value of the observations change as
a function of time [@aghabozorgi2015time].
This high dimensionality leads to
(i) computational and algorithmic challenges regarding handling these data and
building algorithms to perform clustering over long time-series, and 
(ii) open questions regarding the choice of similarity measures in order to cluster 
similar times series objects together considering the whole length of the time-series
and overcoming issues around noise, outliers and shifts [@lin2004iterative;
@aghabozorgi2015time].

Time-series clustering methods utilising the whole length of time-series can be
grouped in three categories.
The first -- model-based approaches -- is based on recovering the underlying model
for each time-series and then applying clustering algorithms on the model parameters
of each time-series [@aghabozorgi2015time].
The main criticism is the cluster accuracy for nearby clusters [@mitsa2010temporal].
The second approach is based the formation of vectors of features derived from the
original time-series.
These new data of reduced dimensionality are then clustered using conventional
clustering algorithms.

For this paper we utilise the third category of time-series clustering methods
known as shape-based approaches.
These methods match two sperate time-series objects based on the similarity of
their shapes through the calculation of distances between the shapes, 
and are thus better equipped to capture similarities between short length
time-series [@aghabozorgi2015time].
This approach serves best this paper because (i) we aim to identify clusters of
UK local authorities with similar temporal signatures -- i.e. shapes -- of
experienced internet speeds and (ii) the length of our time-series is short
(see the data discussion in this section).

Another import element of time-series clustering is the actual clustering algorithm.
Similar to the clustering of cross-sectional data, we can employ partitioning
algorithms, which lead to non-overlapping clusters, hierarchical clustering, 
which classifies clusters at different levels, and fuzzy algorithms, which
create overlapping clusters [@sardatime].
Because of the simplicity of the implementation and the interpretability of the
results, we utilise here partitioning clustering based on the widely used *k-Means*
algorithm.
This is an iterative algorithm, which begins with defining the desired number of
clusters *k*.
Then each observation is randomly assigned to a cluster from the $[1,k]$ space.
This initial cluster assignment is followed by iterations in order to minimise 
the distance between the centroids of the clusters and the observations
assigned to these clusters [@james2013introduction].

<!-- difference between clustering cross-sectional and time-series data -->
There are a number of differences between the above described application of
*k-Means* for cross-sectional data and its application for times series data.
Instead of creating clusters around centroids, a common approach is to create
clusters around *medoids*, which are representative time-series objects with a
minimal distance to all other cluster objects [@sardatime].
Also, instead of calculating the Euclidean distance between centroids and data
points, more complex distance measures need to be employed in order to capture
the similarity between a time-series object and a medoid.
A common distance measure for shape-based time-series clustering is 
Dynamic Time Warping (DTW).
Using its underpinning dynamic programming algorithm, DTW compares two time-series
objects to find the optimum warping path between them.
DTW is widely used in order to overcome limitations linked to the use of Euclidean
distance [@sardatime; @berndt1994using; @ratanamahatana2004everything].
The `R` package `dtwclust` has been used for the time-series clustering [@dtwclust].

<!-- introduction of our broadband speeds data -->
To assess the quality and reliability of internet across local authorities
in the UK during the time when the population were told to work from home if at
all possible we utilise unique data comprising individual internet speed tests from
Speedchecker Ltd^[[https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)].
This a private company that allows internet users to check their own broadband
upload and download speeds, and stores every speed-check with timestamp and
geolocation information.
These data have been used before to assess digital divides [@riddlesden2014broadband]
and the impact of local loop unbundling regulatory processes [@nardotto2015unbundling].
These volunteered geographic data enable to assess the *experienced* internet
speed by users, which may differ from the *advertised* maximum speeds provided
by Internet Service Providers.
We are particularly interested in upload speeds and the frequency of speed tests.
Upload speeds are less associated with internet-based leisure activities such
as video streaming, which involve downloading large amounts of data, but they
can be linked more with work-related activities such as uploading documents or
two-way communication and interaction systems.
Meanwhile, the frequency of speed tests was important in identifying 
the temporal profile which woud give us most insight into 
experienced internet service and resilience.
Although we know from government statements the start and end of the period of
interest, namely, March to May 2020, the frequency of speed tests within that period 
provides an indication of the volume of experience over any particular units of time.

<!-- data transformation -->
Thus, the first step in the workflow after dropping some outliers following
@riddlesden2014broadband was to transform the individual, geolocated and time-stamped 
tests to more meaningful aggregates both in terms of space and time.
The frequency of testing indicates that whilst there is an overall trend of increased
testing or increased experience of internet service from March to April and then a 
slight reduction from April to May, this trend masks substantial variation 
by not only the day of the week, but also time of day, as can be seen in **[graph]**.
Thus, a daily aggregation of upload speeds would also mask this pattern.
Furthermore, the importance of this pattern is highlighted by a comparison 
with the same period in 2019 **see graph**, 
when the volume of testing and thus of experience of internet service quality
was much more tied to the evening leisure peak and presumably to download speeds.
Since the increase in testing during the working day in 2020 is an indication that 
users have a greater perception of the variable quality of internet service,
particularly during a new morning peak of testing for service reliability,
we decided to include a measure of hourly variation in our temporal profiles.
However, although this is a large data set ($241,088$ individual tests during the
working hours of study period), there are not enough observations for each
Local Authority District (LAD) and for each working hour of each working day
($631$ speed tests per LAD on average) to profile speeds at that level of detail.
Therefore, we aggregate all the speed-checks during the $13$ weeks of March to May 
inclusive for weekdays in $2020$ by each hour of the day and day of the week.
As our research aims to identify the geography of internet service resilience for
work purposes, bank holidays and the hours between midnight and 6am were excluded,
as well as weekend days.
The composite week time-series thus comprise 18 hours multiplied by 5 weekdays or
90 time points per series.
We also aggregate these data spatially because we could not follow individuals or 
households and connect data points.
The time-series were calculated for each of the 382 LADs in the UK, standardised, 
and then a *k*-means partitioning around medoids clustering algorithm 
was applied using DTW.
We initially run the algorithm for $k = 5, 10, 15 and 20$ and used cluster validity
indices (CVIs) to pick the optimal solution of $k = 10$.
Following @sardatime the following CVIs were employed: Silhouette (max), Score
function (max), Calinski-Harabasz (max), Davies-Bouldin (min), Modified Davies-Bouldin (DB*, min), Dunn (max), COP (min).

```{r echo=FALSE, eval=FALSE, message=FALSE}

# Code to calculate the above figures about N.

#upload data file
path.data <- paste(path, "./data/temp/TSbb19_20sp.csv", sep = "")
TSbb19_20sp <- read.csv(path.data)
TSbb20sp <- TSbb19_20sp %>%
  filter(dateOnly> "2019-12-31")

# N. of observations for study period  
print(nrow(TSbb20sp))

# observations for study period per local authority
print(nrow(TSbb20sp)/length(unique(TSbb20sp$LAD19NM)))
```

<!-- metrics -->
We then review the characteristics of each cluster, including number of LADs, 
and the temporal profile of upload speed by hour of the day 
and day of the composite week.
Since the quality and reliability of internet services vary in time and space due
to both supply and demand-side influences, we use a number of different measures
to then describe experienced upload speeds per cluster.
These include: a) mean, experienced connection speed,
b) standard deviation or the amount of fluctuation from the mean, and
c) the variation in speeds at particular times of day when working from home is
more likely to take place.
We take account of all three measurements in our descriptives statistics of
upload speeds in order to determine how resilient broadband speeds are 
as experienced in different parts of the UK during a time of extreme demand.

**Data details and some figures, descriptive stats - include whole sample time profile for 2019 and 2020 frequency of tests run as part of why we chose to create the time profiles by hour of the day and day of the week rather than daily over the whole period.**

**Just looking back, it was partly the frequency of tests, but that was also by geography, so do we need maps? And I feel there was something else I'm forgetting...**

<!-- HB old PARA4 - link to auxillary regressions method, or put in results to introduce results of auxillary regressions? -->
The cause of these different experiences of broadband resilience may be different in
different areas, as they may reflect either similarities in patterns of demand
or similar quality of infrastructure.
Our approach is also limited by potential endogeneity, as for example, better
quality connections with high mean speeds may enable more working from home, but
greater demand may cause slower speeds, less reliability and greater variability
of speed at different times of day or week.
Therefore, we avoid attributing any cause to our analysis of the experienced level
of quality and reliability of upload speeds.
Instead, we run an auxiliary regression in order to understand how the spatial and
temporal patterns of internet service relate to the economic geography of the UK.
We discuss how the different patterns might support or undermine efforts to work
from home and maintain safe productivity and whether they reinforce existing
spatial and social inequalities.
More specifically, we estimate the following multinomial model:

\begin{align}
Pr(Y_{i}=j) = \frac{exp^{X_{i}\beta_{j}}}{\sum_{i=1}^j exp^{X_{i}\beta_{j}}}
\begin{cases}
    i = 1, 2, ... , N \\t
    j = 1, 2, ... , J
\end{cases}\label{eq1}
\end{align}

The identification strategy is as follows.
Based on the outcomes of the time-series clustering, we identify $J$ distinct
and crisp clusters.
We then regress this cluster membership against a vector $X_{i}$ of
socio-economic and geographic variables, which are discussed in detail in
**section XX**.
This analysis enables us to provide a more nuanced understanding of how
telecommuting and technology intersect at a time of extreme demand, and what
lessons this time has for a future where telecommuting is likely to remain a
common means of accessing work and broadband services, as well as infrastructure,
must be fit for purpose.

# Results

## Upload Clusters / cluster description

The temporal profiles used to cluster the local authorities have been summarised in
graph [.], which
shows a composite profile of mean upload speeds per hour per day for each cluster.
For upload speeds, 345 of 382 local authorities, or over 62 million people,
fall into cluster 6 or cluster 9.
Graph [.] suggests that these two clusters have relatively similar temporal profiles,
which are flatter than the other, smaller clusters,
suggesting better reliability of service.
However, the upload speeds at all times for cluster 9 are substantially higher
than for cluster 6, which is an indication of better quality of service.
This difference is reflected not only in the mean speeds for these clusters for the
whole sample, but also the mean upload speeds in the morning peak from 9:00-10:59,
as well as the evening peak period from 19:00-20:59.
In comparison, the time profile in graph [.] shows upload speeds in cluster 1 are
on average lower at certain times of day during the study period than any other
cluster, whilst the profile for cluster 3 appears to show speeds fluctuating as much
as cluster 1, but at levels usually higher even than cluster 9.

The variability of the smaller clusters may be related to the fewer speed tests from
fewer local authorities that have been averaged, whilst averaging greater numbers of
speed tests could artificially flatten the profile.
This appears not to be the case for cluster 9, as there is negligible difference
between morning and evening upload speeds, at 1% slower in the morning, confirming a
high level of reliability of service.
In comparison, upload speeds in cluster 6 are 4% slower in the morning than in the
evening and experienced the joint highest ratio of standard deviation to mean across
the time period under assessment.
This suggests that although the time profile is relatively flat in graph [.],
the experience is one of speeds that fluctuate from a lower mean,
and therefore might more often impact on online activities.
Still, in most of the smaller clusters the reliability of service is worse than in
cluster 9 in terms of the ratio of standard deviation to mean.
The smaller clusters are mainly worse than clusters 9 and 6 in terms of
the ratio of upload speed in the morning peak compared to the evening peak.
In five of the smaller clusters, which are home to almost 3 million people,
speeds are 12-19% slower between 9:00-10:59 compared to 19:00-20:59.

The quality and reliability of broadband service is much better in
the 115 Local Authority areas in cluster 9,
which are mostly in urban or suburban areas, compared to cluster 6.
These include 13 London Boroughs (of 32), 8 of the 10 local authorities of Greater Manchester, 5 of the 7 constituent authorities of the West Midlands Combined Authority, as well as cities like Glasgow, Leicester, Nottingham, Sheffield, and the Portsmouth and Southampton conurbation. There are also some notable medium-sized cities, including Aberdeen, Cardiff, Oxford, Milton Keynes, and York, and many suburban districts from the South East of England to South Tyneside. Meanwhile, the 230 local authorities in cluster 6, which have lower speeds on average and more variation in service still include major urban areas, such as Bristol, Liverpool and Leeds, and many suburban areas, but also include some of the most rural areas in the country. Meanwhile, Cluster 1, with 10 local authorities that are home to over 1 million people has the second slowest speeds in the morning compared to the evening 'peaks' and the second highest ratio of standard deviation to mean. This cluster's most populous area is Westminster in central London.

**further description of temporal profiles of other clusters - include all in graph?** Indeed, the only cluster where upload speeds were slower in the evening than in the morning was cluster 5, made up of 2 local authorities with less than 200,000 people: Three Rivers, a suburban district north of London, and Fylde, a seaside suburb of Blackpool. However, these are likely to be outliers and may not have many tests from which the clusters are calcualted. **Take out smaller clusters?** The exceptions can be found in seven of the eight much smaller clusters including 35 local authority districts, where AM peak upload speeds are between 6% and 18% slower than PM peak upload speeds, although the mean speeds for each cluster are higher than cluster 6. Indeed, in 25 local authorities with a combined population of almost 3 million, speeds are 13% or more slower in the morning than in the evening. Included in this latter group are central London borough of Westminster and the London Borough of Newham, rural authorities like Eden and West Devon, and small cities like Dundee and Carlisle.

## aux regressions

Auxiliary regressions indicate that the speed tests in cluster 9 authorities are also more likely to have been run on services provided by Virgin Media, suggesting they are in the half of the country with the most lucrative ICT market, which originally attracted the cable TV provider (OfCom...). For example, although auxiliary regressions show that Cluster 6 local authorities are more likely to be in the South of the UK than Cluster 9, the cluster notably includes Southern rural districts from Cornwall to North Norfolk. Slower speeds could reflect the lower quality of service in rural areas compared to urban and suburban areas. There is frustration, however, as auxiliary regressions show that those living in cluster 6 had the highest probability of testing their broadband between 9:00 and 11:00 of any of the 10 upload speed clusters.

the finding in the auxiliary regressions that those in cluster 9 ran the fewest speed tests per person during the morning period of any cluster. Now this may be an indication of fewer people working from home, less contention, and less resultant frustration. Cluster 9 comprises many central urban areas and has the lowest number of established businesses per inhabitant, which could be interpreted as a dominance of large employers. However, the job density is lower than cluster 6, meaning there are not as many jobs per resident in these areas. As the cluster includes many suburban areas too, which may be largely residential, could the quality and reliability of internet service be reinforcing patterns of telecommuting by those in wealthier suburbs who can work from home? Earnings in cluster 9 are second highest of all the clusters, with only Cluster 2 (comprising just North East Lincolnshire and East Lothian, population 265k) earning more per person.

The auxiliary regressions show that compared to the two authorities in cluster 5, all the other clusters had a lower percentage of working people in managerial, professional and administrative jobs.??? Yet the auxiliary regression suggests that there are not many tests being run during the am peak in cluster 1. This may be because there are fewer people working at home checking their broadband than in most other clusters. Indeed, the auxiliary regressions indicate that cluster 1 has the highest job density or proportion of jobs to working-age population, which is likely to due to the presence of Westminster, central London, cluster 1's most populous local authority. Westminster not only has more workplaces than residents, but it is reasonable to presume that many who would normally work in Westminster, but be able to work from home during lockdown are likely to live outside central London and not be subject to the fluctuating speeds there. Workplaces, meanwhile, some of which would still have been open, could be running programmes that cause the slowdown and variation, but would be more likely to have their own in-house diagnostics, rather than using a service like Speedchecker Ltd.

# Conclusions

<!-- WFH broader consequences-->
The long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they span through various aspects of economy and society: from changes to
transportation planning due to altered commuting patters, to changes in land
use and urban planning to accommodate people who work from home
[@BUDNITZ2020102713]**also 2020 Swedish article from JTG**; and from productivity and innovation changes, to changes
in agglomeration externalities and the attraction of large cities [@econobs]
just to name a few. This paper is positioned to support endeavours in
understanding the effects of increased telecommuting by
exposing the spatial and social dimensions of telecommuting including the
resilience of broadband speeds in terms of both quality and reliability of
service, and whether this reinforces or redresses prior digital divisions.
**took this bit you wrote to put in the discussion at the end?**

# Acknowledgement(s) {-}

An unnumbered section, e.g.\ `\section*{Acknowledgements}`, may be used for thanks, etc.\ if required and included _in the non-anonymous version_ before any Notes or References.

# Funding {-}

An unnumbered section, e.g.\ `\section*{Funding}`, may be used for grant details, etc.\ if required and included _in the non-anonymous version_ before any Notes or References.

# References