### 2020-12-22 paper/v1_sage/broadband.speed.covid.Rmd

---
title: WFH and broadband speed (title needs rework)
author:
- name: A. Uthor*
  num: 1,2
- name: O. Tro
  num: 2
- name: O. Vriga
  num: 3
address:
- num: 1
  org: Department of Incredible Research, University A, City A, Country A
- num: 2
  org: Department of Applied Things, University B, City B, Country B
- num: 3
  org: Very Important Stuff Committee, Institute C, City C, Country C
corrauth: "Corresponding author name, This is sample corresponding address."
email: correspondingauthor@protonmail.com
abstract: "TBC"
keywords:   covid; internet; working from home; broadband speed; time-series clusters;
classoption:
  - Royal
  - times
bibliography: bibliography
bibliographystyle: sageh
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{dcolumn}
  \def\tightlist{}
  \newcommand{\beginappendix}{\setcounter{table}{0} \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
output:
  rticles::sage_article:
    keep_tex: yes
---

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)
library(stargazer)
library(png)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction {#sec:1}

<!-- [our google doc](https://docs.google.com/document/d/1PWjkmgzWGYKR9wFogKYw7l-8mZLoORt593x-Tu-f2-M/edit#heading=h.i5om1o8wpcd9) -->

During the pandemic, working from home using digital technologies, whether partially
or exclusively, was transformed from a niche means of accessing work, albeit one that
had been on a slow, upward trend, to a widespread way of life in many countries.
The ability to work from home or telecommute meant millions retained their jobs and,
to a varying extent, maintained productivity during periods of strict
lockdowns around the world.
However, this ability has not been evenly distributed socially or spatially,
creating a new type of digital divide.
On one side are those who can work from home, supported by digital technologies,
and have thus been able to enjoy both economic resilience and greater personal safety.
On the other side, previously employed individuals have been forced to accept
furlough or redundancy packages unless they are part of the cadre of essential
workers, who are potentially at high risk of infection.
Whilst the basis for this new digital divide has been viewed as mainly occupational,
here we consider whether the divide is also technological.

<!-- our approach -->
Using the UK as a case study, this paper aims to understand how the quality and
reliability of internet service, as reflected in *experienced* internet speeds,
may reinforce or redress the spatial and social dimensions of the digital division
exposed by the pandemic.
To do so, we employ volunteered geographic data on individual broadband speed tests
and state-of-the-art time-series clustering methods to create clusters of UK local
authorities with similar temporal signatures of experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore how they overlap with or diverge from
the existing economic and digital geography of the UK.
Our analysis enables us to better understand how the spatial and social distribution
of both occupations and online accessibility intersect to enable or hinder the
practice of telecommuting at a time of extreme demand.
We will also consider what lessons can be learned from this time for a future where
telecommuting is likely to remain a more common means of accessing work,
at least in comparison to the pre-Covid era,
and broadband services and infrastructure must be fit for purpose.
**LET'S LEAVE IT FOR NOW, BUT I THINK WE CAN CRYSTALISE MORE THE RQ**

<!-- capability to WFH -->
The capability to work from home has previously been studied from the
perspective of whether work tasks in a given occupation both can be and are
allowed to be performed using digital technologies independently of
location or co-location with colleagues, including supervisors
[@allen2015effective; @singh2013modeling].
However, successful telecommuting also requires that the quality and
reliability of digital services, particularly home internet connection speeds,
enable the completion of work tasks with a minimum of delay or interruption.
Prior to the pandemic, the performance of broadband services with respect to
telecommuters was never tested at scale, as working from home and connecting to
colleagues and workplace resources via the internet was the purview of a small minority of workers.
Instead, leisure use in the evening, when video streaming services are at their peak,
has been used to benchmark broadband performance and service delivery by different
Internet Service Providers (ISPs), at least in the UK [@ofcom2017].

Yet the shift towards telecommuting during various stages of lockdown around the world
has been drastic and there are speculations that post-Covid, the tendency to work from
home will be much higher than pre-Covid, raising questions around whether internet
services can accommodate the increased demand.
For example, $47$% of people in employment in the UK worked solely from home in April
$2020$, whilst the same figure only reached $5$% the year before [@ons2020;
@ons2020lm2019].
A back of the envelope calculation suggests that up to 40% of the working force
could work from home on an ongoing basis [@batty2020editorial].
Similar figures have been reported for other countries [@felstead2020homeworking].
For instance, $37$% of the European workforce worked from home in April $2020$ with
countries like Finland reaching $60$% [@eurofound2020].
In the US, almost half of the working population worked from home during the same
period because of the pandemic [@brynjolfsson2020covid],
and a recent estimate indicated that $37$% of all jobs in the US can be permanently
performed entirely from home [@NBERw26948].

<!-- ICT infrastructure and digital divides-->
None of these changes could have happened in the absence of reliable information
and communication technology (ICT) infrastructure -- both in terms of software
and hardware.
But while software innovations are easily diffused across space and
society^[See for example the huge success of videoconferencing apps such as Zoom
[@marks2020zoom].],
the same does not apply for ICT hardware infrastructure such as internet
broadband connectivity.
The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest in
different geographies in the UK [@riddlesden2014broadband; @philip2017digital].
Second level digital divides consider the presence or lack of the necessary skills
to effectively utilise digital technologies and the internet
[@blank2014dimensions; @van2011internet].
The third level focuses on the heterogenous returns of internet usage among
different socioeconomic groups and, consequently, how digital technologies can
assist in bridging or further enhancing existing socioeconomic divides.
[@stern2009levels; @van2014digital; @van2015third].
The capability to telecommute is related to all three levels of digital divides,
but more importantly leads to differentiated outcomes regarding the economic
resilience of people and places to overcome a systemic shock such as the current
pandemic.
The extreme level of demand for telecommuting fundamentally alters the
potential returns of internet use for the user and wider community,
assumes skills or functions that are present in some occupations but not in others,
and relies upon access to high quality internet services.

As the quality of internet infrastructure and services, as well as variation in
occupations are spatially dependent and clustered in space, our approach offers a
framework for understanding the impact of and interactions between
the different levels of digital division in different places
with different characteristics.
By asking how resilient broadband speeds, and particularly upload speeds are
as experienced in different parts of the UK during a time of extreme demand,
we interrogate which places benefit from the greater economic resilience
digital technologies can offer, not only during the pandemic,
but also into the future.
The structure of this paper is as follows.
First we review the literature on telecommuting and digital divides to better
understand the structural and spatial development of these practices pre-pandemic,
and thus their importance to the economic resilience of different places.
We then describe our data and methodology.
Our results section first offers a classification of how internet services vary
across the UK local authorities and then assesses whether these clusters replicate
or repudiate other socio-economic and geographic patterns of economic resilience.
The paper ends with a conclusions section.

# Literature review {#sec:2}

## From telecommuting to #WFH {#sec:2.1}

<!-- Who works from home-->
In this analysis, the terms ‘telecommuting’ and ‘working from home’ are used
interchangeably, as most remote labour during the Covid-19 crisis was carried out
in the homes of individual employees rather than any other location [@eurofound2020].
However, it should be noted that previous research has explored how telecommuting
can occur in other places, including satellite offices or on public transport
[@felstead2012rapid; @siha2006telecommuting].
Previous research has also used a variety of definitions to measure the level of
telecommuting within different workforces, distinguishing, for example, between those
directly employed, indirectly employed, self-employed, full-time or part-time, and
those who use digital technologies to work remotely full-days or part-days
[@allen2015effective; @bailey2002review; @haddad2009examination].
No matter the definition, the option and capability to telecommute or work from home
has never been equally distributed spatially or socio-economically any more than
different industries and employment opportunities have.
For example, studies from the United States, the Netherlands, and the UK found
that telecommuters are most likely to hold professional, managerial, and technical
occupations where the workforce is better educated and wealthier, and that there is
suppressed demand among women and part-time workers
[@headicar2016move; @peters2004employees; @singh2013modeling].

Opportunities for working from home during the current pandemic have likewise not
been equally spread across the workforce.
@NBERw26948 indicated that in the US, managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from home,
and that occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in more spatially fixed occupations,
from farming, construction and manufacturing to hospitality and care services.
In the US, these occupations tend to be lower-income, non-white,
without a university degree, live in rental accommodation and lack health insurance
[@NBERw27085].
Similar trends can be observed for other countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, whilst only $34$% of workers with secondary education and
$14$% of those primary education did so [@eurofound2020].

## Digital divides and economic resilience {#sec:2.2}
<!-- WFH and digital divides-->
Our understanding of telecommuting as a product of enabled occupations can be
described as a manifestation of the third level digital divide, as those who are able
to use digital technologies to work from home benefit from a high rate of return on
their use of the internet in terms of autonomy, flexibility, and time saved from
commuting [@peters2004employees; @siha2006telecommuting; @singh2013modeling].
These returns have been even greater during the Covid-19 crisis,
when those with the ability to telecommute also have the ability to maintain
their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the first
level digital divide, which is associated with access and quality of internet
connectivity at a time of extremely high demand.
@SALEMINK2017360 provides a systematic review of the pre-pandemic,
first level digital divide in infrastructure quality between urban and rural areas
in various advanced economies.
Rural areas, predictably, fare worse.
Yet whether this variation in infrastructure quality affects the spatial footprint of
telecommuting has not previously been measured, in part because telecommuting has
not previously been the cause of greatest demand and pressure on internet services.

There are indications that those who purchase high speed connections consume
more data of all sorts and use their connections for a variety of purposes
[@hauge2011consumer],
and that there is a correlation between access to internet services and a
reduction in household transport spend [@bris2017ict].
Whether the implication is more internet use and less travel
because of increased telecommuting, these studies suggest that better internet
services enable households to make savings and efficiencies,
an example of the first level digital divide reinforcing the third level.
Yet did the purchase of high speed connections and increased internet access also
prepare households for long-term home-working, enforced by government restrictions?
The extreme demand during the pandemic provides a new opportunity to understand
how infrastructure accessibility, quality, and reliability affects telecommuting,
particularly in light of the high volumes of bandwidth-intensive video conferencing
required in order to avoid the face-to-face contact that could increase
the spread of infection.
We seek to answer how internet service resilience contributes to or reduces
economic resilience when the latter is dependent upon
the capability to work from home.
We also aim to improve our understanding of the impact
of first level digital division on telecommuting, and whether this results in
much more fundamental third level digital division than has previously been perceived.

<!-- Covid-19 and working from home, cities, urban structure -->
Furthermore, these multi-layered digital divides intersect with material divides
and the economic geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south regions of England regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only are all three levels of digital divides associated, to a certain extent,
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome
some of the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

<!-- ## other subsection? -->
<!-- **I THINK WE HAVE ENOUGH OF A LIT REV** -->
<!-- - circumstances where work can be carried out more flexibly in space and time -->
<!-- - how relate to digital divides -->
<!-- - capability theory of mobility? -->
<!-- - broadband tech stuff? broadband studies / divides / resilience -->

<!-- Some new papers google recommended to me: -->

<!-- - https://urbanstudies.uva.nl/binaries/content/assets/subsites/centre-for-urban-studies/working-paper-series/wps_43.pdf -->
<!-- - https://link.springer.com/article/10.1007/s11116-020-10136-6 -->
<!-- - https://www.sciencedirect.com/science/article/pii/S0966692319311305 -->
<!-- - check who cites the above and what they cite -->

<!-- How can we observe observe WFM? -->
<!-- removed for now -->
<!-- Directly through surveys (@felstead2020homeworking).
Add survey limitations (cost, representativeness, time lag between survey and data availability)
Data gap.
Passive data collection through internet speeds
Another option could be mobility data.
How is internet speed data a better source than mobility?-->

<!-- **MORE SOURCES:** -->

<!-- - https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas -->
<!-- - [EPB commentaries](https://journals.sagepub.com/toc/EPB/current) -->
<!-- - https://www.coronavirusandtheeconomy.com/question/what-has-coronavirus-taught-us-about-working-home -->
<!-- - https://www.coronavirusandtheeconomy.com/question/who-can-work-home-and-how-does-it-affect-their-productivity -->
<!-- - https://www.coronavirusandtheeconomy.com/question/how-will-economic-effects-coronavirus-vary-across-areas-uk -->
<!-- - https://www.coronavirusandtheeconomy.com/which-parts-uk-have-been-hit-hardest-covid-19-crisis -->
<!-- - https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas -->

# Methods and data {#sec:3}

## Time-Series Clustering {#sec:3.1}
<!-- clustering in general -->
The starting point of our methodological framework is cluster analysis, which can
be defined within the modern machine learning framework as an unsupervised
learning task, partitioning unlabelled observations into homogeneous groups
known as clusters [@montero2014tsclust].
The key idea is that observations within clusters tend to be more similar than
observations between clusters.
Clustering is particularly useful for exploratory studies as it identifies
structures within the data [@aghabozorgi2015time].
Therefore, cluster analysis is a widely used family of techniques in geography
[@gordon1977classification; @everitt1974cluster].
For instance, clustering methods are the basis of *geodemographics*, a research domain
which aims to create small area indicators or typologies of neighbourhoods
based on various and sometimes diverse variables
[@SINGLETON2009289; @harris2005geodemographics].
Clustering techniques have also been employed to solve *regionalisation* problems
[@niesterowicz2016].

<!-- time-series clustering -->
Common characteristics of these studies are the cross-sectional nature of the data
they employ.
Indeed, most clustering problems in geography deal with observations that are
fixed in time.
However, for this paper we are interested in internet speeds, which vary over time
and, therefore, create clusters of local authorities in the UK with similar temporal
signatures of experienced internet speeds.
Hence, we deviate from the established geographical clustering tools and
employ time-series clustering methods.

Time-series clustering methods have been developed in order to deal with clustering
problems linked to, for instance, stock or other financial data, economic,
governmental or medical data as well as machine monitoring
[@aggarwal2013time; @aggarwal2001surprising; @hyndman2015large; @WARRENLIAO20051857].
The main challenge -- and also the difference with cross-sectional clustering
problems -- is data dimensionality given the multiplicity of data points for every
individual object, local authorities in our case, included in the data set.
Time-series are dynamic data as the value of the observations change as
a function of time [@aghabozorgi2015time].
This high dimensionality leads to
(i) computational and algorithmic challenges regarding handling these data and
building algorithms to perform clustering over long time-series, and
(ii) open questions regarding the choice of similarity measures in order to cluster
similar times series objects together considering the whole length of the time-series
and overcoming issues around noise, outliers and shifts [@lin2004iterative;
@aghabozorgi2015time].

<!-- Time-series clustering methods utilising the whole length of time-series can be grouped in three categories. The first -- model-based approaches -- is based on recovering the underlying model for each time-series and then applying clustering algorithms on the model parameters of each time-series [@aghabozorgi2015time]. The main criticism is the cluster accuracy for nearby clusters [@mitsa2010temporal].The second approach is based the formation of vectors of features derived from the original time-series. These new data of reduced dimensionality are then clustered using conventional clustering algorithms. -->

For this paper we utilise a category of time-series clustering methods
known as shape-based approaches.
These methods match two sperate time-series objects based on the similarity of
their shapes through the calculation of distances between the shapes,
and are thus better equipped to capture similarities between short length
time-series [@aghabozorgi2015time].
This approach serves best this paper because (i) we identify clusters of
UK local authorities with similar temporal signatures -- i.e. shapes -- of
experienced internet speeds and (ii) the length of our time-series is short
(see the data discussion in this section).

Another import element of time-series clustering is the actual clustering algorithm.
Similar to the clustering of cross-sectional data, we can employ partitioning
algorithms, which lead to non-overlapping clusters, hierarchical clustering,
which classifies clusters at different levels, and fuzzy algorithms, which
create overlapping clusters [@sardatime].
Because of the simplicity of the implementation and the interpretability of the
results, we utilise here partitioning clustering based on the widely used *k*-means
algorithm.
This is an iterative algorithm, which begins with defining the desired number of
clusters *k*.
Then each observation is randomly assigned to a cluster from the $[1,k]$ space.
This initial cluster assignment is followed by iterations in order to minimise
the distance between the centroids of the clusters and the observations
assigned to these clusters [@james2013introduction].

<!-- difference between clustering cross-sectional and time-series data -->
There are a number of differences between the above described application of
*k*-means for cross-sectional data and its application for times series data.
Instead of creating clusters around centroids, a common approach is to create
clusters around *medoids*, which are representative time-series objects with a
minimal distance to all other cluster objects [@sardatime].
Also, instead of calculating the Euclidean distance between centroids and data
points, more complex distance measures need to be employed in order to capture
the similarity between a time-series object and a medoid.
A common distance measure for shape-based time-series clustering is
Dynamic Time Warping (DTW).
Using its underpinning dynamic programming algorithm, DTW compares two time-series
objects to find the optimum warping path between them.
DTW is widely used in order to overcome limitations linked to the use of Euclidean
distance [@sardatime; @berndt1994using; @ratanamahatana2004everything].
The `R` package `dtwclust` has been used for the time-series clustering [@dtwclust].

## Experienced Broadband Speeds {#sec:3.2}
<!-- introduction of our broadband speeds data -->
To assess the quality and reliability of internet across local authorities
in the UK during the time when the population were told to work from home if at
all possible we utilise unique data comprising individual internet speed tests from
Speedchecker Ltd^[[https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)].
This is a private company that allows internet users to check their own broadband
upload and download speeds, and stores every speed-check with timestamp and
geolocation information.
These data have been used before to assess digital divides [@riddlesden2014broadband]
and the impact of local loop unbundling regulatory processes [@nardotto2015unbundling].
By using volunteered geographic data, we are able to assess the *experienced*
internet speed by users, which may differ from the *advertised* maximum speeds of ISPs.

We are particularly interested in upload speeds and the frequency of speed tests
over the period from March to May $2020$, as government statements indicate this is
when UK workers were told to work from home if at all possible [@GovUK2020].
Average upload speeds are slower than average download speeds, at $9.3$Mb/s
mean upload speed for the whole sample during the period of interest,
compared to $29.6$Mb/s for download speeds, but they are also less associated
with internet-based, high-demand, leisure activities such as video streaming.
Therefore, upload speeds are more relevant to work-related activities such as
uploading documents or two-way audio, video, and text-based communication systems.
Meanwhile, the frequency of speed tests was important in identifying
the temporal profile which would give us most insight into
experienced internet service and resilience, and provide
an indication of the volume of experience over particular units of time.

<!-- data transformation -->
The first step in the workflow after dropping some outliers following
@riddlesden2014broadband was to transform the individual, geolocated and time-stamped
tests to more meaningful aggregates both in terms of space and time.
The frequency of testing indicates that whilst there is an overall trend of increased
testing from March to April and then a slight reduction from April to May,
this trend masks substantial variation by not only the day of the week,
but also time of day, as can be seen in Figure \ref{test2020}.
Thus, a daily aggregation of upload speeds would mask the variation in experienced
service over the course of each weekday.
Furthermore, the importance of this variation is highlighted by a comparison
with the same period in $2019$, as in Figure \ref{test2019},
when the volume of testing and thus of experience of internet service quality
was much more tied to the evening leisure peak and presumably to download speeds.
Since the increase in testing during the working day in $2020$ is an indication that
users have a greater perception of the variable quality of internet service,
particularly during a new morning peak of testing for service reliability,
we decided to include a measure of hourly variation in our temporal profiles.

However, although this is a large data set -- $241,088$ individual tests performed
during weekday hours across the study period --
there are not enough observations for each Local Authority District (LAD)
and for each working hour of each working day -- $631$ speed tests per LAD on average
-- to profile speeds at that level of detail.
Therefore, we aggregate all the speed-checks during the $13$ weeks of March to May
inclusive for weekdays in $2020$ by each hour of the day and day of the week.
As our research aims to identify the geography of internet service resilience for
work purposes, bank holidays and the hours between midnight and $6:00$ were excluded,
as well as weekend days.
The composite week time-series thus comprise $18$ hours multiplied by $5$ weekdays or
$90$ time points per series.
We also aggregate these data spatially because we could not follow individuals or
households and connect data points.
The time-series were calculated for each of the $382$ LADs in the UK, standardised,
and then a *k*-means partitioning around medoids clustering algorithm
was applied using DTW.
We initially run the algorithm for $k \in \mathbb{N} \bigcap [5,15]$
and used cluster validity indices (CVIs) to pick the optimal solution of $k = 13$.
Following @sardatime the majority vote for the following CVIs was used: Silhouette
(max), Score function (max), Calinski-Harabasz (max), Davies-Bouldin (min), Modified
Davies-Bouldin (DB*, min), Dunn (max), COP (min).

![Speed tests over time, 2020 \label{test2020}](figures/time.var.plot2020.png){height=40% width=75%}

![Speed tests over time, 2019 \label{test2019}](figures/time.var.plot2019.png){height=40% width=75%}

```{r echo=FALSE, eval=FALSE, message=FALSE}

# Code to calculate the above figures about N.

#upload data file
path.data <- paste(path, "./data/temp/TSbb19_20sp.csv", sep = "")
TSbb19_20sp <- read.csv(path.data)
TSbb20sp <- TSbb19_20sp %>%
  filter(dateOnly> "2019-12-31")

# N. of observations for study period  
print(nrow(TSbb20sp))

# observations for study period per local authority
print(nrow(TSbb20sp)/length(unique(TSbb20sp$LAD19NM)))
```

<!-- metrics -->
In Section [4.1](#sec:4.1), we review the temporal profile of upload speed
by hour of the day and day of the composite week, as well as
the experienced speed characteristics of each cluster.
Since the quality and reliability of internet services vary in time and space due
to both supply and demand-side influences, we use a number of different measures
to describe experienced upload speeds per cluster.
These include: a) mean, experienced connection speed,
b) standard deviation or the amount of fluctuation from the mean, and
c) the variation in speeds at particular times of day when working from home is
more likely to take place.
We take account of all three measurements in our descriptive statistics of
upload speeds in order to determine how resilient broadband speeds are
as experienced in different parts of the UK during a time of extreme demand.

<!-- link to auxillary regressions method -->
The cause of these different experiences of broadband resilience may be different in
different areas, as they may reflect either similarities in patterns of demand
or similar quality of infrastructure.
Our approach is also limited by potential endogeneity, as for example, better
quality connections with high mean speeds may enable more working from home, but
greater demand may cause slower speeds, less reliability and greater variability
of speed at different times of day or week.
Therefore, we avoid attributing any cause to our analysis of the experienced level
of quality and reliability of upload speeds.
Instead, we run an auxiliary regression in order to understand how the spatial and
temporal patterns of internet service relate to the economic geography of the UK.
More specifically, we estimate the following multinomial model:

\begin{align}
Pr(Y_{i}=j) = \frac{exp^{X_{i}\beta_{j}}}{\sum_{i=1}^j exp^{X_{i}\beta_{j}}}
\begin{cases}
    i = 1, 2, ... , N \\  
    j = 1, 2, ... , J
\end{cases}\label{eq1}
\end{align}

Based on the outcomes of the time-series clustering, we identify $J$ distinct
and crisp clusters.
We then regress this cluster membership against a vector $X_{i}$ of
socio-economic and geographic variables, which are discussed in detail in
the relevant Section [4.2]](#sec:4.2), in order to explore how the different patterns
might support or undermine efforts to work from home and maintain safe productivity
and whether they reinforce existing spatial and social inequalities.
This analysis enables us to provide a more nuanced understanding of how
telecommuting and technology intersect at a time of extreme demand, and what
lessons this time has for a future where telecommuting is likely to remain a
common means of accessing work and broadband services, as well as infrastructure,
must be fit for purpose.

# Results {#sec:4}

## Upload Clusters / cluster description {#sec:4.1}

The temporal profiles of the local authority clusters have been summarised in
Figures \ref{UpClusterL} and \ref{UpClusterS} and Table \ref{up.cluster.descr}.
The graphs show a composite profile of mean upload speeds per hour
per day for each cluster, with the largest, in terms of the LAD membership and
population, five clusters in Figure \ref{UpClusterL},
and the next six in Figure \ref{UpClusterS}.
These figures and table provide a comprehensive overview of the quality and
reliability of experienced broadband in different parts of the UK,
the temporal clusters offering a novel approach to understanding spatial disparity.

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{UpClusterL}Temporal profilies for upload speed large clusters"}
img1_path <- "figures/upClusterL.png"
include_graphics(img1_path)
```

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{UpClusterS}Temporal profilies for upload speed small clusters"}
img1_path <- "figures/upClusterS.png"
include_graphics(img1_path)
```

``````{r echo=FALSE, message=FALSE, results = "asis"}
path.clustersLADup <- paste(path, "./data/temp/LAD.cluster.up.csv", sep = "")

clustersLADup <- read_csv(path.clustersLADup)

# clustersLADup <- clustersLADup %>%
#   dplyr::rename('N. of LADs' = up.LADs,
#          'mean upload speed' = upMean,
#          'SD upload speed' = upSD,
#          'mean AM upload speed' = upAMmean)

clustersLADup <- clustersLADup %>%
  mutate(across(4:7, round, 0))

stargazer(clustersLADup[1:7],
          summary = F,
          font.size = "footnotesize",
          type = "latex",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          covariate.labels=c("Cluster",
                             "N. of LADs",
                             "LAD population",
                             "mean speed",
                             "SD speed",
                             "mean AM speed",
                             "mean PM speed"),
                             #"AM_slower",
                             #"SDtoMean"),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
          notes = "Note: All speed measures are upload speeds",
          title = "Upload speed cluster characteristics\\label{up.cluster.descr}")
```
The second largest cluster, comprising $126$ local authorities and over $20$ million
people, is cluster $13$.
Cluster $13$ has the slowest aggregate mean upload speed of any of the clusters,
and the second highest ratio of the standard deviation to the mean.
This suggests that those living in local authorities in this cluster experienced
some of the lowest quality broadband services in terms of upload speeds in the UK.
However, the variation in upload speeds in cluster $13$,
which can be an indication of its reliability, does not seem to disproportionately
affect the morning peak from $9:00$-$10:59$, as upload speeds are, on average,
only $2.6$% slower than in the evening peak period between $19:00$ and $20:59$,
when entertainment purposes are likely to be using the most bandwidth.
In comparison, the five LADs that are home to $343$ thousand people in cluster $1$
not only experience the second slowest mean upload speeds
and the highest ratio of standard deviation to the mean,
but are also much more affected during the morning peak.

Meanwhile, those living in the largest cluster -- $6$, with $155$ LADs home to $29.5$
million people, experience aggregate mean upload speeds of about $1$Mb/s faster
than those in cluster $13$, but still lower than the other three large clusters
and most of the smaller clusters, suggesting a middling quality of service.
However, the time profile for cluster $6$ in Figure \ref{UpClusterL} shows
that upload speeds are highest at $6:00$ on a Monday morning
and peak between $23:00$ and midnight on Wednesday and Thursday,
but tend to be lower during the working day.
Indeed, experienced mean upload speeds in the morning peak are $4.4$% lower
than in the evening peak -- a greater, more noticeable change than any of the other
large clusters experience, but smaller than any of the clusters with temporal
profiles shown in Figure \ref{UpClusterS}.
Clusters $8$ and $12$ also have mean upload speeds under $10$Mb/s, but higher than
clusters $1$ and $13$, and their standard deviation is not dissimilar.
However, this masks great variation in when lower speeds are experienced,
with the mean upload speeds much lower between $9:00-10:59$ than
between $19:00-20:59$ in cluster $8$, but slightly faster in the morning in
cluster $12$.

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{map.up.clusters}Upload speed clusters for LADs"}
img1_path <- "figures/map.up.clusters.png"
include_graphics(img1_path)
```

Indeed cluster $12$ is the only cluster to experience higher speeds in the evening
peak, compared to the morning peak, suggesting that widespread telecommuting
has generally changed the temporal profile of internet activity throughout the UK.
Yet even if all but one cluster is showing slower speeds in the morning than
the evening, the reliability of internet services in different clusters
during the working day still varies considerably.
Interpreting this variation from the large spikes and dips shown on
Figures \ref{UpClusterL} and \ref{UpClusterS} is difficult,
but the statistics in Table \ref{up.cluster.descr} show that clusters $9$ and $11$
have the most reliable internet services.
The ratio of standard deviation to mean in both these clusters is below 55%,
and the ratio of upload speeds in the peak periods indicates that speeds are
only about 2% slower in the morning than the evening.
Mean speeds are also higher than in any other cluster, excluding cluster $2$,
where measures of reliability suggests poorer performance.

Thus, broadband services in clusters $9$ and $11$, home to over twelve million people
performed the best during the study period, in terms of both quality and reliability.
In Figure \ref{UpClusterL}, cluster $11$ shows more noticeable peaks and troughs,
but the lowest points are not at the peak times described in Table \ref{up.cluster.descr}.
Rather, the slowest upload speeds on average occur between $6:00-7:00$ on
Monday morning, $14:00-15:00$ on Friday, and $16:00-17:00$ on Thursday.
These slowest times are still mostly faster than the average hourly upload speeds
in cluster $13$.
Finally, ignoring the smallest clusters in terms of population,
that is clusters $4$ and $5$,
clusters $3$, $7$ and $10$ also have relatively high mean upload speeds.
Clusters $3$ and $10$ pair high mean speeds with low standard deviations
relative to the mean speeds, suggesting reliability and resilience,
as well as quality broadband services.
Cluster $7$ has a higher ratio of standard deviation to mean, but there is less
difference in average speeds between the morning and evening peaks than in clusters
$3$ and $10$.

In summary, the local authorities in clusters $9$ and $11$ experienced resilient
broadband internet that could support high levels of telecommuting.
Those in clusters $2$, $3$, $7$, and $10$ also experience higher than average
mean speeds and rank middle to high on measures of service reliability.
These LADs are *not* on the wrong side of the first level digital divide,
but how likely are they to be able to take advantage of their resilient ICT
infrastructure and services?
Meanwhile, cluster $6$ is not only the largest in terms of number of LADs
and population, it has the closest mean upload speed to the pre-clustered average
for the whole sample.
As well as average quality internet services, those in cluster $6$ also experience
average reliability for work purposes, ranking fifth behind the four other clusters
with populations over one million, but ahead of the smaller clusters.
Clusters $8$ and $12$ are also close to average mean upload speeds, but show very
different patterns in terms of reliability, whilst clusters $1$ and $13$ appear to
suffer most from a lack of quality internet services, with slow speeds and
high standard deviations.
With those in cluster $1$ in particular more likely to experience that poor
reliability during the morning peak, is this first level digital divide occurring
in areas where few are occupationally able to telecommute anyway, and what are the
implications for economic resilience?

## Post-clustering regression analysis {#sec:4.2}

```{r echo=FALSE, message=FALSE, results = "asis"}
path.LAs_Clusters <- paste0(path, "/data/temp/LAs_Clusters.RData")
load(path.LAs_Clusters)

stargazer(aux.up,
          type = "latex",
          nobs = T,
          digits = 3,
          font.size = "tiny",
          float.env = "sidewaystable",
          dep.var.caption = "",
          title = "Auxiliary multinomial regression of upload speed clusters on socio-economic and geographic LAD variables\\label{aux}",
          header = F,
          covariate.labels=c(#"pop. 16-64, 2018",
                             "pop, 2018",
                             #"pop. densuty, 2018",
                             "job density, 2018",
                             "distance to nearest met. area",
                             "distance to London",
                             "South of the UK",
                             "managerial jobs, 2020",
                             "tech jobs, 2020",
                             "skilled trade jobs, 2020",
                             "professional jobs, 2020",
                             "administrative jobs, 2020",
                             "leisure jobs, 2020",
                             "machine operation jobs, 2020",
                             "earnings, 2019",
                             "n. business est. per hab., 2019",
                             "NVQ4+",
                             "AM tests per hab., 2020",
                             "Virgin Media \\%, 2020"),
          add.lines = list(c("McFadden's R squared",
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3)
                             , round(r2.up[1],3), round(r2.up[1],3)),
                           (c("N",
                            n.obs, n.obs, n.obs,
                            n.obs, n.obs, n.obs,
                            n.obs, n.obs, n.obs
                            , n.obs, n.obs))))
```

Using auxiliary regressions, we test whether the clusters that have higher mean
speeds and more reliable services consist of LADs that are more urban, affluent,
and / or more likely
to benefit from a choice of high quality internet services.
We also estimate which of our clusters are more likely to have a higher proportion of
occupations which enable telecommuting because of the nature of the work.
The results of these auxiliary regressions are presented in Table \ref{aux}.
The dependent variable is the LAD cluster membership as described in the
[methods and data](#sec:3) section
and equation \ref{eq1} and each column represent a different cluster.
The reference case is cluster $4$, which includes only the local authority of
Hambleton in North Yorkshire, a rural area of just over ninety thousand people.
Mean, experienced upload speeds in cluster $4$ are close to both the average speeds
for the $13$ clusters ($9.9$Mb/s) and the pre-clustered average for the whole sample
($9.3$Mb/s) as shown in Table \ref{up.cluster.descr}.
However, the standard deviation for cluster $4$ and the difference between average
speeds in the morning compared to the evening peak periods are indications of
worse reliability than many of the other clusters.
Hence, the results in Table \ref{aux} should be seen as relative
rather than absolute probabilities.

First, the number of speed tests run per cluster inhabitant between $9:00$-$10:59$
is an indication of satisfaction or at least a lack of concern over broadband speeds
and quality of service.
People in LADs assigned to clusters $6$ and $13$ were more likely to run speed tests
(at a per capita basis)
at this time of day, whilst people in cluster $11$ were the least likely to do so.
This high level of satisfaction matches the high quality and reliability of service
we would expect from the analysis in Section [4.1]](#sec:4.1).
Those in clusters $2$, $3$, $9$ and $10$ were also less likely to ran speed tests per
capita than those in the other six multi-authority clusters.
Cluster $7$ was the only cluster with higher mean speeds and middle to high measures
of reliability which was more likely to run more speed tests per capita.
This might be a reflection of the choice of ISP and connection options in those
LADs, as, along with cluster $12$, those in cluster $7$ were least likely to have
Virgin Media connections.
Fast Virgin Media cable connections have historically only been available
to $45$% of premises in the UK [@ofcom2016], where the more lucrative and
competitive market originally attracted the cable TV provider.
Those in clusters $2$, $3$, $9$ and $11$ benefit from a higher proportion of
Virgin connections, which is an indication that people in these clusters are
more likely to live in urban areas, with more choice of broadband services,
or in other words, are more likely to be on the right side of the first level,
infrastructure-based digital divide.

We employ distance from London and for the nearest metropolitan area (including
London) as two more variables depicting peripherality, urban structure and,
potentially, first level digital divides.
The broadband speed tests run in the authorities in cluster $3$ are more likely to
be taking place close to London than those run in any of the other clusters,
a result that makes sense considering two of the four authorities in cluster $3$
are the London commuter towns of Harlow and Luton.
However, even though London was also included in the variable calculating distance
from the centre of one of either the ten largest metropolitan areas in England,
or Glasgow or Cardiff, tests run in cluster $3$ are likely to be furthest away.
Still, whilst significant, the effect is tiny and the other members of cluster $3$
are Corby and Eastbourne, two towns which, whilst not part of metropolitan areas,
are also not in remote parts of the country, and are home to an estimated population
of $175$ thousand people between them. **move maps to here?**

Meanwhile, LADs in cluster $7$ are most likely to be near the centre of a large
metropolitan area, even though the four local authorities of cluster $7$ include
no central urban boroughs and only one LAD that is part of a metropolitan area of
governance - Tameside in Greater Manchester.
This may be why those in cluster $7$ are not served by Virgin Media.
It also is a demonstration of the complexity of experienced broadband upload speeds
as captured by time-series clustering, and their likely interaction with first level
digital divides.
Cluster $1$, which our analysis suggests lacks broadband resilience, contains five,
mainly rural authorities.
However, they are scattered around the country --
Ceredigion in West Wales, Darlington in the Northeast, Eden between the Lake District
and the North Penines, Rossendale in the Northwest, and Rutland in the East Midlands
-- and therefore, the results in Table \ref{aux} indicate that these authorities
are not as far from the centre of a metropolitan area as authorities in cluster $3$,
but furthest from London.

Internet resilience, as measured here, seems to be more nuanced than our --
arguably crude -- dummy variable depicting the North-South economic divide either,
which assigns $1$ to LADs locates in Greater London, Southeast, Southwest and East
of England regions, and $0$ to the rest.
The authorities most likely to be in the South are those in cluster $13$,
which was identified as having the slowest mean upload speeds of any of the clusters,
and thus a low level of service.
However, cluster $13$ does include some rural, remote areas of the country,
such as Northwest Scotland, Cornwall and Powys in Wales as shown on Figure
\ref{map.up.clusters}.
It also includes the major metropolitan centres of Bristol, Liverpool,
and Newcastle, as well as nine (of $32$) London Boroughs, 
including the City of London and plenty of home county and suburban areas.
Although by the standard deviation measure cluster $13$ has unreliable internet
services, speed variation was less during the morning peak, suggesting that 
the estimates for reliability are inconsistent. 
Considering that this is one of the largest clusters, and thus the averages 
incorporate more noise than some of the smaller clusters, 
it may be that the LADs in this cluster do not all suffer equally from 
a first level digital divide.

Yet we need to consider the results for other variables in Table \ref{aux} to better
determine whether clusters $1$ and $13$, which appear to suffer most from a lack 
of quality internet services, are also more likely to have a low skilled workforce, 
less able to benefit from telecommuting. 
Cluster $1$ has the lowest propotion of educated people of any of the clusters,
and the lowest earnings, despite recording the highest proportion of managerial, 
professional jobs, and the second highest proportion of tech jobs.
Rural areas such as those in cluster $1$ are often home to many retired people,
which might explain these results or perhaps, as these figures are relative,
we could note that cluster $1$ also has a greater proportion of skilled trades
than other clusters.
Either way, it seems likely that the first level digital divide reinforces 
other inequities in cluster $1$.
In comparison, those cluster $13$ are more likely to earn more and have a better
education.
Cluster $13$ also has the most businesses per inhabitant, but is somewhere in the 
middle in terms of job density.
If this is an indication of a high number of SMEs, it might give some explanation
for the variable internet quality, considering that small businesses have not been
seen as the most valuable customers for higher speed broadband services **insert ref**.

Cluster $6$ is the largest cluster, and thus, like cluster $13$, there is more 
noise within the averages we use to measure internet quality and reliability,
and the result indicates an average level of service, even if better than LADs
in cluster $13$.
Cluster $6$ also falls towards the middle of the clusters on many of the socio-
economic variables in Table \ref{aux}. 
It ranks third or fourth out of the eleven clusters in the table in terms of
having a higher proportion of managerial, professional, and tech jobs, as well as 
higher earnings, but is fourth from bottom in terms of education.
The LADs in cluster $6$ are also diverse, with few truly remote areas, 
but many urban areas in both the North, including Birmingham, Leeds and Sheffield, 
and South, including twelve London Boroughs, and many suburban areas and 
smaller cities like Oxford and Cambridge.
The capital cities of the other UK nations, Belfast, Cardiff and Edinburgh, are
also in this cluster, suggesting perhaps that the lower level of reliability 
discussed in Section [4.1](#sec:4.1) may be as a result of increased demand,
e.g. for telecommuting.

In comparison, clusters $9$ and $11$, which benefit from the most resilient
internet connections, are characterised by lower share of occupations that are  
may benefit from the ability to telecommute (e.g. managerial and technical). 
Cluster $9$ is more highly educated and higher earning than cluster $11$, 
has fewer businesses per inhabitant and less job density.
Indeed, cluster $11$ has a negative coefficient for the NVQ4 variable, whilst
for cluster $9$ the coefficient is positive, and cluster $11$ has the second
highest job density and cluster $9$ the second lowest. 
The former might be an indication of the higher returns that resilient
broadband infrastructure generates for cluster $9$ versus cluster $11$. 
Indeed, we did detect slightly more slowdown in the AM peak in cluster $9$ than in
cluster $11$ (see Table \ref{up.cluster.descr}).
Meanwhile, a propensity for a high job density might indicate more labour intensive 
economic structures in cluster $11$ than in the LADs in cluster $9$.

The LADs in clusters $9$ and $11$ also have opposite signs for their likelihood 
of being in the South of England.
Those in cluster $9$ less likely to be in the South and those in cluster $11$ 
more likely, although the coefficient for cluster $9$ might be skewed by 
the presence of the Scottish cities of Glasgow and Aberdeen, 
as Scotland has a different economic profile than England.
Still, both clusters consist of larger LADs in terms of population, 
including the urban centre of Manchester and three of the nine other districts of
Greater Manchester, Nottingham, both Portsmouth and Southampton,
eight London Boroughs, and four of the seven constituent authorities of the
West Midlands Combined Authority.
Among the $65$ LADs are also a number of other tightly bounded urban areas, 
such as Blackpool, Ipswich, Norwich, Slough, and Stevenage, and urban areas 
at the centre of less confined districts like Burnley, 
the digital hub of Milton Keynes, and Northampton.
Thus, our findings confirm the importance of geography and urban locations 
for being on the right side of the first level digital divide, and suggest that 
internet resilience cuts across other digital and economic divides.

A similarly divided picture can be built for the next tier of clusters: $2$, $3$,
$7$, $10$. LADs in these clusters experience higher than average mean speeds and
rank high to middle on measures of service reliability. 
LADs in these clusters are on the right side of the first level digital divide. 
However, only LADs in cluster $2$ and $10$ are more likely to have highly skilled
work force, which might more easily take advantage of working from home 
opportunities.
Cluster $2$ is comprised of just two LADs, with the lowest job density of any
cluster -- these are East Lothian near Edinburgh, and North East Lincolnshire, 
home to Grimsby.
Cluster $10$ is comprised of four LADs, including two peripheral suburban areas 
of Birmingham, the East London Borough of Newham, and Dundee.
Suburbs are considered the most likely urban form in which telecommuters live, 
and Dundee has a reputation for tech startups **insert ref**.
Yet although those in clusters $2$ and $10$ may be making the most of their 
resilient internet, those in cluster $7$ have the highest earnings.
The four LADs in cluster $7$ include the suburbs South of Belfast, a suburban 
district of Greater Manchester, the villages between York and Leeds, and a
couple of small West Midlands towns.
LADs in cluster $3$ are associated more with lower skills and have the highest
proportion of individuals in machine operation jobs. 
They include the Southeast commuter towns of Harlow and Luton and the slightly more
distant towns of Eastbourne and Corby.
Arguably, LADs in clusters $3$ benefit the least from their resilient internet 
infrastructure in an era when working from home became a vehicle for economic 
resilience.

Finally, clusters $8$ and $12$, are characterised by LADs enjoying close to average 
mean upload speeds, but different patterns of internet reliability. 
LADs in cluster $8$ are likely to host more individuals employed in tech and
administrative occupations than any other cluster, as well as many in managerial 
occupations, whilst the opposite applies to skilled trade and leisure occupations. 
These LADs are characterised by a very small positive likelihood of more educated 
residents, but are not significantly likely to earn more than other clusters.
Including suburbs of Leicester, Cardiff, Birmingham, the green belt area south of 
London straddling the M25 motorway, and rural West Devon, the LADs in cluster $8$
seem to likely to have the skills and occupations that would benefit from 
quality internet services, but suffer from poor internet resilience and reliability.

In comparison, LADs in cluster $12$ are less likely to be able to benefit from
quality internet services if they had them, with fewer individuals achieving 
NVQ4 or better and lower levels of occupations that would benefit from homeworking. 
Yet LADs in this cluster benefit from the second highest level of earnings, and
have the largest stock of businesses per inhabitant and the highest job density. 
This density of businesses could be why cluster $12$ is the only cluster with higher
speeds in the morning peak during the study period.
If people are at home, business premises might well be abandoned.
Cluster $12$ is home to $1.5$million people spread across twelve LADs, including
the City of Westminster and the London Borough of Hammersmith.
However, it also includes a LAD in Northern Ireland, two in Wales, two suburban areas
North of London, Preston in the Northwest, East Staffordshire in the Midlands, and
Fenland in the East of England.
This spatial diversity demonstrates that the temporal clustering of internet 
resilience does not create spatial clusters, and digital divides do not necessarily
overlap with economic ones.

# Conclusions {#sec:5}

Upload speeds are not used as benchmarks in the UK, nor have been seen as integral to
universal service, considering there has never before been such extreme demand for
telecommuting and operations such as video calls.

Clusters $3$, $9$, $10$ and $11$ seem to benefit most from high quality and
resilient broadband services.
The dips in mean upload speeds in clusters $6$, $3$ and $10$ during the morning
peak are suggestive of more use during the working day, and potentially more
telecommuting.

<!-- WFH broader consequences-->
The long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they span through various aspects of economy and society: from changes to
transportation planning due to altered commuting patters, to changes in land
use and urban planning to accommodate people who work from home
[@BUDNITZ2020102713]**also 2020 Swedish article from JTG**; and from productivity 
and innovation changes, to changes
in agglomeration externalities and the attraction of large cities [@econobs]
just to name a few. This paper is positioned to support endeavours in
understanding the effects of increased telecommuting by
exposing the spatial and social dimensions of telecommuting including the
resilience of broadband speeds in terms of both quality and reliability of
service, and whether this reinforces or redresses prior digital divisions.
**took this bit you wrote to put in the discussion at the end?**

# Acknowledgement(s) {-}

An unnumbered section, e.g.\ `\section*{Acknowledgements}`, may be used for thanks, etc.\ if required and included _in the non-anonymous version_ before any Notes or References.

# Appendix {#sec:appendix}

```{r echo=FALSE, message=FALSE, results = "asis"}

# Cluster table
clusters <- paste(path, "/data/temp/clusters_nodiff.csv", sep = "")
clusters <- read_csv(clusters)
clusters <- clusters %>%
  dplyr::select(LAD, cluster.up) %>%
  rename('cluster membership' = cluster.up) 
  #group_by('cluster membership') %>%
  #dplyr::arrange()
  #sort('cluster membership', decreasing = F)

stargazer(clusters,
          summary = F,
          font.size = "footnotesize",
          type = "latex",
          header = F,
          title = "LAD cluster membership for upload speed timeseries\\label{cluster.membership}")  
```

```{r echo=FALSE, message=FALSE, results = "asis"}

# Descriptive statistics table
path.LAs_Clusters <- paste0(path, "/data/temp/LAs_Clusters.RData")
load(path.LAs_Clusters)

for.descr <- data %>%
  dplyr::select(pop, job.dens2018, distMet,
                distLondon, south, 
                managers, tech, skilled, prof,
                admin, caring, plant, 
                earnings, nvq4,
                virgin,
                total.busi, LAD.tests.am20, pop) %>%
  mutate('total.busi/pop' = total.busi/pop,
         'LAD.tests.am20/pop' = LAD.tests.am20/pop) %>%
  dplyr::select(-total.busi, -pop, -LAD.tests.am20)

stargazer(for.descr,
          summary = T,
          font.size = "footnotesize",
          type = "latex",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          covariate.labels=c("pop, 2018",
                             "job density, 2018",
                             "distance to nearest met. area",
                             "distance to London",
                             "South of the UK",
                             "managerial jobs, 2020",
                             "tech jobs, 2020",
                             "skilled trade jobs, 2020",
                             "professional jobs, 2020",
                             "administrative jobs, 2020",
                             "leisure jobs, 2020",
                             "machine operation jobs, 2020",
                             "earnings, 2019",
                             "NVQ4+",
                             "Virgin Media \\%, 2020",
                             "n. business est. per hab., 2019",
                             "AM tests per hab., 2020"),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
          #notes = "",
          title = "Descriptive statistics for the auxiliary regression explanatory variables\\label{descr.aux}")

```