### 2020-12-30 paper/v1/broadband.speed.covid.Rmd

---
title: "Working from home and digital divides: resilience in the time of a pandemic"
author:
- name: A. Uthor*
  num: 1,2
- name: O. Tro
  num: 2
- name: O. Vriga
  num: 3
address:
- num: 1
  org: Department of Incredible Research, University A, City A, Country A
- num: 2
  org: Department of Applied Things, University B, City B, Country B
- num: 3
  org: Very Important Stuff Committee, Institute C, City C, Country C
corrauth: "Corresponding author name, This is sample corresponding address."
email: correspondingauthor@protonmail.com
abstract: "This paper offers a new perspective on telecommuting from the viewpoint of the complex web of digital divides. Using the UK as a case study, this paper studies how the quality and reliability of internet services, as reflected in *experienced* internet upload speeds during the spring 2020 lockdown, may reinforce or redress the spatial and social dimensions of digital divisions. Fast, reliable internet connections are necessary for the population to be able to work from home. Although not every place hosts individuals in occupations which allow for telecommuting nor with the necessary skills to effectively use the internet to telecommute, good internet connectivity is also essential to local economic resilience in a period like the current pandemic. Employing data on individual broadband speed tests and state-of-the-art time-series clustering methods, we create clusters of UK local authorities with similar temporal signatures of experienced upload speeds. We then associate these clusters of local authorities with their socioeconomic and geographic characteristics to explore how they overlap with or diverge from the existing economic and digital geography of the UK. Our analysis enables us to better understand how the spatial and social distribution of both occupations and online accessibility intersect to enable or hinder the practice of telecommuting at a time of extreme demand."
keywords:   covid; internet; working from home; broadband speed; time-series clusters;
classoption:
  - Royal
  - times
bibliography: bibliography
bibliographystyle: sageh
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{dcolumn}
  \def\tightlist{}
  \newcommand{\beginappendix}{\setcounter{table}{0} \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
output:
  rticles::sage_article:
    keep_tex: yes
---

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)
library(stargazer)
library(png)

# This is the project path
path <- find_rstudio_root_file()
```

# Introduction {#sec:1}

<!-- [our google doc](https://docs.google.com/document/d/1PWjkmgzWGYKR9wFogKYw7l-8mZLoORt593x-Tu-f2-M/edit#heading=h.i5om1o8wpcd9) -->

During the pandemic, working from home using digital technologies, whether partially
or exclusively, was transformed from a niche means of accessing work, albeit one that
had been on a slow, upward trend, to a widespread way of life in many countries.
The ability to work from home or telecommute meant millions retained their jobs and,
to a varying extent, maintained productivity during periods of strict
lockdowns around the world.
However, this ability has not been evenly distributed socially or spatially,
creating new intersections of economic and digital divisions.
On one side are those who can work from home, supported by digital technologies,
and have thus been able to enjoy both economic resilience and greater personal safety.
On the other side, previously employed individuals have been forced to accept
furlough or redundancy packages unless they are part of the cadre of essential
workers, who are potentially at high risk of infection.
Whilst the basis for this pandemic-generated divide has been viewed as mainly
occupational, here we consider whether it is also technological and, consequently, geographical.

<!-- our approach -->
Using the UK as a case study, this paper aims to understand how the quality and
reliability of internet service, as reflected in *experienced* internet speeds
during the spring 2020 lockdown,
may reinforce or redress the spatial and social dimensions of digital divisions.
We employ volunteered geographic data on individual broadband speed tests
and state-of-the-art time-series clustering methods to create clusters of UK local
authorities with similar temporal signatures of experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore how they overlap with or diverge from
the existing economic and digital geography of the UK.
Our analysis enables us to better understand how the spatial and social distribution
of both occupations and online accessibility intersect to enable or hinder the
practice of telecommuting at a time of extreme demand.
We also consider what lessons can be learned from this time for a future where
telecommuting is likely to remain a more common means of accessing work,
at least in comparison to the pre-Covid era,
and broadband services and infrastructure must be fit for purpose.

<!-- capability to WFH -->
The capability to work from home has previously been studied from the
perspective of whether work tasks in a given occupation both can be and are
allowed to be performed using digital technologies independently of
location or co-location with colleagues, including supervisors
[@allen2015effective; @singh2013modeling].
However, successful telecommuting also requires that the quality and
reliability of digital services, particularly home internet connection speeds,
enable the completion of work tasks with a minimum of delay or interruption.
Prior to the pandemic, the performance of broadband services with respect to
telecommuters was never tested at scale, as working from home and connecting to
colleagues and workplace resources via the internet was the purview of a small minority of workers.
Instead, leisure use in the evening, when video streaming services are at their peak,
has been used to benchmark broadband performance and service delivery by different
Internet Service Providers (ISPs) [@ofcom2017].

Yet the shift towards telecommuting during various stages of lockdown around the world
has been drastic and there are speculations that post-Covid, the tendency to work from
home will be much higher, raising questions around whether internet
services can accommodate the increased demand.
For example, $47$% of people in employment in the UK worked solely from home in April
$2020$, whilst the same figure only reached $5$% the year before [@ons2020;
@ons2020lm2019].
A back of the envelope calculation suggests that up to 40% of the labour force
could work from home on an ongoing basis [@batty2020editorial].
Similar figures have been reported for other countries [@felstead2020homeworking].
Approximately $37$% of the European workforce worked from home in April $2020$ with
countries like Finland reaching $60$% [@eurofound2020].
In the US, almost half of the working population worked from home during the same
period because of the pandemic [@brynjolfsson2020covid],
and a recent estimate indicated that $37$% of all jobs in the US can be permanently
performed entirely from home [@NBERw26948].
None of these changes could have happened in the absence of reliable information
and communication technology (ICT) infrastructure -- both in terms of software
and hardware.
But while software innovations are easily diffused across space and
society^[See for example the huge success of videoconferencing apps such as Zoom
[@marks2020zoom].], the same does not apply for ICT hardware infrastructure such as
internet broadband connectivity.

<!-- ICT infrastructure and digital divides-->
The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest in
different geographies in the UK [@riddlesden2014broadband; @philip2017digital].
Second level digital divides consider the presence or lack of the necessary skills
to effectively utilise digital technologies and the internet
[@blank2014dimensions; @van2011internet].
The third level focuses on the heterogenous returns of internet usage among
different socioeconomic groups and, consequently, how digital technologies can
assist in bridging or further enhancing existing socioeconomic divides.
[@stern2009levels; @van2014digital; @van2015third].
The capability to telecommute is related to all three levels of digital divides,
but more importantly leads to differentiated outcomes regarding the economic
resilience of people and places to overcome a systemic shock such as the current
pandemic.

As the quality of internet infrastructure and services, as well as the concentration
of different occupations are spatially dependent and clustered in space,
our approach offers a framework for understanding the impact of and interactions
between digital divisions geographically and socioeconomically.
By asking how resilient broadband speeds, and particularly upload speeds, are
experienced in different parts of the UK during a time of extreme demand,
we interrogate which places benefit from the greater economic resilience
digital technologies can offer.
The structure of this paper is as follows.
First we review the literature on telecommuting and digital divides to better
understand their structural and spatial development pre-pandemic,
and thus their importance to the economic resilience of different places.
We then describe our data and methodology.
Our results section first offers a classification of how internet services vary
across clusters of UK local authorities and then assesses whether these clusters
replicate or repudiate other socio-economic and geographic patterns of economic
resilience.
We conclude with a discussion of the insights we have gained from our new
perspective on digital divisions.

# Literature review {#sec:2}

## From telecommuting to #WFH {#sec:2.1}

<!-- Who works from home-->
In this analysis, the terms ‘telecommuting’ and ‘working from home^[See also the popular social media hashtag #WFM]’ are used
interchangeably, as most remote labour during the Covid-19 crisis was carried out
in the homes of individual employees rather than any other location [@eurofound2020].
However, previous research has explored how telecommuting
can occur in other places, including satellite offices or on public transport
[@felstead2012rapid; @siha2006telecommuting].
Previous research has also used a variety of definitions to measure the level of
telecommuting within different workforces, distinguishing between those
directly employed, indirectly employed, self-employed, full-time or part-time, and
those who use digital technologies to work remotely full-days or part-days
[@allen2015effective; @bailey2002review; @haddad2009examination].
No matter the definition, the option and capability to telecommute or work from home
has never been equally distributed spatially or socio-economically any more than
different industries and employment opportunities have.
Studies from the United States, the Netherlands, and the UK found
that telecommuters are most likely to hold professional, managerial, and technical
occupations where the workforce is better educated and wealthier, and that there is
suppressed demand among women and part-time workers
[@headicar2016move; @peters2004employees; @singh2013modeling].

Opportunities for working from home during the current pandemic have likewise not
been equally spread across the workforce.
@NBERw26948 indicated that in the US, managers, educators, those
working in computer-related occupations, finance, and law can easily work from home,
and that occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in more spatially fixed occupations,
from farming, construction and manufacturing to hospitality and care services.
In the US, these occupations tend to be lower-income, non-white,
without a university degree, live in rental accommodation and lack health insurance
[@NBERw27085].
Similar trends can be observed for other countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, whilst only $34$% of workers with secondary education and
$14$% of those primary education did so [@eurofound2020].

## Digital divides and economic resilience {#sec:2.2}
<!-- WFH and digital divides-->
Our understanding of telecommuting as a product of enabled occupations can be
described as a manifestation of the third level digital divide, as those who are able
to use digital technologies to work from home benefit from a high rate of return on
their use of the internet in terms of autonomy, flexibility, and time saved from
commuting [@peters2004employees; @siha2006telecommuting; @singh2013modeling].
These returns have been even greater during the Covid-19 crisis,
when those with the capability to telecommute also have the ability to maintain
their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the first
level digital divide, which is associated with access to and quality of internet
connectivity.
@SALEMINK2017360 provides a systematic review of the pre-pandemic,
first level digital divide in infrastructure quality between urban and rural areas
in various advanced economies.
Rural areas, predictably, fare worse.
Yet whether this variation in infrastructure quality affects the spatial footprint of
telecommuting has not previously been investigated, in part because telecommuting has
not previously been a cause of concentrated demand and pressure on internet services.

There are indications that those who purchase high speed connections consume
more data of all sorts and use their connections for a greater variety of purposes
[@hauge2011consumer].
There is also a correlation between access to internet services and a
reduction in household transport spend [@bris2017ict].
Independently of whether the additional internet use and reduced travel is
because of increased telecommuting, these studies suggest that better internet
services enable households to make savings and efficiencies,
an example of the first level digital divide reinforcing the third level.

<!-- Covid-19 and working from home, cities, urban structure -->
Multi-layered digital divides may also intersect with material divides
and the economic geography of the UK.
The regional economic resilience literature underlines the differentiated capacity 
of cities and regions to escape or recover from economic crises 
[@martin2012regional; @kitsos2018economic].
As different places have different industrial and occupational profiles, 
these affect their aggregated potential capacity for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south regions of England in terms of skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers from some of the highest levels of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
All three levels of digital divides are associated with or shaped by the
geography of the UK. Yet this is the first time that the intersection of digital 
and material divides is relevant to understanding the economic resilience of places 
and large swathes of the population, as digital technologies became an essential
tool of productivity during the Covid-19 pandemic.

The extreme demand during the pandemic thus provides a new opportunity to understand
how internet infrastructure quality, and reliability affects telecommuting, 
particularly in light of the high volumes of bandwidth-intensive 
video conferencing required in order to avoid the face-to-face contact that could 
increase the spread of infection.
We seek to answer how internet service resilience contributes to or reduces economic
resilience when the latter is dependent upon the capability to work from home.
We also aim to improve our understanding of the impact of first level digital
division on telecommuting, and whether this results in more fundamental third
level digital divisions than has previously been perceived.

# Methods and data {#sec:3}

## Time-Series clustering {#sec:3.1}
<!-- clustering in general -->
Our chosen methodological framework is cluster analysis, which can
be defined within machine learning approaches as an unsupervised
learning task, partitioning unlabelled observations into homogeneous groups
known as clusters [@montero2014tsclust].
The key idea is that observations within clusters tend to be more similar than
observations between clusters.
Clustering is particularly useful for exploratory studies as it identifies
structures within the data [@aghabozorgi2015time].
Cluster analysis is a widely used in geography
[@gordon1977classification; @everitt1974cluster], for example to solve 
*regionalisation* problems [@niesterowicz2016].
Clustering methods are also the basis of *geodemographics*, a research
domain which aims to create small area indicators or typologies of neighbourhoods
based on diverse variables [@SINGLETON2009289; @harris2005geodemographics].
These studies usually employ cross-sectional data, and most clustering problems in 
geography deal with observations that are fixed in time.
However, for this paper we are interested in internet speeds, which vary over time.
Therefore, we create clusters of local authorities in the UK with similar
temporal signatures of experienced internet speeds.

<!-- time-series clustering -->
To do so, we employ time-series clustering methods, which have been developed
to deal with clustering problems linked to, for instance, stock or other financial
data, economic, governmental or medical data as well as machine monitoring
[@aggarwal2013time; @aggarwal2001surprising; @hyndman2015large; @WARRENLIAO20051857].
The main challenge, which does not apply to cross-sectional clustering
problems, is data dimensionality, with a multiplicity of data points for every
individual object included in the data set, and how their value changes dynamically
as a function of time [@aghabozorgi2015time].
This high dimensionality leads to
(i) computational and algorithmic challenges regarding handling these data and
building algorithms to perform clustering over long time-series, and
(ii) open questions regarding the choice of similarity measures in order to cluster
similar times-series objects together considering the whole length of the time-series
and overcoming issues around noise, outliers and shifts [@lin2004iterative;
@aghabozorgi2015time].

For this paper we utilise a category of time-series clustering methods
known as shape-based approaches.
These methods match two separate time-series objects based on the similarity of
their shapes through the calculation of distances between the shapes,
and are better equipped to capture similarities between short length time-series
[@aghabozorgi2015time], such as our data.
We thus identify clusters of UK local authorities with similar temporal signatures --
i.e. shapes -- of experienced internet speeds.
The clusters are identified using the common partitioning algorithm, 
where no clusters overlap, known as *k*-means.
This iterative algorithm is popular because of the simplicity of the implementation 
and the interpretability of the results.
It begins with defining the desired number of clusters: *k*.
Then each observation is randomly assigned to a cluster from the $[1,k]$ space.
This initial cluster assignment is followed by iterations in order to minimise
the distance between the centroids of the clusters and the observations
assigned to these clusters [@james2013introduction].

<!-- difference between clustering cross-sectional and time-series data -->
There are a number of differences between the application of *k*-means for
cross-sectional and times-series data.
Instead of creating clusters around centroids, a common approach is to create
clusters around *medoids*, which are representative time-series objects with a
minimal distance to all other cluster objects [@sardatime].
Also, instead of calculating the Euclidean distance between centroids and data
points, more complex distance measures need to be employed to capture
the similarity between a time-series object and a medoid.
A common distance measure for shape-based time-series clustering is
Dynamic Time Warping (DTW), an algorithm comparing two time-series objects to find
the optimum warping path between them.
DTW is widely used in order to overcome limitations linked to the use of Euclidean
distance [@sardatime; @berndt1994using; @ratanamahatana2004everything].
The `R` package `dtwclust` has been used for the time-series clustering [@dtwclust].

## Experienced Broadband Speeds {#sec:3.2}
<!-- introduction of our broadband speeds data -->
To assess the internet quality and reliability across local authorities in the UK,
we utilise unique data comprising individual internet speed tests from
Speedchecker Ltd^[[https://www.broadbandspeedchecker.co.uk/](https://www.broadbandspeedchecker.co.uk/)].
This is a private company that allows internet users to check their own broadband
upload and download speeds, and stores every speed-check with timestamp and
geolocation information.
These data have been used before to assess digital divides [@riddlesden2014broadband]
and the impact of local loop unbundling regulatory processes
[@nardotto2015unbundling], and we followed the former's approach to remove outliers.
By using this volunteered geographic data, we are able to assess the internet speed
*experienced* by users, which may differ from the maximum speeds *advertised* by
ISPs.
We are particularly interested in upload speeds and the frequency of speed tests
over the period from March to May $2020$, as government statements indicate this is
when UK workers were first told to work from home if at all possible [@GovUK2020].
Average upload speeds are slower than average download speeds, at $9.3$Mb/s
mean upload speed for the whole sample,
compared to $29.6$Mb/s for download speeds, but they are also less associated
with internet-based, high-demand, leisure activities such as video streaming.
Therefore, upload speeds are more relevant to work-related activities such as
uploading documents or two-way audio, video, and text-based communication systems.

<!-- data transformation -->
The frequency of speed tests was important in identifying the temporal profile which
would give us most insight into experienced  internet service and resilience over
units of time.
Whilst there is an overall trend of increased testing
from March to April and then a slight reduction from April to May,
this trend masks substantial variation by not only the day of the week,
but also time of day, as can be seen in Figure \ref{test2020}.
Thus, a daily aggregation of upload speeds would mask the variation in experienced
service over the course of each weekday.
Furthermore, the importance of this variation is highlighted by a comparison
with the same period in $2019$, as in Figure \ref{test2019}, when the volume of
testing and thus of experienced internet service quality peaked in the evening,
presumably in response to demand for leisure activities and download speeds.
In contrast, the majority of the increase in testing in $2020$ is during the working
day, creating a new morning peak in Figure \ref{test2020}.
Therefore, we include a measure of hourly variation in our temporal profiles to
reflect the change in users' perception of the workday reliability of internet
services.

However, there were insufficient observations -- only $631$ speed
tests per LAD on average -- for each for each working hour of each working day in
each Local Authority District (LAD) to profile speeds at that level of detail.
Spatial aggregation was also necessary because we could not follow individuals or
households and connect data points.
Therefore, we aggregate the $241,088$ individual, geolocated and time-stamped
speed-checks during the $13$ weeks of March to May inclusive for weekdays in $2020$
by each hour of the day and day of the week.
As our research aims to identify the geography of internet service resilience for
work purposes, bank holidays and the hours between midnight and $6:00$ were excluded,
as well as weekend days.
The composite week time-series thus comprise $18$ hours multiplied by $5$ weekdays or
$90$ time points per series.
The time-series were calculated for each of the $382$ LADs in the UK, standardised,
and then a *k*-means partitioning around medoids clustering algorithm
was applied using DTW.
We initially run the algorithm for $k \in \mathbb{N} \bigcap [5,15]$
and used cluster validity indices (CVIs) to pick the optimal solution of $k = 13$.
Following @sardatime the majority vote for the following CVIs was used: Silhouette
(max), Score function (max), Calinski-Harabasz (max), Davies-Bouldin (min), Modified
Davies-Bouldin (DB*, min), Dunn (max), COP (min).

![Speed tests over time, 2020 \label{test2020}](figures/time.var.plot2020.png){height=40% width=75%}

![Speed tests over time, 2019 \label{test2019}](figures/time.var.plot2019.png){height=40% width=75%}

```{r echo=FALSE, eval=FALSE, message=FALSE}

# Code to calculate the above figures about N.

#upload data file
path.data <- paste(path, "./data/temp/TSbb19_20sp.csv", sep = "")
TSbb19_20sp <- read.csv(path.data)
TSbb20sp <- TSbb19_20sp %>%
  filter(dateOnly> "2019-12-31")

# N. of observations for study period  
print(nrow(TSbb20sp))

# observations for study period per local authority
print(nrow(TSbb20sp)/length(unique(TSbb20sp$LAD19NM)))
```

<!-- metrics -->
In Section [4.1](#sec:4.1), we review the temporal profile of upload speed
by hour of the day and day of the composite week for each cluster.
Since the quality and reliability of internet services vary in time and space due
to both supply and demand-side influences, we also use a number of different measures
to describe experienced upload speeds per cluster.
These include: i) mean, experienced connection speed,
ii) standard deviation or the amount of fluctuation from the mean, and
iii) the variation in speeds during the new morning peak of testing when working
from home is more likely to take place.
We take account of all three measurements
in order to determine how resilient broadband speeds are
as experienced in different parts of the UK during a time of extreme demand.

<!-- link to auxillary regressions method -->
The cause of these different experiences of broadband resilience may vary between
and within clusters, as they may reflect either patterns of demand or quality 
of infrastructure.
Our approach is also limited by potential endogeneity, as for example, better
quality connections with high mean speeds may enable more working from home, but
greater demand can cause slower speeds, less reliability, or greater variability
of speed at different times of day or week.
Therefore, we avoid attributing any cause to our analysis of the experienced level
of quality and reliability of upload speeds.
Instead, we run an auxiliary regression to understand how the spatial and
temporal patterns of internet service relate to the economic geography of the UK.
More specifically, we estimate the following multinomial logit model:

\begin{align}
Pr(Y_{i}=j) = \frac{exp^{X_{i}\beta_{j}}}{\sum_{i=1}^j exp^{X_{i}\beta_{j}}}
\begin{cases}
    i = 1, 2, ... , N \\  
    j = 1, 2, ... , J
\end{cases}\label{eq1}
\end{align}

Based on the outcomes of the time-series clustering, we identify $J$ distinct
and crisp clusters.
We then regress this cluster membership against a vector $X_{i}$ of
socio-economic and geographic variables, which are discussed in detail in
the relevant Section [4.2](#sec:4.2).
<!-- limitations-->
Because we cannot identify individuals or households and consequently aggregated our 
data at the LAD level, our results offer correlations between the socioeconomic
characteristics of certain geographic locations and internet service quality,
not a record of who was telecommuting.
Such individual data could be found though surveys, but these offer less
detailed information about the experience of internet resilience due to enforced
demand, which is the main contribution of this paper.
Our auxiliary regression, therefore, provides an indication of how
internet connectivity can reinforce or redress existing spatial and social
inequalities in different places.
However, it opens a path to future research by highlighting the importance of 
understanding of how telecommuting capabilities and digital infrastructure divisions
intersect.

# Results {#sec:4}

## Upload Clusters / cluster description {#sec:4.1}

The temporal profiles of the local authority clusters have been summarised in
Figures \ref{UpClusterL} and \ref{UpClusterS} and Table \ref{up.cluster.descr}.
The graphs show a composite profile of mean upload speeds per hour
per day for each cluster, with the largest five clusters, 
in terms of the LAD membership and population, in Figure \ref{UpClusterL},
and the next six in Figure \ref{UpClusterS}.
These figures and table provide a comprehensive overview of the quality and
reliability of experienced broadband in different parts of the UK.

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{UpClusterL}Temporal profilies for upload speed large clusters"}
img1_path <- "figures/upClusterL.png"
include_graphics(img1_path)
```

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{UpClusterS}Temporal profilies for upload speed small clusters"}
img1_path <- "figures/upClusterS.png"
include_graphics(img1_path)
```

``````{r echo=FALSE, message=FALSE, results = "asis"}
path.clustersLADup <- paste(path, "./data/temp/LAD.cluster.up.csv", sep = "")

clustersLADup <- read_csv(path.clustersLADup)

# clustersLADup <- clustersLADup %>%
#   dplyr::rename('N. of LADs' = up.LADs,
#          'mean upload speed' = upMean,
#          'SD upload speed' = upSD,
#          'mean AM upload speed' = upAMmean)

clustersLADup <- clustersLADup %>%
  mutate(across(4:7, round, 0))

stargazer(clustersLADup[1:7],
          summary = F,
          font.size = "footnotesize",
          type = "latex",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          covariate.labels=c("Cluster",
                             "N. of LADs",
                             "LAD population",
                             "mean speed",
                             "SD speed",
                             "mean AM speed",
                             "mean PM speed"),
                             #"AM_slower",
                             #"SDtoMean"),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
          notes = "Note: All speed measures are upload speeds",
          title = "Upload speed cluster characteristics\\label{up.cluster.descr}")
```
The second largest cluster, comprising $126$ local authorities and over $20$ million
people, is cluster $13$, which has the slowest aggregate mean upload speed,
and the second highest ratio of the standard deviation to the mean.
This suggests that those living in local authorities in this cluster experienced
some of the lowest quality broadband services in terms of upload speeds in the UK.
However, Figure \ref{UpClusterL} shows that the high standard deviation,
which is one indication of unreliable internet, did not disproportionately affect
the morning peak from $9:00$-$10:59$, when upload speeds were, on average,
only $2.6$% slower than in the evening peak period between $19:00$ and $20:59$,
when entertainment purposes are likely to be using the most bandwidth.
In comparison, the five LADs in cluster $1$, home to $343$ thousand people, not only
experience the second slowest mean upload speeds and the highest ratio of standard
deviation to the mean, but are also much more affected during the morning peak.

Meanwhile, those living in the largest cluster -- $6$, with $155$ LADs and $29.5$
million people -- experienced aggregate mean upload speeds of about $1$Mb/s faster
than those in cluster $13$, but still lower than the other three large clusters
and most of the smaller clusters, suggesting a middling quality of service.
The temporal profile for cluster $6$ in Figure \ref{UpClusterL} shows that upload
speeds are highest at $6:00$ on a Monday morning and between $23:00$ and midnight 
on Wednesday and Thursday, but tend to be lower during the working day.
Furthermore, experienced mean upload speeds in the morning peak are $4.4$% lower
than in the evening peak -- a greater, more noticeable change than any of the other
large clusters experience, suggesting poorer reliability during the working day.
This difference is less, however than any of the clusters
included in Figure \ref{UpClusterS}.

Clusters $8$ and $12$ also have mean upload speeds under $10$Mb/s, but higher than
clusters $1$, $13$, and $6$. However, mean upload speeds are much lower between
$9:00-10:59$ than between $19:00-20:59$ in cluster $8$, but slightly faster in the
morning than in the evening in cluster $12$.
Indeed, cluster $12$ is the only cluster to experience higher speeds in the morning
peak compared to the evening, and thus the only cluster where the temporal profile
of internet use is closer to what might have been expected pre-pandemic.
And yet even among the other clusters, the reliability of internet services
during the working day still varies considerably.
Interpreting this variation from the large spikes and dips shown on
Figures \ref{UpClusterL} and \ref{UpClusterS} is difficult,
but the statistics in Table \ref{up.cluster.descr} show that clusters $9$ and $11$
have the most reliable internet services.
The ratio of standard deviation to mean in both these clusters is lowest,
and speeds are only about 2% slower in the morning than the evening.
Mean speeds are also higher than in any other cluster, excluding cluster $2$,
where measures of reliability suggest poorer performance.

Thus, broadband services in clusters $9$ and $11$, 
home to over twelve million people, performed the best during the study period, 
in terms of both quality and reliability.
In Figure \ref{UpClusterL}, cluster $11$ shows more noticeable peaks and troughs,
with the lowest points occurring, on average, between $6:00-7:00$ on Monday morning, 
$14:00-15:00$ on Friday, and $16:00-17:00$ on Thursday, not the workday peak. 
Furthermore, these slow periods offer better speeds than the average hourly
profile of cluster $13$. Finally, ignoring the smallest clusters in terms of
population, that is clusters $4$ and $5$, clusters $3$, $7$ and $10$ also have
relatively high mean upload speeds. Clusters $3$ and $10$ pair high mean speeds with
low standard deviations relative to the mean speeds, suggesting reliability and
resilience, as well as quality broadband services. Cluster $7$ has a higher ratio of
standard deviation to mean, but there is less difference in average speeds between
the morning and evening peaks than in clusters $3$ and $10$.

In summary, LADs in clusters $9$ and $11$ experienced resilient
broadband internet that could support high levels of telecommuting.
Those in clusters $2$, $3$, $7$, and $10$ also experience higher than average
mean speeds and rank middle to high on measures of service reliability.
These LADs are *not* on the wrong side of the first level digital divide,
but how likely are they to be able to take advantage of their resilient ICT
infrastructure and services?
Meanwhile, cluster $6$ is not only the largest in terms of number of LADs
and population, it has the closest mean upload speed to the pre-clustered average
for the whole sample.
As well as average quality internet services, those in cluster $6$ also experience
average reliability for work purposes, ranking fifth behind the four other clusters
with populations over one million, but ahead of the smaller clusters.
Clusters $8$ and $12$ are also close to average mean upload speeds, but show very
different patterns in terms of reliability, whilst clusters $1$ and $13$ appear to
suffer most from a lack of quality internet services, with slow speeds and
high standard deviations.
With those in cluster $1$ more likely to experience that poor reliability during the
morning peak, is this first level digital divide occurring in areas where few are
occupationally able to telecommute anyway, and what are the implications for
economic resilience?

## Post-clustering regression analysis {#sec:4.2}

```{r echo=FALSE, message=FALSE, results = "asis"}
path.LAs_Clusters <- paste0(path, "/data/temp/LAs_Clusters.RData")
load(path.LAs_Clusters)

stargazer(aux.up,
          type = "latex",
          nobs = T,
          digits = 3,
          font.size = "tiny",
          float.env = "sidewaystable",
          dep.var.caption = "",
          title = "Auxiliary multinomial regression of upload speed clusters on socio-economic and geographic LAD variables\\label{aux}",
          header = F,
          covariate.labels=c(#"pop. 16-64, 2018",
                             "pop, 2018",
                             #"pop. densuty, 2018",
                             "job density, 2018",
                             "distance to nearest met. area",
                             "distance to London",
                             "South of the UK",
                             "managerial jobs, 2020",
                             "tech jobs, 2020",
                             "skilled trade jobs, 2020",
                             "professional jobs, 2020",
                             "administrative jobs, 2020",
                             "leisure jobs, 2020",
                             "machine operation jobs, 2020",
                             "earnings, 2019",
                             "n. business est. per hab., 2019",
                             "NVQ4+",
                             "AM tests per hab., 2020",
                             "Virgin Media \\%, 2020"),
          add.lines = list(c("McFadden's R squared",
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3)
                             , round(r2.up[1],3), round(r2.up[1],3)),
                           (c("N",
                            n.obs, n.obs, n.obs,
                            n.obs, n.obs, n.obs,
                            n.obs, n.obs, n.obs
                            , n.obs, n.obs))))
```

Using an auxiliary multinomial logit regression, we test whether the clusters that
have higher mean speeds and more reliable services consist of LADs that are more
urban, affluent, and / or more likely to benefit from a choice of high quality
internet services.
We also estimate which of our clusters are more likely to have a higher proportion of
occupations where the nature of the work enables telecommuting.
The results of the auxiliary regression are presented in Table \ref{aux}.
The dependent variable is the LAD cluster membership as described in the
[methods and data](#sec:3) section
and equation \ref{eq1}. Each column represents a different cluster.
The reference case is cluster $4$, which includes only the local authority of
Hambleton in North Yorkshire, a rural area of just over ninety thousand people.
Mean, experienced upload speeds in cluster $4$ (see Table \ref{up.cluster.descr})
are between the average speeds for the $13$ clusters ($9.9$Mb/s) and the
pre-clustered average for the whole sample ($9.3$Mb/s).
However, the standard deviation for cluster $4$ and the difference between average
speeds in the morning compared to the evening peak periods are indications of
worse reliability than many of the other clusters. Hence, the results in Table
\ref{aux} should be seen as relative rather than absolute probabilities.

First, we control for the number of speed tests run per cluster inhabitant
between $9:00$-$10:59$ as well as the share of fast Virgin Media internet
connections. Regarding the former, we expect people in LADs with more unreliable
connections to test their internet speeds more often, and the results validate our
priors. Meanwhile, fast Virgin Media cable connections have historically
only been available to $45$% of premises in the UK [@ofcom2016], where the more
lucrative and competitive market originally attracted the cable TV provider.
Those in clusters $2$, $3$, $9$ and $11$ benefit from a higher proportion of
Virgin connections, which is an indication that people in these clusters are
more likely to live in urban areas, with more choice of broadband services.
In other words, they are more likely to be on the right side of the first level,
infrastructure-based digital divide, as we expected from the analysis in Section
[4.1](#sec:4.1).

We employ distance from London and from the nearest metropolitan area
(including London) as two variables depicting peripherality, urban structure and,
potentially, first level digital divides.
The broadband speed tests run in the authorities in cluster $3$ are more likely to
be taking place close to London than those run in any of the other clusters, and
two of the four authorities in cluster $3$ are the London commuter towns of Harlow
and Luton.
However, even though London was also included in the variable calculating distance
from the centre of one of either the ten largest metropolitan areas in England,
or Glasgow or Cardiff, tests run in cluster $3$ are likely to be furthest away.
Thus, it is important to consider the membership of each cluster as well as the
regression results. Corby and Eastbourne, the other two authorities in cluster $3$
are large, accessible towns, although not part of metropolitan areas.

```{r echo=FALSE, out.width = "95%",fig.cap="\\label{map.up.clusters}Upload speed clusters for LADs"}
img1_path <- "figures/map.up.clusters.png"
include_graphics(img1_path)
```

Meanwhile, LADs in cluster $7$ are most likely to be near the centre of a large
metropolitan area, even though the four local authorities of cluster $7$ include
no central urban boroughs and only one LAD that is part of a metropolitan area of
governance - Tameside in Greater Manchester. This may explain why those in cluster
$7$ are the second least likely to be served by Virgin Media.
It is also a demonstration of the complexity of both experienced broadband 
upload speeds as captured by time-series clustering, and the geography of 
first level digital divides as a product of quality as well as availability.
In comparison, cluster $1$, which our analysis suggests lacks broadband resilience, 
contains five, mainly rural authorities, but they are closer to a metropolitan area 
than authorities in cluster $3$, although furthest from London, perhaps because 
they are scattered around the country.

Internet resilience is also more nuanced than our -- arguably crude
-- dummy variable depicting the North-South economic divide, which assigns $1$ to
LADs located in Greater London, Southeast, Southwest and East of England regions,
and $0$ to the rest.
The authorities most likely to be in the South are those in cluster $13$,
which was identified as having the slowest mean upload speeds of any of the clusters,
and thus a low level of service.
However, cluster $13$ includes some rural, remote areas of the country, such as
Northwest Scotland, Cornwall and Powys in Wales as shown on Figure
\ref{map.up.clusters}.
It also includes major metropolitan centres in the North -- Liverpool, Newcastle --
and South -- Bristol, and nine (of $32$) London Boroughs.
There are also plenty of Southern home county and suburban areas.
The standard deviation measure for cluster $13$ is high, but speed variation is
low during the morning peak, suggesting that the estimates for reliability are
inconsistent. Considering that this is one of the largest clusters, and thus the
averages incorporate more noise than some of the smaller clusters,
it may be that the LADs in this cluster do not all suffer equally from
a first level digital divide.

Yet we need to consider the results for other variables in Table \ref{aux} to better
determine whether clusters $1$ and $13$, which appear to suffer most from poor
quality internet services, are also more likely to have a low skilled workforce,
less able to benefit from telecommuting.
Cluster $1$ has the lowest proportion of educated people,
and the lowest earnings, despite recording the highest proportion of managerial and
professional jobs, and the second highest proportion of tech jobs.
Rural areas such as those in cluster $1$ are home to many retired people,
which might explain these results or perhaps, as these figures are relative,
we could note that cluster $1$ also has a greater proportion of skilled trades
than other clusters. It seems likely that the first level digital divide reinforces
other inequities in cluster $1$, whilst those in cluster $13$ are more likely
to earn more and have a better education despite poor internet services.
Cluster $13$ also has the most businesses per inhabitant, but is somewhere in the
middle in terms of job density. If this is an indication of a high number of SMEs,
it could explain the variable internet quality, considering that small businesses
have not been seen as the most valuable customers for higher speed broadband services
if they are not located near residential customers [@ofcom2016].

Cluster $6$ is the largest cluster, and thus, like cluster $13$, there is more
noise within the averages we use to measure internet quality and reliability.
Our results indicate average mean speeds, and Table \ref{aux} shows that
cluster $6$ also falls towards the middle of the clusters on many of the
socioeconomic variables. It ranks third or fourth out of the eleven in terms of
the likelihood of having a higher proportion of managerial, professional, and
tech jobs, as well as higher earnings, but is fourth from bottom for educational
attainment. The LADs in cluster $6$ are also diverse, with few truly remote areas,
but urban areas throughout England, including Birmingham, Leeds, Sheffield,
twelve London Boroughs, and many suburban areas and smaller cities like Oxford
and Cambridge. The capital cities of the other UK nations, Belfast, Cardiff and
Edinburgh, are also in this cluster, suggesting perhaps that the lower level of
reliability discussed in Section [4.1](#sec:4.1) may be due to increased demand,
e.g. for telecommuting, when it is not supported by the most resilient internet
connections.

In contrast, Clusters $9$ and $11$ enjoy resilient internet connections, 
but LADs in Cluster $9$ are more likely to host highly educated individuals 
with higher earnings than cluster $11$, which has a negative coefficient for 
the NVQ4 variable. Cluster $11$ has fewer businesses per inhabitant and the 
second highest job density and cluster $9$ the second lowest.
These coefficients might indicate that resilient broadband infrastructure
generates higher returns for those in cluster $9$, where slightly more slowdown in
the AM peak was detected (see Table \ref{up.cluster.descr}).
LADs in cluster $9$ are also less likely to be in the South and those in cluster $11$
more likely, although the coefficient for cluster $9$ might be skewed by
the presence of the Scottish cities of Glasgow and Aberdeen. Scotland has a
different economic profile than England. Still, both clusters consist of larger
LADs in terms of population, including the city centre among four of the ten
districts of Greater Manchester, eight London Boroughs, four of the seven
constituent authorities of the West Midlands Combined Authority, Nottingham, and
both Portsmouth and Southampton. Among the $65$ LADs are also a number of other
tightly bounded urban areas, such as Blackpool, Ipswich, Norwich, Slough, and
Stevenage, and urban areas at the centre of less confined districts like Burnley,
the digital hub of Milton Keynes, and Northampton.
These urban locations are on the right side of the first level digital divide,
and the regression results suggest that internet resilience is supporting a wide
range of small and large urban economies.

LADs in clusters $2$, $3$, $7$, $10$ are also on the right side of the first level
digital divide, experiencing higher than average mean speeds and ranking high to
middle on measures of service reliability.
However, only LADs in cluster $2$ and $10$ are more likely to have highly skilled
workers to take advantage of working from home opportunities.
Cluster $2$ is comprised of just two LADs, with the lowest job density of any
cluster. Cluster $10$ is comprised of four LADs, including two
peripheral suburban areas of Birmingham, the East London Borough of Newham,
and Dundee. Suburbs are considered the most likely urban form in which telecommuters
live [@e2018does], and Dundee has a reputation for tech startups [@technation2017].
LADs in cluster $7$, which include the suburbs South of Belfast, a suburban
district of Greater Manchester, the villages between York and Leeds, and a couple of
small West Midlands towns have the highest earnings, and thus may also be making the
most of their resilient internet. In contrast, LADs in cluster $3$ are associated
more with lower skills and have the highest proportion of individuals in machine
operation jobs. Arguably, LADs in clusters $3$ benefit the least from their
resilient internet infrastructure in an era when working from home became a vehicle
for economic resilience.

Finally, clusters $8$ and $12$ consist of LADs enjoying close to average
mean upload speeds, but opposing patterns of internet reliability during the morning
peak. LADs in cluster $8$ are likely to host more individuals employed in tech and
administrative occupations than any other cluster, as well as many in managerial
occupations, whilst the opposite applies to skilled trade and leisure occupations.
These LADs are characterised by a very small positive likelihood of more educated
residents, but are not significantly likely to earn more than other clusters.
Including suburbs near Leicester, Cardiff, Birmingham, the green belt area south of
London straddling the M25 motorway, and rural West Devon, the LADs in cluster $8$
seem to likely to have the skills and occupations that would benefit from
quality internet services, but suffer from poor internet resilience and reliability.

In comparison, LADs in cluster $12$ are less likely to be able to benefit from
quality internet services if they had them, with fewer individuals achieving
NVQ4 or better and lower levels of occupations that would benefit from homeworking.
Yet LADs in this cluster benefit from the second highest level of earnings, and
have the largest stock of businesses per inhabitant and the highest job density.
This density of businesses could be why cluster $12$ is the only cluster with higher
speeds in the morning peak during the study period. If people are at home, business
premises might well have been abandoned. Cluster $12$ is home to $1.5$ million people
spread across twelve LADs, including the City of Westminster, 
another London Borough, and two others outside London, but also a LAD in
Northern Ireland, two in Wales, Preston in the Northwest, and Fenland in the East of
England. This spatial diversity demonstrates that the temporal clustering of internet
resilience is not necessarily spatially dependent, and digital divides do not
necessarily overlap with economic ones.

# Conclusions {#sec:5}

This paper offers a new perspective on telecommuting from the viewpoint of the
complex web of digital divides. We employ novel data regarding
experienced upload speeds and time-series clustering methods,
a family of unsupervised machine learning techniques which are rarely utilised
in geographical research. Fast, reliable internet connections are
necessary for the population to be able to work from home.
Although not every place hosts individuals in occupations which allow for
telecommuting nor with the necessary skills to effectively use the internet to
telecommute, this paper raises the issue that without good internet connectivity,
these places would struggle more to achieve economic resilience in a period like
the current pandemic in any case.
Indeed, our analysis demonstrated that the temporal profiles of twelve of our
thirteen clusters had slower upload speeds in the morning than in the evening.
The opposite is likely to have been the norm prior to the pandemic, as the level of
demand and bandwidth management is the most common cause of temporal variation in
experienced speeds, and why evening download speeds, rather than daytime upload
speeds, have been used to benchmark the performance of internet services.
Thus, the new patterns can be taken as evidence of widespread telecommuting
and other daytime internet use which changed the temporal profile of
internet activity throughout the UK, not just in areas with more digital industry
or better skills.

Upload speeds have not previously been seen as integral to universal service,
considering there has never before been such extreme demand for telecommuting
and operations such as video calls.
This may be why the average upload speeds for the largest two clusters
-- $6$ and $13$ -- contain so many LADs that are centres of the knowledge economy,
from Oxford and Cambridge in the former, to areas like Reading in the latter, 
which has the highest concentration of digital businesses in the country 
[@technation2017].
Thus, whilst some areas suffer from an intersection of digital and economic divides,
such as those in cluster $1$, others found their digital infrastructure to be less
reliable when confronted with the sudden change in the timing and type of demand.
Still other LADs, such as Milton Keynes in cluster $9$ were able to benefit both
from reliable internet connections and populations which were familiar with working
from home and could capitalise on their digital infrastructure.
Yet the nuanced picture we gained through our analysis of the UK case study suggests
that quality internet connectivity would also enable other LADs to gain ground
during the pandemic despite lower skills or earnings currently, especially in a
future where telecommuting might be a more common means of accessing work
and broadband services must be fit for purpose.

<!-- WFH broader consequences-->
Indeed, the long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they span various aspects of economy and society: from changes to
transportation planning due to altered commuting patterns to changes in land
use and urban planning to accommodate people who work from home
[@BUDNITZ2020102713; @ELLDER2020102777], and from productivity
and innovation changes to changes in agglomeration externalities and the
attraction of large cities [@econobs].
Further research may be able to measure the economic resilience of the different
clusters of places discussed in this paper once this pandemic is firmly past.
However, our analysis demonstrates that the economic resilience made possible by
working from home cannot be understood without considering the underpinning digital
divides and cannot be achieved without planning for how the levels of digital,
social and economic divides might intersect.

<!-- # Acknowledgement(s) {-} -->

<!-- An unnumbered section, e.g.\ `\section*{Acknowledgements}`, may be used for thanks, etc.\ if required and included _in the non-anonymous version_ before any Notes or References. -->

<!-- \pagebreak -->

# Appendix 1 {#appendix1}

This is the LAD cluster membership for the upload speed timeseries.

```{r echo=FALSE, message=FALSE, results = "asis"}

# Cluster table
clusters <- paste(path, "/data/temp/clusters_nodiff.csv", sep = "")
clusters <- read_csv(clusters)
clusters <- clusters %>%
  dplyr::select(LAD, cluster.up) %>%
  #rename('cluster membership' = cluster.up) %>%
  #group_by('cluster membership') %>%
  group_split(cluster.up, .keep = F)
  #dplyr::arrange()
  #sort('cluster membership', decreasing = F)

cluster1 <- as.vector(unlist(clusters[1]))
cluster2 <- as.vector(unlist(clusters[2]))
cluster3 <- as.vector(unlist(clusters[3]))
cluster4 <- as.vector(unlist(clusters[4]))
cluster5 <- as.vector(unlist(clusters[5]))
cluster6 <- as.vector(unlist(clusters[6]))
cluster7 <- as.vector(unlist(clusters[7]))
cluster8 <- as.vector(unlist(clusters[8]))
cluster9 <- as.vector(unlist(clusters[9]))
cluster10 <- as.vector(unlist(clusters[10]))
cluster11 <- as.vector(unlist(clusters[11]))
cluster12 <- as.vector(unlist(clusters[12]))
cluster13 <- as.vector(unlist(clusters[13]))
```

**Cluster 1: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster1, type="cmd2"), collapse=", "))
```

**Cluster 2: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster2, type="cmd2"), collapse=", "))
```

**Cluster 3: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster3, type="cmd2"), collapse=", "))
```

**Cluster 4: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster4, type="cmd2"), collapse=", "))
```

**Cluster 5: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster5, type="cmd2"), collapse=", "))
```

**Cluster 6: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster6, type="cmd2"), collapse=", "))
```

**Cluster 7: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster7, type="cmd2"), collapse=", "))
```

**Cluster 8: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster8, type="cmd2"), collapse=", "))
```

**Cluster 9: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster9, type="cmd2"), collapse=", "))
```

**Cluster 10: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster10, type="cmd2"), collapse=", "))
```

**Cluster 11: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster11, type="cmd2"), collapse=", "))
```

**Cluster 12: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster12, type="cmd2"), collapse=", "))
```

**Cluster 13: **
```{r echo=FALSE, message=FALSE, results = "asis"}
cat(paste(shQuote(cluster13, type="cmd2"), collapse=", "))
```

```{r echo=FALSE, message=FALSE, results = "asis", eval=FALSE}
# LADs <- do.call(cbindX, clusters) # this is with library(gdata)
# LADs1 <- LADs[,1:7]
# LADs2 <- LADs[,8:13]
#
# names(LADs1) <- c("Cluster 1", "Cluster 2", "Cluster 3",
#                      "Cluster 4", "Cluster 5", "Cluster 6",
#                      "Cluster 7")
# names(LADs2) <- c("Cluster 8", "Cluster 9", "Cluster 10",
#                      "Cluster 11", "Cluster 12")
#
# opts <- options(knitr.kable.NA = "") # kableExtra.auto_format = FALSE
#
# # LADs1 %>%
# # kable(caption = "LAD cluster membership for upload speed timeseries (1 out of 2)\\label{cluster.membership1}") %>%
# #   kable_styling(font_size = 5,fixed_thead = T)
#
# LADs1 %>%
# kbl(caption = "LAD cluster membership for upload speed timeseries (1 out of 2)\\label{cluster.membership1}") %>%
#    kable_classic(font_size = 5,
#                 fixed_thead = T,
#                 full_width = F
#                 )
#
#
# # stargazer(LADs1,
# #           summary = F,
# #           font.size = "footnotesize",
# #           type = "latex",
# #           header = F,
# #           title = "LAD cluster membership for upload speed timeseries (1 out of 2)\\label{cluster.membership1}")  
#
# stargazer(LADs2,
#           summary = F,
#           font.size = "footnotesize",
#           type = "latex",
#           header = F,
#           title = "LAD cluster membership for upload speed timeseries (2 out of 2)\\label{cluster.membership2}")  
```

# Appendix 2 {#appendix2}

```{r echo=FALSE, message=FALSE, results = "asis"}

# Descriptive statistics table
path.LAs_Clusters <- paste0(path, "/data/temp/LAs_Clusters.RData")
load(path.LAs_Clusters)

for.descr <- data %>%
  dplyr::select(pop, job.dens2018, distMet,
                distLondon, south,
                managers, tech, skilled, prof,
                admin, caring, plant,
                earnings, nvq4,
                virgin,
                total.busi, LAD.tests.am20, pop) %>%
  mutate('total.busi/pop' = total.busi/pop,
         'LAD.tests.am20/pop' = LAD.tests.am20/pop) %>%
  dplyr::select(-total.busi, -LAD.tests.am20)

stargazer(for.descr,
          summary = T,
          font.size = "footnotesize",
          type = "latex",
          header = F,
          digit.separator = ",",
          digit.separate = c(3,3),
          covariate.labels=c("pop, 2018",
                             "job density, 2018",
                             "distance to nearest met. area",
                             "distance to London",
                             "south of the UK",
                             "managerial jobs, 2020",
                             "tech jobs, 2020",
                             "skilled trade jobs, 2020",
                             "professional jobs, 2020",
                             "administrative jobs, 2020",
                             "leisure jobs, 2020",
                             "machine operation jobs, 2020",
                             "earnings, 2019",
                             "NVQ4+",
                             "Virgin Media \\%, 2020",
                             "n. business est. per hab., 2019",
                             "AM tests per hab., 2020"),
          rownames = F,
          no.space=TRUE,
          column.sep.width = "0pt",
          #notes = "",
          title = "Descriptive statistics for the auxiliary regression explanatory variables\\label{descr.aux}")

```

\pagebreak