### 2021-03-04 src/ts_clusters_k5.Rmd

---
title: "Time series clusters, k = 5"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(tsbox)
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)
library(imputeTS)
library(tsbox)
library(dtwclust)
library(leaflet)
library(ggmap)
library(cowplot)

# This is the project path
path <- find_rstudio_root_file()
```

```{r include = FALSE}

# load the data created by Data_Spatial.Rmd

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")

TSbb19_20sp <- read_csv(path.data)

#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs
```

## Time series clusters

I create clusters based on upload, download and number of tests time series for tests working hours and days from 16/2/2020 onwards.

More work needs to be done to select the optimal *k* as well as the clustering methods.
The `dtwclust` package comes with quite some readings, saved in the `.\literature\`.

Also, we need to think about upload / download / differnces / period + **imputation**.

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# download 2020
ts.down2020 <- TS2020 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(mean.down = mean(speeddown))#, mean.down = mean(speeddown), n.tests = n())

# upload 2020
ts.up2020 <- TS2020 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(mean.up = mean(speedup))

# tests 2020
ts.tests2020 <- TS2020 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(n.tests = n())

# download 2019
ts.down2019 <- TS2019 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(mean.down = mean(speeddown))#, mean.down = mean(speeddown), n.tests = n())

# upload 2019
ts.up2019 <- TS2019 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(mean.up = mean(speedup))

# tests 2019
ts.tests2019 <- TS2019 %>%
  group_by(weekday, hour, LAD19NM) %>%
  summarise(n.tests = n())

```

### Upload speed

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# create a 'fake' week for a ts id
help.dates <- tibble(help.date.time = c(make_datetime(year = 2020, month = 3, day = 2),
                                        make_datetime(year = 2020, month = 3, day = 3),
                                        make_datetime(year = 2020, month = 3, day = 4),
                                        make_datetime(year = 2020, month = 3, day = 5),
                                        make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
ts.wide <- merge(ts.up2020, help.dates, by = "weekday")
ts.wide$hour <- paste(ts.wide$hour, ":00", sep="")
ts.wide$help.date.time <- with(ts.wide, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide <- ts.wide %>%
  dplyr::select(help.date.time, LAD19NM, mean.up) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# Define k instead of selecting it
k <- 5

pc_k <- tsclust(ts.wide, k = k, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))
plot(pc_k)

cluster_up <- data.frame(LAD=ts.wide.names,
                         cluster=pc_k@cluster)
```

### Download speed

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

ts.wide <- merge(ts.down2020, help.dates, by = "weekday")
ts.wide$hour <- paste(ts.wide$hour, ":00", sep="")
ts.wide$help.date.time <- with(ts.wide, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide <- ts.wide %>%
  dplyr::select(help.date.time, LAD19NM, mean.down) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# Define k instead of selecting it

pc_k <- tsclust(ts.wide, k = k, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))
plot(pc_k)

cluster_down <- data.frame(LAD=ts.wide.names,
                           cluster=pc_k@cluster)
```

### N. of speed tests

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# tests 
ts.wide <- merge(ts.tests2020, help.dates, by = "weekday")
ts.wide$hour <- paste(ts.wide$hour, ":00", sep="")
ts.wide$help.date.time <- with(ts.wide, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide <- ts.wide %>%
  dplyr::select(help.date.time, LAD19NM, n.tests) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)
#sapply(ts.wide, function(x) sum(is.na(x)))

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# Define k instead of selecting it

pc_k <- tsclust(ts.wide, k = k, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))
plot(pc_k)

cluster_tests <- data.frame(LAD=ts.wide.names,
                            cluster=pc_k@cluster)
```

### Upload speed diff 2019-2020

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# 2020 copied from above

# create a 'fake' week for a ts id
help.dates <- tibble(help.date.time = c(make_datetime(year = 2020, month = 3, day = 2),
                                        make_datetime(year = 2020, month = 3, day = 3),
                                        make_datetime(year = 2020, month = 3, day = 4),
                                        make_datetime(year = 2020, month = 3, day = 5),
                                        make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
ts.wide <- merge(ts.up2020, help.dates, by = "weekday")
ts.wide$hour <- paste(ts.wide$hour, ":00", sep="")
ts.wide$help.date.time <- with(ts.wide, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide <- ts.wide %>%
  dplyr::select(help.date.time, LAD19NM, mean.up) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)

# 2019

# create a 'fake' week for a ts id
help.dates <- tibble(help.date.time = c(make_datetime(year = 2019, month = 3, day = 4),
                                        make_datetime(year = 2019, month = 3, day = 5),
                                        make_datetime(year = 2019, month = 3, day = 6),
                                        make_datetime(year = 2019, month = 3, day = 7),
                                        make_datetime(year = 2019, month = 3, day = 8)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
ts.wide2019 <- merge(ts.up2019, help.dates, by = "weekday")
ts.wide2019$hour <- paste(ts.wide2019$hour, ":00", sep="")
ts.wide2019$help.date.time <- with(ts.wide2019, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide2019 <- ts.wide2019 %>%
  dplyr::select(help.date.time, LAD19NM, mean.up) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide2019 <- ts_wide(ts.wide2019)

# impute with means of each TS
# TODO check sensitivity
ts.wide2019 <- na_mean(ts.wide2019)
#sapply(ts.wide, function(x) sum(is.na(x)))

# diff 2020-2019
ts.wide <- mapply('-', ts.wide, ts.wide2019, SIMPLIFY = FALSE)
ts.wide <- as.data.frame(ts.wide)

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# Define k instead of selecting it

pc_k <- tsclust(ts.wide, k = k, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 20L)))
plot(pc_k)

cluster_up.diff <- data.frame(LAD=ts.wide.names,
                              cluster=pc_k@cluster)

```

### download speed diff 2019-2020

```{r eval=TRUE, echo=TRUE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# 2020 copied from above

# create a 'fake' week for a ts id
help.dates <- tibble(help.date.time = c(make_datetime(year = 2020, month = 3, day = 2),
                                        make_datetime(year = 2020, month = 3, day = 3),
                                        make_datetime(year = 2020, month = 3, day = 4),
                                        make_datetime(year = 2020, month = 3, day = 5),
                                        make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
ts.wide <- merge(ts.down2020, help.dates, by = "weekday")
ts.wide$hour <- paste(ts.wide$hour, ":00", sep="")
ts.wide$help.date.time <- with(ts.wide, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide <- ts.wide %>%
  dplyr::select(help.date.time, LAD19NM, mean.down) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide <- ts_wide(ts.wide)

# impute with means of each TS
# TODO check sensitivity
ts.wide <- na_mean(ts.wide)

# 2019

# create a 'fake' week for a ts id
help.dates <- tibble(help.date.time = c(make_datetime(year = 2019, month = 3, day = 4),
                                        make_datetime(year = 2019, month = 3, day = 5),
                                        make_datetime(year = 2019, month = 3, day = 6),
                                        make_datetime(year = 2019, month = 3, day = 7),
                                        make_datetime(year = 2019, month = 3, day = 8)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
ts.wide2019 <- merge(ts.down2019, help.dates, by = "weekday")
ts.wide2019$hour <- paste(ts.wide2019$hour, ":00", sep="")
ts.wide2019$help.date.time <- with(ts.wide2019, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

ts.wide2019 <- ts.wide2019 %>%
  dplyr::select(help.date.time, LAD19NM, mean.down) #%>%
#filter(date > "2020-2-15")

# turn to a wide format
ts.wide2019 <- ts_wide(ts.wide2019)

# impute with means of each TS
# TODO check sensitivity
ts.wide2019 <- na_mean(ts.wide2019)
#sapply(ts.wide, function(x) sum(is.na(x)))

# diff 2020-2019
ts.wide <- mapply('-', ts.wide, ts.wide2019, SIMPLIFY = FALSE)
ts.wide <- as.data.frame(ts.wide)

# extract names
ts.wide.names <- names(ts.wide[,-1])
ts.wide <- ts.wide[,-1]

# standardise
ts.wide <- zscore(ts.wide)

# convert to list for tsclust
ts.wide <- split(ts.wide, rep(1:ncol(ts.wide), each = nrow(ts.wide)))

# Define k instead of selecting it

pc_k <- tsclust(ts.wide, k = k, seed = 94L, 
                type="partitional", 
                distance = "dtw_basic", centroid = "pam", trace = T, 
                args = tsclust_args(dist = list(window.size = 15L)))
plot(pc_k)

cluster_down.diff <- data.frame(LAD=ts.wide.names,
                                cluster=pc_k@cluster)
```

## Mapping the clusters

```{r include=TRUE, echo=FALSE, results= 'markup', message=FALSE, warning = FALSE, fig.height=10, fig.width=10}

# all clusters data frame
clusters <- merge(cluster_up, cluster_down, by = "LAD")
names(clusters)[2] <- "cluster.up"
names(clusters)[3] <- "cluster.down"

clusters <- merge(clusters, cluster_up.diff, by = "LAD")
names(clusters)[4] <- "cluster.up.diff"

clusters <- merge(clusters, cluster_down.diff, by = "LAD")
names(clusters)[5] <- "cluster.down.diff"

clusters <- merge(clusters, cluster_tests, by = "LAD")
names(clusters)[6] <- "cluster.tests"
#sapply(clusters, function(x) sum(is.na(x)))

# clusters  %>% arrange(cluster.up) %>% kable()

# save clusters file
out.clusters <- paste(path, "/data/temp/clusters_all_5.csv", sep = "")
write_csv(clusters, out.clusters)

# clusters data frame WITHOUT the diff clusters
clusters.nodiff <- merge(cluster_up, cluster_down, by = "LAD")
names(clusters.nodiff)[2] <- "cluster.up"
names(clusters.nodiff)[3] <- "cluster.down"

clusters.nodiff <- merge(clusters.nodiff, cluster_tests, by = "LAD")
names(clusters.nodiff)[4] <- "cluster.tests"
#sapply(clusters, function(x) sum(is.na(x)))

clusters.nodiff  %>% arrange(cluster.up) %>% kable()

# save clusters file
out.clusters <- paste(path, "/data/temp/clusters_nodiff_5.csv", sep = "")
write_csv(clusters.nodiff, out.clusters)

# get LA 
# this is the heaviest version:
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
# path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
# la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC: 
la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
# UK BFC: la <- readOGR("https://opendata.arcgis.com/datasets/1d78d47c87df4212b79fe2323aae8e08_0.geojson")

# source: https://data.gov.uk/dataset/7c387c64-d25f-474a-b07e-b933578caea2/local-authority-districts-april-2019-boundaries-uk-bfe

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))
#la@data$LAD19NM <- as.character(la@data$LAD19NM)
la@data$lad19nm <- as.character(la@data$lad19nm)

# ggplot2 maps
la.f <- fortify(la, region = "lad19nm")
la.f <- merge(la.f, clusters.nodiff, by.x = "id", by.y = "LAD")
la.f <- la.f[order(la.f$order),] # if i don't order merge.nuts.f loses order and the map has gaps

ggplot(la.f, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = as.factor(cluster.tests))) +
  #theme_nothing(legend = TRUE) +
  labs(title = "Test clusters") +
  scale_fill_viridis_d()+
  guides(fill=guide_legend(title="Clusters")) +
  theme_void()

ggplot(la.f, aes(x = long, y = lat)) +
  geom_polygon(aes( group = group, fill = as.factor(cluster.up))) +
  #theme_nothing(legend = TRUE) +
  labs(title = "Upload clusters") +
  scale_fill_viridis_d()+
  guides(fill=guide_legend(title="Clusters")) +
  theme_void()

#ggsave("all_maps2.png") # this is the colour map for the preprint
# out.path.map <- paste0(path, "/paper/v1/figures/map.up.clusters5.png")
# ggsave(out.path.map)#,
#width = 210, #A4
#height = 297, #A4
#units = "mm")
```

```{r, eval=FALSE}
# leaflet maps

# !!!WRONG CATEGORY VISUALISATION!!!

#la@data <- merge(la@data, clusters, by.x = "LAD19NM", by.y = "LAD")
la@data <- merge(la@data, clusters.nodiff, by.x = "lad19nm", by.y = "LAD")

# from handbook
la@data$cluster.tests <- as.factor(la@data$cluster.tests)
factpal <- colorFactor(topo.colors(5), la@data$cluster.tests)

leaflet(la) %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
              color = ~factpal(la@data$cluster.tests)) %>%
  addMarkers(lng=la$long, lat=la$lat, popup=la@data$lad19nm) %>%
  addLegend(pal = factpal, 
            values = ~la@data$cluster.tests,
            opacity = 1,
            #group = "cluster.tests",
            title = "Test clusters")

leaflet(la) %>%
  addPolygons(
    fillColor = factpal(la$cluster.tests),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7) %>%
  addMarkers(lng=la$long, lat=la$lat, popup=la@data$lad19nm)

# circles
leaflet(la@data) %>%
  addTiles() %>%
  addCircles(lng = ~long, lat = ~lat, weight = 1, 
             radius = 500L, popup = ~lad19nm, color = ~factpal(cluster.tests))

# factors
la@data$cluster.up <- as.factor(la@data$cluster.up)
la@data$cluster.down <- as.factor(la@data$cluster.down)
#la@data$cluster.up.diff <- as.factor(la@data$cluster.up.diff)
#la@data$cluster.down.diff <- as.factor(la@data$cluster.down.diff)
la@data$cluster.tests <- as.factor(la@data$cluster.tests)

# palettes
factpal.up <- colorFactor("RdYlBu", la@data$cluster.up)
factpal.down <- colorFactor("RdYlBu", la@data$cluster.down)
#factpal.up.diff <- colorFactor("RdYlBu", la@data$cluster.up.diff)
#factpal.down.diff <- colorFactor("RdYlBu", la@data$cluster.down.diff)
factpal.tests <- colorFactor("RdYlBu", la@data$cluster.tests)

leaflet(la) %>%
  # add polygons as groups
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
              color = ~factpal.up(cluster.up), group = "cluster.up") %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
              color = ~factpal.down(cluster.down), group = "cluster.down") %>%
  #addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
  #  color = ~factpal.up.diff(cluster.up.diff), group = "cluster.up.diff") %>%
  #addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
  #  color = ~factpal.down.diff(cluster.down.diff), group = "cluster.down.diff") %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
              color = ~factpal.tests(cluster.tests), group = "cluster.tests") %>%
  addMarkers(lng=la$long, lat=la$lat, popup=la$lad19nm) %>%
  # hide groups, so they don't appear all on in the begining 
  hideGroup("cluster.up") %>%
  hideGroup("cluster.down") %>%
  #hideGroup("cluster.up.diff") %>%
  #hideGroup("cluster.down.diff") %>%
  hideGroup("cluster.tests") %>%
  # layer control
  addLayersControl(
    #baseGroups = c("Upload)", "Download"),
    overlayGroups = c("cluster.up","cluster.down", "cluster.down", "cluster.tests"), #"cluster.up.diff", "cluster.down.diff", 
    options = layersControlOptions(collapsed = FALSE)) %>%
  # add different legends
  addLegend(pal = factpal.up, 
            values = la@data$cluster.up,
            opacity = 1,
            group = "cluster.up",
            title = "Upload clusters") %>%
  addLegend(pal = factpal.down, 
            values = la@data$cluster.down,
            opacity = 1,
            group = "cluster.down",
            title = "Download clusters") %>%
  #addLegend(pal = factpal.up.diff, 
  #          values = la@data$cluster.up.diff,
  #          opacity = 1,
  #          group = "cluster.up.diff",
  #          title = "Upload diff. clusters") %>%
  #addLegend(pal = factpal.down.diff, 
  #          values = la@data$cluster.down.diff,
  #          opacity = 1,
  #          group = "cluster.down.diff",
  #          title = "Download diff. clusters") %>%
  addLegend(pal = factpal.tests, 
            values = la@data$cluster.tests,
            opacity = 1,
            group = "cluster.tests",
            title = "Test clusters")

```