### 2021-03-04 paper/v2_taylor_francis/response_reviewers.md
v.2

---
output:
  pdf_document: default
---
Dear Prof. Bian,

Thank you for providing the opportunity to revise and resubmit our paper.
We found the reviewer comments very useful and constructive and we believe they enabled us to improve the paper.
We were also very pleased to see that the reviewers found our paper ...

blah blah

Kind Regards,

[removed for anonymity]

## Reviewer: 1

>The paper reports research of detecting clusters of network access speed, especially under the pandemic for resilience. Such a topic is of great importance as we are all in the pandemic and we rely on the network cyberinfrastructure for work as stated, and also for online education for all levels of students. Therefore, it worth supports with a few major revisions needed:

> 1. The VGI approach of obtaining network speed maybe bias given that this probably is the digital guru group that will respond, maybe a discussion on this and picture the steps of selecting/soliciting and how representative this is.

> 2. The results for network speed for the different regions are not very different with 1-2 out of 10Mbps difference and probably can not reflect the social disparities. Not sure if this is the facts in U.K. If so, then it's not a problem?

> 3. The relevance to pandemic is not very straightforward. This may be improved by adding some studies, e.g., how the clusters impact the effectiveness of work from home through survey. Or relates to the employment rate.

> 4. The time series may be enhanced by putting a time range according to different events of the pandemic or other factors that relates to the pandemic.

> 5. Would suggest a review of the spatiotemporal challenges of the covid-19.

> 6. A review of cyberinfrastructure and how network varies in different region would be useful.

> 7. The first part introduction can be shorted to reflect the importance of network bandwidth to resilience of pandemic impacted life.

## Reviewer: 2

>This article presents interesting and timely research in digital geography, which is directly relevant to digital divide associated with the COVID-19 pandemic. The pandemic provides a once-in-a-lifetime opportunity to examine the barriers of broadband usage for telecommuting. The first two sections were very well written. I enjoyed reading them. However, I have a few major concerns on the methodology and a lack of discussion.

> 1. The paper used k-means, a popular unsupervised classification in this study. The optimal number of clusters was determined as 13 based on cluster validation indices (CVIs). While this number could be well justified, I feel it is too high to properly interpret the results. The interpretation and knowledge discovery from clustering results with a high k tend to be tedious. Being affected by the high k, the Results section appears to be very difficult to read and follow. As there are many different CVIs, I feel a lower k will benefit the interpretation of results.

ET

> 2. Another major comment is on the multinomial logit regression (MLR). This is partly related to my first comment on a larger k. While there is no hard limit on the number of categories in MLR, the interpretation of results will be difficult with a high k as well. In addition, if the dependence among independent variables (collinearity) occurs, the estimation of the multinomial logit model parameters becomes inaccurate. According to Table 2, it is very likely some of the variables could present collinearity (e.g., distance to nearest metro and tech jobs).

ET

> 3. Although Fig. 1 and 2 exhibit different speed tests patterns in 2019 and 2020, the data mining outcome lacks a baseline (prior to 2020) to compare with. It is very likely that the clusters in 2019 would be different from that in 2020. It will make the results more robust. Of course, I will not push this given the page limit. But some insights may be helpful in the Discussion (however, a Discussion section is missing).

ET: after we finalise k, we can rerun k-means for 2019 and decide

> 4. No discussion was included in this manuscript. I expect some insights into various important issues, such as problem causes, any broadband policies and socioeconomic solutions, equitable development, and research uncertainties.

> 5.  There are three levels of digital divide. According to the last paragraph in Section 2, only the intersection of level-1 and level-3 was included in this research. I am wondering how level-2 may fit in the general research framework. Could education and demographic variables be used to represent level-2 digital divide?

> 6. A few suggestions on the figures. Fig 1 and 2 have crowded tick labels for x-axis month and weekday. Fig 3 and 4 may be placed side by side with the same y-axis for easy comparison. Actually, I feel “Figure 3 shows that the high standard deviation” may be inaccurate as Fig 4 appears to exhibit higher variability. Fig 5 maps may include the locations of metropolitan or major cities to help interpretation. The warmer-cooler color scheme may match the population size instead of the ordinal cluster numbers. The current color scheme does not help interpretation at all.

> 7. A few minor things. In Sec 2.2, the statement “These returns have been even greater …” appears to be controversial. LAD acronym appears before it was defined. Virgin Media as a major broadband provider should be briefly introduced in the manuscript.