### 2021-03-11 src/LAs_Clusters_k9.Rmd
regressions for k=9

---
title: "LAs_clusters k = 9"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: FALSE
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(rprojroot)
library(lubridate)
library(knitr)
library(rprojroot)
library(nnet)
library(stargazer)
library(openair)
library(rgdal)
library(corrplot)
library(DescTools)
library(geosphere)
library(rgeos)
library(raster)
library(httr)
library(readxl)
library(httr)

# This is the project path
path <- find_rstudio_root_file()
```

## Load cluster data

```{r include = FALSE}

# load the data created by Data_Spatial.Rmd and ts_clusters.Rmd
path.cluster <- paste(path, "/data/temp/clusters_nodiff_9.csv", sep = "")
clusters <- read_csv(path.cluster)
names(clusters) [1] <- "LAD19NM"

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")
TSbb19_20sp <- read_csv(path.data)

#split out 2020 before join to clusters to create descriptives
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) 
TS2019 <- selectByDate(TSbb19_20sp, year = 2019)

# LAD codes and names
# get LA 
# this is the heaviest version:
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
# path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
# la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC: 
la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
codes <- la@data %>%
  dplyr::select(lad19cd, lad19nm) %>%
  rename(LAD19NM = lad19nm)

# LAD broadband characteristics AM
LAD.up.am19 <- TS2019 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.up.am19 = mean(speedup))

LAD.up.am20 <- TS2020 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.up.am20 = mean(speedup))

LAD.down.am19 <- TS2019 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.down.am19 = mean(speeddown))

LAD.down.am20 <- TS2020 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.down.am20 = mean(speeddown))

LAD.tests.am19 <- TS2019 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.tests.am19 = n())

LAD.tests.am20 <- TS2020 %>%
  filter(hour == 9 | hour == 10) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.tests.am20 = n())

# LAD broadband characteristics PM
LAD.up.pm19 <- TS2019 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.up.pm19 = mean(speedup))

LAD.up.pm20 <- TS2020 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.up.pm20 = mean(speedup))

LAD.down.pm19 <- TS2019 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.down.pm19 = mean(speeddown))

LAD.down.pm20 <- TS2020 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.down.pm20 = mean(speeddown))

LAD.tests.pm19 <- TS2019 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.tests.pm19 = n())

LAD.tests.pm20 <- TS2020 %>%
  filter(hour == 19 | hour == 20) %>%
  group_by(LAD19NM) %>%
  summarise(LAD.tests.pm20 = n())

# this is a share of Virgin provider variable
virgin <- TS2020 %>%
  filter(provider == "Virgin Media") %>%
  group_by(LAD19NM) %>%
  summarise(virgin = n())

virgin.denom <- TS2020 %>%
  #filter(provider == "Virgin Media") %>%
  group_by(LAD19NM) %>%
  summarise(virgin.denom = n())

virgin <- merge(virgin.denom, virgin, by = "LAD19NM", all.x = T)
virgin$virgin <- ifelse(is.na(virgin$virgin), 0, virgin$virgin)
virgin$virgin <- virgin$virgin/virgin.denom$virgin.denom
virgin$virgin.denom <- NULL

# LAD characteristcs merge
lad <- list(codes,
            clusters,
            LAD.up.am19,
            LAD.up.am20,
            LAD.down.am19,
            LAD.down.am20,
            LAD.tests.am19,
            LAD.tests.am20,
            LAD.up.pm19,
            LAD.up.pm20,
            LAD.down.pm19,
            LAD.down.pm20,
            LAD.tests.pm19,
            LAD.tests.pm20,
            virgin)

sapply(lad, function(x) dim(x))
sapply(lad, function(x) sum(is.na(x)))

# merge with reduce
lad <- lad %>% 
  reduce(inner_join, by = "LAD19NM")
sapply(lad, function(x) sum(is.na(x)))
```

## furlough

```{r}
url <- "https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/891607/Coronavirus_Job_Retention_Scheme_Statistics_Tables_June_2020.xlsx"
GET(url, write_disk(tf <- tempfile(fileext = ".xlsx")))
f <- readxl::read_excel(tf, 6L, range = "B7:D486", col_names = F) %>% 
  rename(lad19cd = 1,
         lad19nm = 2,
         furlough = 3) %>% 
  filter(!is.na(furlough) & !is.na(lad19cd)) %>% 
  dplyr::select(-lad19nm)
```

## Load socio-econ data and merge

```{r}
data.old.path <- paste0(path, "/data/temp/data_for_aux.csv")
data <- read_csv(data.old.path) %>% 
  dplyr::select(-c(1,4:19)) %>% 
  left_join(x = lad, y = . ) %>% 
  left_join(x = ., y = f )

# NOT RUN
# export file for future reference / backup
# data.out.path <- paste0(path, "/data/temp/data_for_aux_k9.csv")
# write.csv(data, data.out.path)
```

## Descriptive statistics and correlations

There is surprisingly low correlation between upload and download, morning and night. 
Number of tests are the most correlated variables.

```{r correlations, results = 'asis'}
data.cor <- data %>%
  dplyr::select(LAD.up.am19, LAD.down.am19,  
                LAD.tests.am19, LAD.up.pm19, 
                LAD.down.pm19, LAD.tests.pm19, 
                LAD.up.am20, LAD.down.am20,
                LAD.tests.am20, LAD.up.pm20, 
                LAD.down.pm20, LAD.tests.pm20)

m <- cor(data.cor)
corrplot(m, type="upper",method = "number", number.cex = .5, tl.cex = .75)

# data.cor <- data %>%
#   dplyr::select(LAD.tests.am20,
#                 pop, pop16_64,
#                 managers, tech,
#                 skilled, earnings,
#                 total.busi)

data.cor <- data %>%
  dplyr::select(#pop16_64,
                pop,
                #pop.dens,
                job.dens2018, 
                distMet,
                distLondon,
                south,
                managers, 
                tech,
                skilled,
                prof,
                admin, 
                caring,
                plant,
                earnings,
                total.busi,
                nvq4,
                furlough,
                LAD.tests.am20,
                virgin) %>%
  mutate('LAD.tests.am20/pop' = LAD.tests.am20/pop) %>%
  mutate('total.busi/pop' = total.busi/pop) %>%
  mutate('furlough/pop' = furlough/pop) %>% 
  dplyr::select(-LAD.tests.am20,-pop, -total.busi, -furlough)

m <- cor(data.cor, use = "pairwise.complete.obs")
corrplot(m, type="upper",method = "number", number.cex = .5, tl.cex = .75)

stargazer(data.cor,
          type="html",
          #type="text", 
          summary = T,
          nobs = T)
```

## Auxiliary regression, dep. var.: upload clusters

```{r upload regression , results= 'asis', message=F, error=F, warning=F}
# helpful sources: http://www.princeton.edu/~otorres/LogitR101.pdf
# https://thestatsgeek.com/2014/02/08/r-squared-in-logistic-regression/

# convert dependent variable to factor
data$cluster.up <- as.factor(data$cluster.up)
data$cluster.up <- relevel(data$cluster.up, "4") # base category: cluster 4

aux.up <- multinom(cluster.up ~ 
                     #pop16_64 + # non sign, while pop is
                     pop +
                     #pop.dens + 
                     job.dens2018 + # pop.dens is still insignificant when job.dens is excluded
                     distMet + distLondon + south +
                     managers + tech + skilled + prof + admin + caring + plant +
                     earnings + I(total.busi/pop) + nvq4 + I(furlough/pop) +
                     I(LAD.tests.am20/pop) + # non sensitive to addition of other tests pm
                     virgin
                   , data = data, 
                   trace = F) # to avoid messages

# McFadden's R squared
aux.up.null <- multinom(cluster.up ~ 1, data = data)
r2.up <- 1-logLik(aux.up)/logLik(aux.up.null)

n.obs <- dim(aux.up$fitted.values)[1]

stargazer(aux.up, 
          type = "html",
          #type = "text",
          nobs = T,
          digits = 3,
          covariate.labels=c(#"pop. 16-64, 2018",
            "pop, 2018",
            #"pop. density, 2018",
            "job density, 2018",
            "distance to nearest met. area",
            "distance to London",
            "South of the UK",
            "% of managerial jobs, 2020",
            "% of tech jobs, 2020",
            "% of skilled trade jobs, 2020",
            "% of professional jobs, 2020",
            "% of administrative jobs, 2020",
            "% of leisure jobs, 2020",
            "% of machine operation jobs, 2020",
            "earnings, 2019",
            "n. business est. per hab., 2019",
            "% of NVQ4+",
            "% furloughed pop., 2020",
            "n. of AM broadband tests per hab., 2020",
            "% of Virgin Media connections"),
          add.lines = list(c("McFadden's R squared",
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3)
                             , round(r2.up[1],3), round(r2.up[1],3)),
                           (c("N",
                              n.obs, n.obs, n.obs,
                              n.obs, n.obs, n.obs,
                              n.obs, n.obs, n.obs
                              , n.obs, n.obs))))

```

## Auxiliary regression, dep. var.: upload clusters, NO NVQ4

```{r upload regression no NVQ4, results= 'asis', message=F, error=F, warning=F}
aux.up <- multinom(cluster.up ~ 
                     #pop16_64 + # non sign, while pop is
                     pop +
                     #pop.dens + 
                     job.dens2018 + # pop.dens is still insignificant when job.dens is excluded
                     distMet + distLondon + south +
                     managers + tech + skilled + prof + admin + caring + plant +
                     earnings + I(total.busi/pop) + I(furlough/pop) +
                     I(LAD.tests.am20/pop) + # non sensitive to addition of other tests pm
                     virgin
                   , data = data, 
                   trace = F) # to avoid messages

# McFadden's R squared
aux.up.null <- multinom(cluster.up ~ 1, data = data)
r2.up <- 1-logLik(aux.up)/logLik(aux.up.null)

n.obs <- dim(aux.up$fitted.values)[1]

stargazer(aux.up, 
          type = "html",
          #type = "text",
          nobs = T,
          digits = 3,
          covariate.labels=c(#"pop. 16-64, 2018",
            "pop, 2018",
            #"pop. density, 2018",
            "job density, 2018",
            "distance to nearest met. area",
            "distance to London",
            "South of the UK",
            "% of managerial jobs, 2020",
            "% of tech jobs, 2020",
            "% of skilled trade jobs, 2020",
            "% of professional jobs, 2020",
            "% of administrative jobs, 2020",
            "% of leisure jobs, 2020",
            "% of machine operation jobs, 2020",
            "earnings, 2019",
            "n. business est. per hab., 2019",
            "% furloughed pop., 2020",
            "n. of AM broadband tests per hab., 2020",
            "% of Virgin Media connections"),
          add.lines = list(c("McFadden's R squared",
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3),
                             round(r2.up[1],3), round(r2.up[1],3), round(r2.up[1],3)
                             , round(r2.up[1],3), round(r2.up[1],3)),
                           (c("N",
                              n.obs, n.obs, n.obs,
                              n.obs, n.obs, n.obs,
                              n.obs, n.obs, n.obs
                              , n.obs, n.obs))))

```

```{r}
path.out <- paste0(path, "/data/temp/LAs_Clusters_k9.RData")
save.image(path.out)
```