### 2021-03-22 paper/v2_taylor_francis/broadband.speed.covid.tex
ET changes / responses to reviewer comments

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% see https://stackoverflow.com/a/47122900

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{dcolumn}
\def\tightlist{}

\begin{document}

\articletype{}

\title{Working from home and digital divides: resilience during the pandemic}

\author{\name{Hannah Budnitz$^{a}$, Emmanouil Tranos$^{b, c}$}
\affil{$^{a}$Transport Studies Unit, University of Oxford, UK; $^{b}$School of Geographic Sciences, University of Bristol, UK; $^{c}$The Alan Turing Institute, UK}
}

\thanks{CONTACT Hannah Budnitz. Email: , Emmanouil Tranos. Email: \href{mailto:e.tranos@bristol.ac.uk}{\nolinkurl{e.tranos@bristol.ac.uk}}}

\maketitle

\begin{abstract}
This paper offers a new perspective on telecommuting from the viewpoint
of the complex web of digital divides. Using the UK as a case study,
this paper studies how the quality and reliability of internet services,
as reflected in \emph{experienced} internet upload speeds during the
spring 2020 lockdown, may reinforce or redress the spatial and social
dimensions of digital divisions. Fast, reliable internet connections are
necessary for the population to be able to work from home. Although not
every place hosts individuals in occupations which allow for
telecommuting nor with the necessary skills to effectively use the
internet to telecommute, good internet connectivity is also essential to
local economic resilience in a period like the current pandemic.
Employing data on individual broadband speed tests and state-of-the-art
time-series clustering methods, we create clusters of UK local
authorities with similar temporal signatures of experienced upload
speeds. We then associate these clusters of local authorities with their
socioeconomic and geographic characteristics to explore how they overlap
with or diverge from the existing economic and digital geography of the
UK. Our analysis enables us to better understand how the spatial and
social distribution of both occupations and online accessibility
intersect to enable or hinder the practice of telecommuting at a time of
extreme demand.
\end{abstract}

\begin{keywords}
covid; internet; working from home; broadband speed; time-series
clusters
\end{keywords}

\hypertarget{sec:1}{%
\section{Introduction}\label{sec:1}}

During the pandemic, working from home using digital technologies,
whether partially or exclusively, was transformed from a niche means of
accessing work, albeit one that had been on a slow, upward trend, to a
widespread way of life in many countries. The ability to work from home
or telecommute meant millions retained their jobs and, to a varying
extent, maintained productivity during periods of strict lockdowns
around the world. However, this ability has not been evenly distributed
socially or spatially, creating new intersections of economic and
digital divisions. On one side are those who can work from home,
supported by digital technologies, and have thus been able to enjoy both
economic resilience and greater personal safety. On the other side,
previously employed individuals have been forced to accept furlough or
redundancy packages unless they are part of the cadre of essential
workers, who are potentially at high risk of infection. Whilst the basis
for this pandemic-generated divide has been viewed as mainly
occupational, here we consider whether it is also technological and
geographical.

The capability to work from home has previously been studied from the
perspective of whether work tasks in a given occupation both can be and
are allowed to be performed using digital technologies independently of
location or co-location with colleagues, including supervisors
\citep{allen2015effective, singh2013modeling}. However, successful
telecommuting also requires that the quality and reliability of digital
services, particularly home internet connection speeds, enable the
completion of work tasks with a minimum of delay or interruption. High
levels of demand are one of the main causes of reduced reliability and
slower connection speeds, as network bandwidth becomes congested. Yet
prior to the pandemic, the performance of broadband services with
respect to telecommuters was never tested at scale, as working from home
and connecting to colleagues and workplace resources via the internet
was the purview of a small minority of workers. Instead, leisure use in
the evening, when video streaming services are at their peak, has been
used to benchmark broadband performance and service delivery by
different Internet Service Providers (ISPs) \citep{ofcom2017}.

The shift towards telecommuting during various stages of lockdown around
the world has been drastic and there are speculations that post-Covid,
the tendency to work from home will be much higher, raising questions
around whether internet services can accommodate the increased demand.
For example, \(47\)\% of people in employment in the UK worked solely
from home in April \(2020\), whilst the same figure only reached \(5\)\%
the year before \citep{ons2020, ons2020lm2019}. Similar figures have
been reported for other countries \citep{felstead2020homeworking}.
Approximately \(37\)\% of the European workforce worked from home in
April \(2020\) with countries like Finland reaching \(60\)\%
\citep{eurofound2020}. In the US, almost half of the working population
worked from home during the same period because of the pandemic
\citep{brynjolfsson2020covid}, and a recent estimate indicated that
\(37\)\% of all jobs in the US can be permanently performed entirely
from home \citep{NBERw26948}. None of these changes could have happened
in the absence of reliable information and communication technology
(ICT) infrastructure -- both in terms of software and hardware. But
while software innovations are easily diffused across space and
society\footnote{See for example the huge success of videoconferencing
  apps such as Zoom \citep{marks2020zoom}.}, the same does not apply for
ICT hardware infrastructure such as internet broadband connectivity.

Using the UK as a case study, this paper aims to understand how the
quality and reliability of internet service, as reflected in
\emph{experienced} internet speeds during the spring 2020 lockdown, may
reinforce or redress the spatial and social dimensions of digital
divisions. We employ volunteered geographic data on individual broadband
speed tests and state-of-the-art time-series clustering methods to
create clusters of UK local authorities with similar temporal signatures
of experienced internet speeds. We then associate these clusters of
local authorities with their socioeconomic and geographic
characteristics to explore how they overlap with or diverge from the
existing economic and digital geography of the UK. As the quality of
internet infrastructure and services, as well as the concentration of
different occupations are spatially dependent and clustered in space,
our approach offers a framework for understanding the impact of and
interactions between digital divisions geographically and
socioeconomically. Our analysis enables us to better understand how the
spatial and social distribution of education, occupation and online
accessibility intersect to enable or hinder the practice of
telecommuting at a time of extreme demand. We also consider what lessons
can be learned from this time for a future where telecommuting is likely
to remain a more common means of accessing work, at least in comparison
to the pre-Covid era.

By asking how resilient broadband speeds, and particularly upload
speeds, are as experienced in different parts of the UK during a time of
extreme demand, we interrogate which places benefit from the greater
economic resilience digital technologies can offer. The structure of
this paper is as follows. First we review the literature on
telecommuting and digital divides to better understand their structural
and spatial development pre-pandemic, and thus their importance to the
economic resilience of different places. We then describe our data and
methodology. Our results section first offers a classification of how
internet services vary across clusters of UK local authorities and then
assesses whether these clusters replicate or repudiate other
socio-economic and geographic patterns of economic resilience. We
conclude with a discussion of the insights we have gained from our new
perspective on digital divisions.

\hypertarget{sec:2}{%
\section{Literature review}\label{sec:2}}

\hypertarget{sec:2.1}{%
\subsection{From telecommuting to \#WFH}\label{sec:2.1}}

In this analysis, the terms `telecommuting' and `working from
home\footnote{See also the popular social media hashtag \#WFM}' are used
interchangeably, as most remote labour during the Covid-19 crisis was
carried out in the homes of individual employees rather than any other
location \citep{eurofound2020}. However, previous research has explored
how telecommuting can occur in other places, including satellite offices
or on public transport \citep{felstead2012rapid, siha2006telecommuting}.
Previous research has also used a variety of definitions to measure the
level of telecommuting within different workforces, distinguishing
between those directly employed, indirectly employed, self-employed,
full-time or part-time, and those who use digital technologies to work
remotely full-days or part-days
\citep{allen2015effective, bailey2002review, haddad2009examination}. No
matter the definition, the option and capability to telecommute or work
from home has never been equally distributed spatially or
socio-economically any more than different industries and employment
opportunities have. Studies from the United States, the Netherlands, and
the UK found that telecommuters are most likely to hold professional,
managerial, and technical occupations where the workforce is better
educated and wealthier, and that there is suppressed demand among women
and part-time workers
\citep{headicar2016move, peters2004employees, singh2013modeling}.

Opportunities for working from home during the current pandemic have
likewise not been equally spread across the workforce.
\citet{NBERw26948} indicated that in the US, managers, educators, those
working in computer-related occupations, finance, and law can easily
work from home, and that occupations with opportunities to telecommute
are associated with higher earnings. This is not the case for the
workforce occupied in more spatially fixed occupations, from farming,
construction and manufacturing to hospitality and care services. In the
US, these occupations tend to be lower-income, non-white, without a
university degree, live in rental accommodation and lack health
insurance \citep{NBERw27085}. Similar trends can be observed for other
countries. For example, \(75\)\% of workers with tertiary education
worked from home in Europe during spring \(2020\), whilst only \(34\)\%
of workers with secondary education and \(14\)\% of those primary
education did so \citep{eurofound2020}.

\hypertarget{sec:2.2}{%
\subsection{Digital divides and economic resilience}\label{sec:2.2}}

The literature describes first level digital divides in terms of the
availability and quality of internet connectivity, such as that manifest
in different geographies in the UK
\citep{riddlesden2014broadband, philip2017digital}. Second level digital
divides consider the presence or lack of the necessary skills to
effectively utilise digital technologies and the internet
\citep{blank2014dimensions, van2011internet}. The third level focuses on
the heterogenous returns of internet usage among different socioeconomic
groups and, consequently, how digital technologies can assist in
bridging or further enhancing existing socioeconomic divides.
\citep{stern2009levels, van2014digital, van2015third}. The capability to
telecommute is related to all three levels of digital divides, but more
importantly leads to differentiated outcomes regarding the economic
resilience of people and places to overcome a systemic shock such as the
current pandemic.

Our understanding of telecommuting as a product of enabled occupations
can be described as a manifestation of the third level digital divide,
as those who are able to use digital technologies to work from home
benefit from a high rate of return on their use of the internet in terms
of autonomy, flexibility, and time saved from commuting
\citep{peters2004employees, siha2006telecommuting, singh2013modeling}.
In some aspects, these returns may have increased during the Covid-19
crisis, when those with the capability to telecommute also have the
ability to maintain their employment whilst protecting their health.
However, the success of these arrangements has been dependent upon the
first level digital divide, which is associated with access to and
quality of internet connectivity. \citet{SALEMINK2017360} provides a
systematic review of the pre-pandemic, first level digital divide in
infrastructure quality between urban and rural areas in various advanced
economies. Rural areas, predictably, fare worse. This finding broadly
applies in the UK, where studies that also analysed broadband speed
checks concluded that average speeds are lower in rural areas, something
that has not been improved by policy measures to increase competition
\citep{riddlesden2014broadband, nardotto2015unbundling}. Moreover,
\citet{riddlesden2014broadband} found that levels of deprivation did not
correlate with first level digital divides.

However, as \citet{blank2018local} highlight, variation in individual
internet uptake and use is a product of more complex spatial and
demographic characteristics than simple rurality or urbanisation. Dense
urban areas were shown to suffer more from slowdown during peak hours,
although these services were more likely to be improved by increased
competition between providers, such as between new entrants and Virgin
Media cable connections
\citep{riddlesden2014broadband, nardotto2015unbundling}. The latter were
historically available to only \(45\)\% of premises in the UK
\citep{ofcom2016}, where the more lucrative and competitive market
originally attracted the cable TV provider. Whether the variation in
infrastructure quality and reliability affects the spatial footprint of
telecommuting has not previously been investigated, in part because
telecommuting has not previously been a cause of concentrated demand and
pressure on internet services. Other intersections between digital
divides have been subject to study. There are indications that those who
purchase high speed connections consume more data of all sorts and use
their connections for a greater variety of purposes
\citep{hauge2011consumer}. There is also a correlation between access to
internet services and a reduction in household transport spend
\citep{bris2017ict}. These studies suggest that better internet services
enable households to make savings and efficiencies, an example of the
first level digital divide reinforcing the third level.

Currently there is a hotly debated discussion in the literature
regarding these exact changes in transportation and the structure of
cities. We now have enough hard data to observe the drastic change in
the space-time geography of cities around the world
\citep[e.g.~][]{google2020, shibayama2021impact}. Although there is a
broad agreement that these changes during the pandemic have played a
pivotal role in stopping the spread of the virus
\citep{jia2020population, yang2020taking, mu2020interplay}, we still
have no evidence to what extend these increased levels of working from
home, and the consequent decrease in commuting flows and the structure
of cities, will remain post-pandemic.

Multi-layered digital divides may also intersect with material divides
and the economic geography of the UK. The regional economic resilience
literature underlines the differentiated capacity of cities and regions
to escape or recover from economic crises
\citep{martin2012regional, kitsos2018economic}. As different places have
different industrial and occupational profiles, these affect their
aggregated potential capacity for telecommuting. Such profiles are
associated with longstanding inequalities in the UK and their spatial
representation as a North-South divide \citep{martin_north_south}.
Various studies have illustrated severe inequalities between the north
and the south regions of England in terms of skills and human capital,
unemployment, productivity and prosperity
\citep{lee2014grim, mccann2020perceptions, dorling2018peak}. Some
scholars have even argued that the UK suffers from some of the highest
levels of interregional inequalities in the global north
\citep{gal2018reducing, mccann2016uk}. All three levels of digital
divides are associated with or shaped by the geography of the UK. Yet
this is the first time that the intersection of digital and material
divides is relevant to understanding the economic resilience of places
and large swathes of the population, as digital technologies became an
essential tool of productivity during the Covid-19 pandemic.

The extreme demand during the pandemic thus provides a new opportunity
to understand how internet infrastructure quality, and reliability
affects telecommuting, particularly in light of the high volumes of
bandwidth-intensive video conferencing required in order to avoid the
face-to-face contact that could increase the spread of infection. We
seek to answer how internet service resilience could contributes to or
reduce economic resilience when the latter is dependent upon the
capability to work from home. We also aim to improve our understanding
of the impact of first level digital divisions on telecommuting, and
whether this results in more fundamental third level digital divisions
than has previously been perceived.

\hypertarget{sec:3}{%
\section{Methods and data}\label{sec:3}}

\hypertarget{sec:3.1}{%
\subsection{Time-Series clustering}\label{sec:3.1}}

Our chosen methodological framework is cluster analysis, which can be
defined within machine learning approaches as an unsupervised learning
task, partitioning unlabelled observations into homogeneous groups known
as clusters \citep{montero2014tsclust}. The key idea is that
observations within clusters tend to be more similar than observations
between clusters. Clustering is particularly useful for exploratory
studies as it identifies structures within the data
\citep{aghabozorgi2015time}. Cluster analysis is a widely used in
geography \citep{gordon1977classification, everitt1974cluster}, for
example to solve \emph{regionalisation} problems
\citep{niesterowicz2016}. Clustering methods are also the basis of
\emph{geodemographics}, a research domain which aims to create small
area indicators or typologies of neighbourhoods based on diverse
variables \citep{SINGLETON2009289, harris2005geodemographics}. These
studies usually employ cross-sectional data, and most clustering
problems in geography deal with observations that are fixed in time.
However, for this paper we are interested in internet speeds, which vary
over time. Therefore, we create clusters of local authorities in the UK
with similar temporal signatures of experienced internet speeds.

To do so, we employ time-series clustering methods, which have been
developed to deal with clustering problems linked to, for instance,
stock or other financial data, economic, governmental or medical data as
well as machine monitoring
\citep{aggarwal2013time, aggarwal2001surprising, hyndman2015large, WARRENLIAO20051857}.
The main challenge, which does not apply to cross-sectional clustering
problems, is data dimensionality, with a multiplicity of data points for
every individual object included in the data set, and how their value
changes dynamically as a function of time \citep{aghabozorgi2015time}.
This high dimensionality leads to (i) computational and algorithmic
challenges regarding handling these data and building algorithms to
perform clustering over long time-series, and (ii) open questions
regarding the choice of similarity measures in order to cluster similar
times-series objects together considering the whole length of the
time-series and overcoming issues around noise, outliers and shifts
\citep{lin2004iterative, aghabozorgi2015time}.

For this paper we utilise a category of time-series clustering methods
known as shape-based approaches. These methods match two separate
time-series objects based on the similarity of their shapes through the
calculation of distances between the shapes, and are better equipped to
capture similarities between short length time-series
\citep{aghabozorgi2015time}, such as our data. We thus identify clusters
of UK local authorities with similar temporal signatures -- i.e.~shapes
-- of experienced internet speeds. The clusters are identified using the
common partitioning algorithm, where no clusters overlap, known as
\emph{k}-means. This iterative algorithm is popular because of the
simplicity of the implementation and the interpretability of the
results. It begins with defining the desired number of clusters:
\emph{k}. Then each observation is randomly assigned to a cluster from
the \([1,k]\) space. This initial cluster assignment is followed by
iterations in order to minimise the distance between the centroids of
the clusters and the observations assigned to these clusters
\citep{james2013introduction}.

There are a number of differences between the application of
\emph{k}-means for cross-sectional and times-series data. Instead of
creating clusters around centroids, a common approach is to create
clusters around \emph{medoids}, which are representative time-series
objects with a minimal distance to all other cluster objects
\citep{sardatime}. Also, instead of calculating the Euclidean distance
between centroids and data points, more complex distance measures need
to be employed to capture the similarity between a time-series object
and a medoid. A common distance measure for shape-based time-series
clustering is Dynamic Time Warping (DTW), an algorithm comparing two
time-series objects to find the optimum warping path between them. DTW
is widely used in order to overcome limitations linked to the use of
Euclidean distance
\citep{sardatime, berndt1994using, ratanamahatana2004everything}. The
\texttt{R} package \texttt{dtwclust} has been used for the time-series
clustering \citep{dtwclust}.

\hypertarget{sec:3.2}{%
\subsection{Experienced Broadband Speeds}\label{sec:3.2}}

To assess the internet quality and reliability across local authorities
in the UK, we utilise unique data comprising individual internet speed
tests from Speedchecker Ltd\footnote{\url{https://www.broadbandspeedchecker.co.uk/}}.
This is a private company that allows internet users to check their own
broadband upload and download speeds, and stores every speed-check with
timestamp and geolocation information. These data have been used before
to assess digital divides \citep{riddlesden2014broadband} and the impact
of local loop unbundling regulatory processes
\citep{nardotto2015unbundling}, and we followed the former's approach to
remove outliers. By using this volunteered geographic data, we are able
to assess the internet speed \emph{experienced} by users, which may
differ from the maximum speeds \emph{advertised} by ISPs. Indeed, those
who seek to test their broadband are most likely to do so because they
are experiencing slower speeds than expected, although any skew towards
slower speeds is balanced by the likelihood that those who test their
broadband are also more `tech-savvy' and / or have purchased higher
speed packages that are not delivering the promised level of service.

We are particularly interested in upload speeds and the frequency of
speed tests over the period from March to May \(2020\), as government
statements indicate this encompasses the period when UK workers were
told to work from home if at all possible \citep{GovUK2020}. Schools and
various retail, leisure and hospitality businesses were closed from late
March, and restrictions were gradually eased from late May. Average
upload speeds are slower than average download speeds, at \(9.3\)Mb/s
mean upload speed for the whole sample, compared to \(29.6\)Mb/s for
download speeds, but they are also less associated with internet-based,
high-demand, leisure activities such as video streaming. Therefore,
upload speeds are more relevant to work-related activities such as
uploading documents or two-way audio, video, and text-based
communication systems.

The frequency of speed tests was important in identifying the temporal
profile which would give us most insight into experienced internet
service and resilience over units of time. Whilst there is an overall
trend of increased testing from March to April and then a slight
reduction from April to May, this trend masks substantial variation by
not only the day of the week, but also time of day, as can be seen in
Figure \ref{test2020}. Thus, a daily aggregation of upload speeds would
mask the variation in experienced service over the course of each
weekday. Furthermore, the importance of this variation is highlighted by
a comparison with the same period in \(2019\), as in Figure
\ref{test2019}, when the volume of testing and thus of experienced
internet service quality peaked in the evening, presumably in response
to demand for leisure activities and download speeds. In contrast, the
majority of the increase in testing in \(2020\) is during the working
day, creating a new morning peak in Figure \ref{test2020}. Therefore, we
include a measure of hourly variation in our temporal profiles to
reflect the change in users' perception of the workday reliability of
internet services.

However, there were insufficient observations -- only \(631\) speed
tests per Local Authority District (LAD) on average -- for each for each
working hour of each working day in each LAD to profile speeds at that
level of detail. Spatial aggregation was also necessary because we could
not follow individuals or households and connect data points. Therefore,
we aggregate the \(241,088\) individual, geolocated and time-stamped
speed-checks during the \(13\) weeks of March to May inclusive for
weekdays in \(2020\) by each hour of the day and day of the week. As our
research aims to identify the geography of internet service resilience
for work purposes, bank holidays and the hours between midnight and
\(6:00\) were excluded, as well as weekend days. The composite week
time-series thus comprise \(18\) hours multiplied by \(5\) weekdays or
\(90\) time points per series. The time-series were calculated for each
of the \(382\) LADs in the UK, standardised, and then a \emph{k}-means
partitioning around medoids clustering algorithm was applied using DTW.
We initially run the algorithm for \(k \in \mathbb{N} \bigcap [5,15]\),
calculated the cluster validity indices (CVIs) and then run the
subsequent multinomial regression -- see the end of this section for
more details for the different \emph{k}. Following \citet{sardatime}, to
identify the optimal \emph{k} we used the majority vote for the
following CVIs: Silhouette (max), Score function (max),
Calinski-Harabasz (max), Davies-Bouldin (min), Modified Davies-Bouldin
(DB*, min), Dunn (max), COP (min). Nevertheless, we opted against using
the optimal \(k = 13\) solution as it was too large to allow for
communicable LAD clusters. Instead we opted for a smaller \(k = 9\),
which led to a rather similar spatial pattern and, importantly, to a
much higher R-squared in the subsequent explanatory regression (\(0.44\)
instead of \(0.34\)).

\begin{figure}
\centering
\includegraphics[width=0.85\textwidth,height=0.5\textheight]{figures/time.var.plot2020.png}
\caption{Speed tests over time, 2020 \label{test2020}}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.85\textwidth,height=0.5\textheight]{figures/time.var.plot2019.png}
\caption{Speed tests over time, 2019 \label{test2019}}
\end{figure}

In Section \protect\hyperlink{sec:4.1}{4.1}, we review the temporal
profile of upload speed by hour of the day and day of the composite week
for each of the clusters. Since the quality and reliability of internet
services vary in time and space due to both supply and demand-side
influences, we also use a number of different measures to describe
experienced upload speeds per cluster. These include: i) mean,
experienced connection speed, ii) standard deviation or the amount of
fluctuation from the mean, and iii) the variation in speeds during the
new morning peak of testing when working from home is more likely to
take place. We take account of all three measurements in order to
determine how resilient broadband speeds are as experienced in different
parts of the UK during a time of extreme demand.

The cause of these different experiences of broadband resilience may
vary between and within clusters, as they may reflect either patterns of
demand or quality of infrastructure. Our approach is also limited by
potential endogeneity, as for example, better quality connections with
high mean speeds may enable more working from home, but greater demand
can cause slower speeds, less reliability, or greater variability of
speed at different times of day or week. Therefore, we avoid attributing
any cause to our analysis of the experienced level of quality and
reliability of upload speeds. Instead, we run an auxiliary regression to
understand how the spatial and temporal patterns of internet service
relate to the economic geography of the UK. More specifically, we
estimate the following multinomial logit model:

\begin{align}
Pr(Y_{i}=j) = \frac{exp^{X_{i}\beta_{j}}}{\sum_{i=1}^j exp^{X_{i}\beta_{j}}}
\begin{cases}
    i = 1, 2, ... , N \\  
    j = 1, 2, ... , J
\end{cases}\label{eq1}
\end{align}

Based on the outcomes of the time-series clustering, we identify \(J\)
distinct and crisp clusters. We then regress this cluster membership
against a vector \(X_{i}\) of socio-economic and geographic variables,
which are discussed in detail in the relevant Section
\protect\hyperlink{sec:4.2}{4.2}. Because we cannot identify individuals
or households and consequently aggregated our data at the LAD level, our
results offer correlations between the socioeconomic characteristics of
certain geographic locations and internet service quality, not a record
of who was telecommuting. Such individual data could be found though
surveys, but these offer less detailed information about the experience
of internet resilience due to enforced demand, which is the main
contribution of this paper. Our auxiliary regression, therefore,
provides an indication of how internet connectivity can reinforce or
redress existing spatial and social inequalities in different places.
However, it opens a path to future research by highlighting the
importance of understanding of how telecommuting capabilities and
digital infrastructure divisions intersect.

\hypertarget{sec:4}{%
\section{Results}\label{sec:4}}

\hypertarget{sec:4.1}{%
\subsection{Upload Clusters / cluster description}\label{sec:4.1}}

The temporal profiles of the local authority clusters have been
summarised in Figure \ref{UpCluster} and Table \ref{up.cluster.descr},
to provide an overview of the quality and reliability of experienced
broadband in different parts of the UK. Figure \ref{UpCluster} shows a
composite profile of mean upload speeds per hour per day for each of the
largest six clusters, in terms of the LAD membership and population.

\begin{figure}
\includegraphics[width=0.95\linewidth]{figures/UpCluster} \caption{\label{UpCluster}Temporal profilies for upload speed clusters}\label{fig:unnamed-chunk-2}
\end{figure}

\begin{table}[!htbp] \centering 
  \caption{Upload speed cluster characteristics\label{up.cluster.descr}} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{0pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Cluster & N. of LADs & LAD population & mean speed & SD speed & mean AM speed & mean PM speed \\ 
\hline \\[-1.8ex] 
1 & 9 & 903200 & 9564 & 6314 & 8798 & 10457 \\ 
2 & 2 & 162000 & 12085 & 6537 & 11882 & 10866 \\ 
3 & 12 & 1785800 & 11047 & 6079 & 10029 & 11634 \\ 
4 & 1 & 91100 & 9689 & 6122 & 7816 & 9689 \\ 
5 & 3 & 280000 & 10802 & 6116 & 11010 & 10084 \\ 
6 & 229 & 40552800 & 8761 & 5847 & 8555 & 8955 \\ 
7 & 5 & 682500 & 10326 & 6102 & 10045 & 11149 \\ 
8 & 6 & 510000 & 9769 & 6352 & 8989 & 10836 \\ 
9 & 115 & 21467800 & 10328 & 5915 & 10283 & 10333 \\ 
\hline \\[-1.8ex] 
\multicolumn{7}{l}{Note: All speed measures are upload speeds} \\ 
\end{tabular} 
\end{table}

The largest cluster, comprising \(229\) local authorities and over
\(40\) million people, is cluster \(6\), which has the slowest aggregate
mean upload speed of any cluster, and the highest ratio of the standard
deviation to the mean. This suggests that those living in local
authorities in this cluster experienced some of the lowest quality
broadband services in terms of upload speeds and reliability in the UK.
However, as shown in Figure \ref{map.up.clusters}, some of the most
rural areas of the UK are included in this cluster. If these areas
suffer most from first level digital divides as described in the
literature review, the low speeds in rural areas might be pulling down
the averages in other areas in this largest cluster. Also, since the
areas are clustered by their temporal profile across the working week,
the graph in Figure \ref{UpCluster} indicates that speeds in cluster
\(6\) are some of the more reliable. Table \ref{up.cluster.descr}
confirms that upload speeds in cluster \(6\) during the morning peak
from \(9:00\)-\(10:59\) were, on average, only \(4.5\)\% slower than in
the evening peak period between \(19:00\) and \(20:59\), when
entertainment purposes are likely to be using the most bandwidth.

\begin{figure}
\includegraphics[width=0.95\linewidth]{figures/map.up.clusters} \caption{\label{map.up.clusters}Upload speed clusters for LADs}\label{fig:unnamed-chunk-4}
\end{figure}

Those living in the second largest cluster -- \(9\), with \(115\) LADs
and \(21.5\) million people -- experienced aggregate mean upload speeds
of more than \(1.5\)Mb/s faster than those in cluster \(6\), a lower
ratio of standard deviation to mean, and almost the same upload speeds
in the AM Peak as the PM Peak. Indeed, the temporal profile for cluster
\(9\) in Figure \ref{UpCluster}, like that for cluster \(6\), is fairly
flat. Whilst it is likely that the large numbers of tests being
performed by the large populations in these two clusters of LADs result
in less varied averages to show in the graph, the method of shape-based
clustering suggests that these results indicate that the majority of the
population of the UK experience less temporal variation, and so belong
in these two clusters. The results also confirm that those in cluster
\(9\) experienced a consistently better service during lockdown than
those in cluster \(6\), in terms of both average speed and reliability.

Of the smaller clusters, cluster \(7\) experiences quite similar mean
speeds to cluster \(9\), and a not too dissimilar standard deviation,
but mean speeds in the morning peak are almost \(10\)\% lower than the
evening peak. Cluster \(3\) boasts speeds almost \(1\)Mb/s higher again,
but suffers from even greater slowdown in the morning peak. In
comparison, clusters \(2\) and \(5\), home to a little less than half a
million people, experience not only above average mean speeds, but also
faster speeds in the morning peak compared to the evening. Thus, in the
five LADs in these two clusters, the temporal profile of internet use
may be closer to what might have been expected pre-pandemic. In
comparison, the \(15\) LADs in clusters \(1\) and \(8\), home to over
\(1.4\) million people, not only experience fairly average upload speeds
and high ratios of standard deviation to the mean, but also experience
much lower speeds during the morning peak than the evening peak. From
the large spikes and dips shown on Figure \ref{UpCluster}, it is likely
that this poor reliability or consistency of morning internet speeds was
more noticeable to those in cluster \(1\), whilst those in cluster \(8\)
experienced the lack of reliable service as a problem throughout the
day.

In summary, the reliability of internet services during the working day
appears to have altered for the vast majority of locations due to
increased use from residents told to stay at home, and, if they were
fortunate, work from home. Yet this experience has been different in
different locations. In particular, LADs in cluster \(9\) experienced
higher speeds and more resilient broadband internet than those in the
largest cluster, \(6\). Those in clusters \(3\) and \(7\) also
experience higher mean speeds and better service reliability, and can be
more confident that they are on the right side of the first level
digital divide, and that their more resilient ICT infrastructure and
services can robustly support higher levels of telecommuting. On the
other hand, whilst high levels of telecommuting may be the cause of
morning speeds over 15\% lower than those in the evening, those in
clusters \(1\) and \(8\) also experience lower mean speeds than clusters
\(3\), \(7\), and \(9\), but not as low as cluster \(6\). Whilst the
differences may not be large, the lower upload speeds in cluster \(6\)
or the poor reliability in clusters \(1\) and \(8\) could still have had
implications for economic resilience if combined with other types of
digital divide. This potential will be explored in the next section.

\hypertarget{sec:4.2}{%
\subsection{Post-clustering regression analysis}\label{sec:4.2}}

\begin{sidewaystable}[!htbp] \centering 
  \caption{Auxiliary multinomial regression of upload speed clusters on socio-economic and geographic LAD variables\label{aux}} 
  \label{} 
\tiny 
\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
\\[-1.8ex] & 1 & 2 & 3 & 5 & 6 & 7 & 8 & 9 \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 pop, 2018 & $-$0.00004$^{**}$ & 0.00002 & 0.00001 & 0.00002 & 0.00001 & 0.00002 & $-$0.00000 & 0.00001 \\ 
  & (0.00002) & (0.00001) & (0.00001) & (0.00002) & (0.00001) & (0.00001) & (0.00002) & (0.00001) \\ 
  & & & & & & & & \\ 
 job density, 2018 & $-$0.534$^{***}$ & $-$0.362$^{***}$ & 4.211$^{***}$ & $-$0.890$^{***}$ & 3.516$^{***}$ & $-$8.624$^{***}$ & $-$3.259$^{***}$ & 3.337$^{***}$ \\ 
  & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) \\ 
  & & & & & & & & \\ 
 distance to nearest met. area & $-$0.007$^{***}$ & 0.013$^{***}$ & 0.002$^{***}$ & 0.023$^{***}$ & $-$0.001 & $-$0.158$^{***}$ & 0.005$^{***}$ & $-$0.003$^{**}$ \\ 
  & (0.0004) & (0.00001) & (0.0003) & (0.0002) & (0.002) & (0.0001) & (0.0003) & (0.002) \\ 
  & & & & & & & & \\ 
 distance to London & 0.011$^{***}$ & 0.004$^{***}$ & 0.005$^{**}$ & $-$0.013$^{***}$ & 0.010$^{***}$ & 0.012$^{***}$ & 0.008$^{***}$ & 0.008$^{***}$ \\ 
  & (0.001) & (0.00004) & (0.002) & (0.001) & (0.001) & (0.001) & (0.002) & (0.001) \\ 
  & & & & & & & & \\ 
 South of the UK & $-$3.628$^{***}$ & 0.958$^{***}$ & 4.936$^{***}$ & $-$10.243$^{***}$ & 4.830$^{***}$ & 3.919$^{***}$ & $-$1.140$^{***}$ & 3.188$^{***}$ \\ 
  & (0.000) & (0.000) & (0.00000) & (0.00000) & (0.00001) & (0.00000) & (0.00000) & (0.00001) \\ 
  & & & & & & & & \\ 
 managerial jobs, 2020 & 0.944$^{***}$ & 0.628$^{***}$ & 0.593$^{***}$ & $-$0.544$^{***}$ & 0.634$^{***}$ & $-$0.128$^{***}$ & 0.725$^{***}$ & 0.511$^{***}$ \\ 
  & (0.00002) & (0.00000) & (0.00003) & (0.00003) & (0.00002) & (0.00003) & (0.00003) & (0.00002) \\ 
  & & & & & & & & \\ 
 tech jobs, 2020 & 0.350$^{***}$ & $-$0.222$^{***}$ & $-$0.068$^{***}$ & $-$0.151$^{***}$ & 0.018$^{***}$ & $-$0.139$^{***}$ & 0.143$^{***}$ & $-$0.071$^{***}$ \\ 
  & (0.00002) & (0.00000) & (0.00002) & (0.0001) & (0.00003) & (0.0001) & (0.00003) & (0.00003) \\ 
  & & & & & & & & \\ 
 skilled trade jobs, 2020 & 0.521$^{***}$ & 0.171$^{***}$ & 0.214$^{***}$ & $-$0.578$^{***}$ & 0.231$^{***}$ & 1.067$^{***}$ & 0.284$^{***}$ & 0.049$^{***}$ \\ 
  & (0.00002) & (0.00000) & (0.00004) & (0.00004) & (0.00004) & (0.0001) & (0.00003) & (0.00003) \\ 
  & & & & & & & & \\ 
 professional jobs, 2020 & $-$0.577$^{***}$ & $-$0.962$^{***}$ & $-$0.565$^{***}$ & $-$0.860$^{***}$ & $-$0.566$^{***}$ & $-$1.139$^{***}$ & $-$0.389$^{***}$ & $-$0.702$^{***}$ \\ 
  & (0.00003) & (0.00000) & (0.00004) & (0.0001) & (0.00004) & (0.0001) & (0.0001) & (0.00004) \\ 
  & & & & & & & & \\ 
 administrative jobs, 2020 & 0.200$^{***}$ & $-$1.324$^{***}$ & $-$0.050$^{***}$ & 0.136$^{***}$ & $-$0.137$^{***}$ & $-$0.101$^{***}$ & $-$0.018$^{***}$ & $-$0.155$^{***}$ \\ 
  & (0.00002) & (0.00000) & (0.00002) & (0.00004) & (0.00002) & (0.00005) & (0.00002) & (0.00002) \\ 
  & & & & & & & & \\ 
 service jobs, 2020 & $-$0.082$^{***}$ & $-$0.180$^{***}$ & $-$0.659$^{***}$ & $-$1.066$^{***}$ & $-$0.746$^{***}$ & $-$0.944$^{***}$ & $-$0.773$^{***}$ & $-$0.783$^{***}$ \\ 
  & (0.00002) & (0.00000) & (0.00003) & (0.00002) & (0.00002) & (0.00003) & (0.00002) & (0.00001) \\ 
  & & & & & & & & \\ 
 machine operation jobs, 2020 & 0.210$^{***}$ & 0.574$^{***}$ & 0.439$^{***}$ & $-$0.589$^{***}$ & 0.229$^{***}$ & $-$0.260$^{***}$ & 0.080$^{***}$ & 0.127$^{***}$ \\ 
  & (0.00001) & (0.00000) & (0.00002) & (0.00002) & (0.00001) & (0.00002) & (0.00001) & (0.00001) \\ 
  & & & & & & & & \\ 
 earnings, 2019 & $-$0.003$^{***}$ & 0.029$^{***}$ & 0.019$^{***}$ & 0.069$^{***}$ & 0.018$^{***}$ & 0.050$^{***}$ & 0.016$^{***}$ & 0.026$^{***}$ \\ 
  & (0.001) & (0.0001) & (0.001) & (0.002) & (0.001) & (0.002) & (0.001) & (0.001) \\ 
  & & & & & & & & \\ 
 n. business est. per hab., 2019 & 1.930$^{***}$ & $-$0.377$^{***}$ & 0.826$^{***}$ & 0.251$^{***}$ & 0.315$^{***}$ & 1.515$^{***}$ & $-$0.948$^{***}$ & $-$4.011$^{***}$ \\ 
  & (0.00000) & (0.000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) \\ 
  & & & & & & & & \\ 
 furloughed per hab., 2020 & $-$1.518$^{***}$ & $-$0.152$^{***}$ & 11.014$^{***}$ & 0.972$^{***}$ & $-$26.969$^{***}$ & 2.059$^{***}$ & 4.236$^{***}$ & 10.500$^{***}$ \\ 
  & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) \\ 
  & & & & & & & & \\ 
 AM tests per hab., 2020 & 0.002$^{***}$ & $-$0.011$^{***}$ & $-$0.140$^{***}$ & 0.015$^{***}$ & 0.231$^{***}$ & 0.005$^{***}$ & $-$0.025$^{***}$ & $-$0.078$^{***}$ \\ 
  & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) \\ 
  & & & & & & & & \\ 
 Virgin Media \%, 2020 & $-$0.349$^{***}$ & 12.236$^{***}$ & 5.619$^{***}$ & 0.489$^{***}$ & 1.721$^{***}$ & $-$16.289$^{***}$ & $-$5.581$^{***}$ & 6.126$^{***}$ \\ 
  & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) \\ 
  & & & & & & & & \\ 
 Constant & $-$4.570$^{***}$ & $-$1.045$^{***}$ & $-$6.539$^{***}$ & 5.629$^{***}$ & 2.957$^{***}$ & 3.386$^{***}$ & $-$2.531$^{***}$ & 2.979$^{***}$ \\ 
  & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) & (0.00000) \\ 
  & & & & & & & & \\ 
\hline \\[-1.8ex] 
McFadden's R squared & 0.435 & 0.435 & 0.435 & 0.435 & 0.435 & 0.435 & 0.435 & 0.435 \\ 
N & 322 & 322 & 322 & 322 & 322 & 322 & 322 & 322 \\ 
Akaike Inf. Crit. & 742.901 & 742.901 & 742.901 & 742.901 & 742.901 & 742.901 & 742.901 & 742.901 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{8}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{sidewaystable}

Using an auxiliary multinomial logit regression, we test whether the
clusters that have higher mean speeds and more reliable services do
indeed consist of LADs that are more urban or closer to major urban
areas and are more likely to benefit from a choice of high quality
internet services. This part of the regression aims to confirm any first
level digital divides. Next, to better understand how the clusters fare
in relation to the second level digital divide, we consider which LADs
in which clusters have a higher proportion of occupations where the
nature of the work and the skills that occupation employs enable
telecommuting. Finally, we consider the rate of return on internet use,
or third level digital divides, by reviewing which clusters have the
highest earnings and job density, and also which clusters experienced
the highest share of population furloughed during this period in the
pandemic.

The results of the auxiliary regression are presented in Table
\ref{aux}. The dependent variable is the LAD cluster membership as
described in the \protect\hyperlink{sec:3}{methods and data} section and
equation \ref{eq1}. Each column represents a different cluster. The
reference case is cluster \(4\), which includes only the local authority
of Hambleton in North Yorkshire, a rural area of just over ninety
thousand people. Mean, experienced upload speed in cluster \(4\) (see
Table \ref{up.cluster.descr}) is close to the pre-clustered average for
the whole sample (\(9.3\)Mb/s). However, the standard deviation for
cluster \(4\) and the difference between average speeds in the morning
compared to the evening peak periods are indications of worse
reliability than many of the other clusters. Hence, the results in Table
\ref{aux} should be seen as relative rather than absolute probabilities.

First, we control for the number of speed tests run per cluster
inhabitant between \(9:00\)-\(10:59\) as well as the share of fast
Virgin Media internet connections. Regarding the former, we expect
people in LADs with more unreliable connections to test their internet
speeds more often, and the results show that those in cluster \(6\) are
by far the most concerned about reliability at that time of day.
Meanwhile, those in clusters \(2\), \(3\), and \(9\) benefit from a
higher proportion of Virgin connections, which is an indication that
people in these clusters are more likely to live in urban areas, with
more profitable broadband markets, and a better choice of broadband
services. We also employ distance from London and from the nearest
metropolitan area (including London) as two variables depicting
peripherality and urban structure. However, whilst significant, the
pattern of coefficients of these variables is inconclusive in confirmed
first level digital divides. Even though London was one of the ten
largest metropolitan areas in England, which, along with Glasgow and
Cardiff, were identified to calculate the variable estimating the impact
of distance from the centre of a metropolitan area, the coefficients for
the two variables are mostly small and the signs for some are opposite.

When looking at the constituent authorities (see full list of LADs in
Appendix \protect\hyperlink{appendix1}{1} as well as Figure
\ref{map.up.clusters}), however, Cluster \(9\) is clearly more urban,
including \(13\) of \(32\) London Boroughs, five of the seven
constitutent LADs of the West Midlands conurbation, nine of the ten
boroughs of Greater Manchester, and five of nine other large
metropolitan areas coded in for the `distance to nearest met area'
variable. Cluster \(9\) also includes the other main cities in the East
Midlands, Leicester and Derby, as well as smaller cities known for their
knowledge economy, such as Oxford and Milton Keynes. However, as
previously mentioned, there is more noise within the variables for the
two largest clusters than the smaller ones, and cluster \(6\) also
contains many urban areas. These include \(16\) London boroughs,
Birmingham City, Bolton in Greater Manchester, Leeds, Liverpool,
Newcastle and Bristol, as well as small cities with knowledge economies
such as Cambridge, Edinburgh, and Reading and its neighbours in the
high-tech agglomeration of the Thames Valley. And yet cluster \(6\) also
includes some of the most rural areas in the country, showing how
important it is to interpretation to consider the membership of the
cluster as well as the regression results.

The coefficients in Table \ref{aux} demonstrate that those in cluster
\(6\) were more likely to hold managerial, tech and professional jobs
than those in cluster \(9\), and were the least likely to be furloughed
of any cluster. Thus, although speeds were slow and not as reliable
during the morning peak as in cluster \(9\), the skills that enabled
telecommuting also enabled greater returns from doing so, despite lower
than average earnings. Cluster \(6\) also has the second highest job
density, with more jobs to keep. Likewise, those in LADs in clusters
\(1\) and \(8\) had the highest and second highest proportions of
residents working in tech or managerial occupations despite suffering
from lower speeds and unreliable services, another mismatch between
first and second level digital divides. Yet unlike cluster \(6\) and
despite occupations with better digital skills, cluster \(8\), with its
five peripheral suburban areas in the Midlands, Scotland and south of
London, as well as rural West Devon saw the third highest proportion of
its working population furloughed. It also had the second lowest average
earnings in 2019, suggesting that its population was more likely to be
on the wrong side of the third level digital divide before, as well as
during the pandemic. In contrast, the nine LADs of cluster \(1\), which
include more rural areas, plus Westminster in central London had the
lowest earnings of any cluster, but also lowest level of furlough after
cluster \(6\).

The regression results for cluster \(7\) also reveal the complexity of
intersecting and diverging digital divides. Along with \(3\) and \(9\),
the analysis in Section \protect\hyperlink{sec:4.1}{4.1} suggested this
cluster was on the `right' side of any first level digital divides, yet
the regression shows that it has the lowest proportion of Virgin
connections. It is furthest from a metropolitan area, but closest to
London. The five LADs in cluster \(7\) are all suburban, making it
sensible that the cluster is in the group with better infrastructure,
but none are central urban boroughs and, along with a London suburb,
Leeds suburb, Manchester suburb, and West Midlands suburb, the fifth LAD
is outside Belfast. Belfast was not included among the metropolitan
areas, and whilst there has been substantial investment in high speed
broadband in Northern Ireland in recent years, including by Virgin,
take-up may still lag behind historic networks in other urban markets.

Cluster \(7\) has the lowest proportion of residents currently employed
in professional occupations, and among the lowest proportions in
managerial or technical occupations. This suggests potentially a low
level of skills to take advantage of the quality internet service
available, a second level digital divide. Cluster \(7\) also has the
lowest job density. However, with the largest proportion of skilled
tradespeople, and the second highest number of businesses per
inhabitant, those in cluster \(7\) ranked second in terms of earnings in
2019. On the other hand, the lack of some of the occupations most likely
to telecommute during lockdown may have contributed both to a
substantial proportion being furloughed in 2020, and even to the
relative reliability of broadband speeds compared to many of the other
smaller clusters.

Cluster \(3\), meanwhile, had the greatest proportion of its workforce
furloughed of any cluster, despite having the highest job density and
the second highest average speeds. Ranking in the middle for most of the
occupations estimated, at least some of its population should have the
skills to work from home. The geography would suggest economic advantage
as well, with eight of its twelve LADs in the South of England,
including two London Boroughs, four suburban areas north of London, and
the two tightly bounded urban areas of Eastbourne and Ipswich -- a
greater proportion than any other cluster. Yet geographic position and
job density do not appear to equate to particularly high average
earnings in 2019, nor did they protect residents from furlough in 2020.
In contrast, cluster \(5\), comprised of three suburban districts in the
Northwest, Midlands, and the London green belt, had the highest average
earnings in 2019, despite low levels of residents working in managerial,
tech and professional occupations. Suburbs are considered the most
likely urban form in which telecommuters live \citep{e2018does}, yet the
contrast between these small, mainly suburban clusters -- \(7\), \(3\),
and \(5\), suggest that they are not all equally economically resilient
in a pandemic.

Cluster \(2\), which, like cluster \(5\), had faster morning broadband
speeds than evening speeds, as shown in Table \ref{up.cluster.descr}
contains two remote LADs, the Isles of Scilly and Northeast
Lincolnshire. Whilst the low demand may be due to the least residents in
tech occupations, it may also be due to high numbers of retired people.
Rural areas such as those in cluster \(2\) are often home to many older,
retired people \citep{blank2018local}, which may also explain some of
the contradictions in cluster \(1\) with its low levels of furlough and
low earnings, whilst the presence of Westminster in that cluster might
be why there are the highest levels of tech and managerial occupations.

In summary, the regression results indicate the complexity of measuring
second and third level digital divides within spatial aggregations where
the geography of first level digital divides has been captured using
time-series clustering of experienced broadband upload speeds as a
product of reliability not just availability. Whilst the differences in
mean speeds between clusters were not large, the temporal clustering of
internet resilience showed much greater variation, and was not spatially
dependent upon distance to large urban areas, or relative location.
Internet resilience supports a wide range of small and large urban
economies, but also fails, or at least frustrates a wide range of other
urban economies, as demonstrated by the number of tests per capita run
in cluster \(6\). Yet our analysis of the likelihood of furlough within
the population, especially in the smaller suburban clusters \(7\),
\(3\), \(5\), and also \(8\) shows that digital divides were as likely
to diverge as intersect during the pandemic, and did not necessarily
overlap with prior economic divides. With a much greater share of the
population likely to continue to work from home in the future, and with
changing attitudes towards residential locations, our analysis suggests
that first level digital divides should be seen as a function of
reliability as well as availability, of upload speeds as well as
download speeds, and of greater spatial variation than might have
previously been considered.

\hypertarget{sec:5}{%
\section{Discussion and Conclusions}\label{sec:5}}

This paper offers a new perspective on telecommuting from the viewpoint
of the complex web of digital divides. We employ novel data regarding
experienced upload speeds and time-series clustering methods, a family
of unsupervised machine learning techniques which are rarely utilised in
geographical research. Fast, reliable internet connections are necessary
for the population to be able to work from home. Although not every
place hosts individuals in occupations which allow for telecommuting nor
with the necessary skills to effectively use the internet to
telecommute, this paper raises the issue that places may depend upon
good internet reliability as well as connectivity to achieve economic
resilience in a period like the current pandemic when internet
resilience is so vital.

Our analysis demonstrated that the temporal profiles of seven of our
nine clusters had slower upload speeds in the morning than in the
evening. The opposite is likely to have been the norm prior to the
pandemic, as level of demand and bandwidth management is the most common
cause of temporal variation in experienced speeds, and why evening
download speeds, rather than daytime upload speeds, have been used to
benchmark the performance of internet services. Thus, the new patterns
can be taken as evidence of widespread telecommuting and other daytime
internet use which changed the temporal profile of internet activity
throughout the UK, not just in areas with more digital industry or
better skills. Furthermore, upload speeds have not previously been seen
as integral to universal service, considering there has never before
been such extreme demand for telecommuting and operations such as video
calls. Yet those in the largest cluster, \(6\), clearly experienced
lower upload speeds. Their speeds were also less reliable, perhaps
because they had higher levels of demand than the second largest
cluster, \(9\), where many more employees were furloughed.

Almost all of the largest urban areas in the UK, as well as many smaller
cities, were split between the largest two clusters -- \(6\) and \(9\)
-- and both contain LADs that are centres of the knowledge economy, from
Cambridge in the former to Oxford in the latter, or have high
concentrations of digital businesses, like Reading in the former and
Milton Keynes in the latter \citep{technation2017}. Yet there are both
intersections and divergences of digital and economic divides. LADs in
cluster \(9\) were able to benefit both from reliable internet
connections and populations which were familiar with working from home
and could capitalise on their digital infrastructure, although overall,
they appeared to have a lower rate of return in the pandemic, with far
greater numbers furloughed. Is this an indication of both historic
economic division, as well as the third level digital divide in cities
in the North and Midlands unable to capitalise on their digital
infrastructure? Meanwhile, did the greater presence in London and the
South of LADs in cluster \(6\), as well as the inclusion of many rural
areas less impacted by the pandemic help those in this cluster be more
resilient and gain greater returns, despite using less reliable internet
services? The nuanced picture we gained through our analysis of the UK
case study suggests that in cluster \(1\) as well as \(6\), being on the
right side of the second level digital divide had a greater impact on
economic resilience and therefore the third level digital divide, than
having quality internet connectivity.

Digital infrastructure which considers upload speeds and working day
reliability as well as availability are likely to be particularly
important in a future where telecommuting might be a more common means
of accessing work and broadband services must be fit for purpose,
although the long-term effects of such drastic changes in telecommuting
and attitudes towards working from home are difficult to predict.
Nevertheless, they span various aspects of economy and society: from
changes to transportation planning due to altered commuting patterns to
changes in land use and urban planning to accommodate people who work
from home \citep{BUDNITZ2020102713, ELLDER2020102777}, and from
productivity and innovation changes to changes in agglomeration
externalities and the attraction of large cities \citep{econobs}.
Further research may be able to measure the economic resilience of the
different clusters of places discussed in this paper once this pandemic
is firmly past. However, our analysis demonstrates that the economic
resilience made possible by working from home cannot be understood
without considering the underpinning digital divides and cannot be
achieved without planning for how the levels of digital, social and
economic divides might intersect.

Early research on Covid-19 and cities speculated the changes that
potential extensive post-Covid working from home patterns might generate
for spatial structure: from more walkable cities to more localised
production and consumption patterns and from more extensive urban sprawl
to decline in public transportation and increased private car usage
\citep{batty2020editorial}. Despite the essential role it played during
the pandemic, less effort has been spent in understanding the current
and future role of digital infrastructure. Our results indicate that
probably for the first time the future of cities and spatial structure
are so much intertwined with digital infrastructure. If the post-covid
world is a world with extensive working from home, then we need to
build, among other things, resilient digital infrastructure capable of
bridging the first layer of the digital divide. Contrary to previous
broadband business and deployment plans, emphasis should be placed not
only on download, but also on upload speeds {[}\citet{brake2020lessons}.
In essence, Covid-19 accentuated the old, but more valid than ever
argument: digital infrastructure, just like any other network
infrastructure, only becomes visible when it stops working
\citep{star1999, tranos2013geography}.

\hypertarget{appendix1}{%
\section*{Appendix 1}\label{appendix1}}
\addcontentsline{toc}{section}{Appendix 1}

This is the LAD cluster membership for the upload speed timeseries.

\textbf{Cluster 1: } Ceredigion, Darlington, Eden, Erewash, Kettering,
Rossendale, Runnymede, Rutland, Westminster

\textbf{Cluster 2: } Isles of Scilly, North East Lincolnshire

\textbf{Cluster 3: } Broxbourne, Carlisle, Corby, Dundee City,
Eastbourne, Harlow, Hertsmere, Hounslow, Ipswich, Luton, Newham, South
Northamptonshire

\textbf{Cluster 4: } Hambleton

\textbf{Cluster 5: } Fylde, Rugby, Three Rivers

\textbf{Cluster 6: } Aberdeenshire, Adur, Allerdale, Amber Valley,
Angus, Antrim and Newtownabbey, Argyll and Bute, Armagh City, Banbridge
and Craigavon, Arun, Ashford, Aylesbury Vale, Babergh, Barnet,
Barrow-in-Furness, Basildon, Bassetlaw, Bath and North East Somerset,
Bedford, Belfast, Birmingham, Blackburn with Darwen, Blackpool, Blaenau
Gwent, Bolton, Boston, Bournemouth, Christchurch and Poole, Bracknell
Forest, Bradford, Braintree, Breckland, Brentwood, Bridgend, Brighton
and Hove, Bristol, City of, Broadland, Bromley, Calderdale, Cambridge,
Camden, Canterbury, Carmarthenshire, Causeway Coast and Glens, Central
Bedfordshire, Chelmsford, Cherwell, Cheshire East, Cheshire West and
Chester, Chesterfield, Chichester, Chiltern, City of Edinburgh, City of
London, Clackmannanshire, Conwy, Copeland, Cornwall, Cotswold, County
Durham, Coventry, Craven, Croydon, Dacorum, Dartford, Denbighshire,
Derbyshire Dales, Derry City and Strabane, Dorset, Dover, Dumfries and
Galloway, Ealing, East Cambridgeshire, East Devon, East Dunbartonshire,
East Hampshire, East Hertfordshire, East Lindsey, East Renfrewshire,
East Riding of Yorkshire, East Suffolk, Elmbridge, Epping Forest, Epsom
and Ewell, Exeter, Fareham, Fenland, Fermanagh and Omagh, Fife,
Flintshire, Folkestone and Hythe, Forest of Dean, Gateshead, Gloucester,
Gosport, Great Yarmouth, Greenwich, Gwynedd, Hackney, Harborough,
Haringey, Hartlepool, Hastings, Havering, Herefordshire, County of, High
Peak, Highland, Hillingdon, Horsham, Huntingdonshire, Inverclyde, Isle
of Anglesey, King's Lynn and West Norfolk, Kingston upon Hull, City of,
Kingston upon Thames, Kirklees, Lambeth, Lancaster, Leeds, Lincoln,
Liverpool, Maidstone, Malvern Hills, Mansfield, Melton, Merthyr Tydfil,
Mid Devon, Mid Suffolk, Mid Ulster, Midlothian, Mole Valley,
Monmouthshire, Moray, Na h-Eileanan Siar, Neath Port Talbot, New Forest,
Newcastle upon Tyne, Newry, Mourne and Down, North Devon, North East
Derbyshire, North Lanarkshire, North Lincolnshire, North Norfolk, North
Somerset, North Tyneside, North Warwickshire, North West Leicestershire,
Northumberland, Orkney Islands, Pembrokeshire, Pendle, Perth and
Kinross, Peterborough, Plymouth, Powys, Preston, Reading, Redcar and
Cleveland, Reigate and Banstead, Rhondda Cynon Taf, Ribble Valley,
Richmondshire, Rother, Rotherham, Rushcliffe, Rushmoor, Ryedale,
Scottish Borders, Sedgemoor, Sefton, Sevenoaks, Shetland Islands,
Shropshire, Somerset West and Taunton, South Ayrshire, South Bucks,
South Cambridgeshire, South Gloucestershire, South Hams, South Holland,
South Lakeland, South Norfolk, South Oxfordshire, South Ribble, South
Somerset, South Staffordshire, Southend-on-Sea, Southwark, St Albans,
Stafford, Stirling, Stoke-on-Trent, Stratford-on-Avon, Stroud, Swale,
Swansea, Swindon, Teignbridge, Tendring, Test Valley, Tewkesbury,
Thanet, Tonbridge and Malling, Torbay, Torfaen, Torridge, Tower Hamlets,
Tunbridge Wells, Uttlesford, Wakefield, Waltham Forest, Wandsworth,
Warrington, Watford, Waverley, Wellingborough, West Berkshire, West
Lancashire, West Lindsey, West Oxfordshire, Wiltshire, Winchester,
Windsor and Maidenhead, Wirral, Wokingham, Worcester, Worthing, Wrexham,
Wychavon, Wycombe, Wyre

\textbf{Cluster 7: } Lisburn and Castlereagh, Selby, Tameside, Welwyn
Hatfield, Wyre Forest

\textbf{Cluster 8: } Cannock Chase, East Lothian, Lichfield, Oadby and
Wigston, Tandridge, West Devon

\textbf{Cluster 9: } Aberdeen City, Ards and North Down, Ashfield,
Barking and Dagenham, Barnsley, Basingstoke and Deane, Bexley, Blaby,
Bolsover, Brent, Bromsgrove, Broxtowe, Burnley, Bury, Caerphilly,
Cardiff, Castle Point, Charnwood, Cheltenham, Chorley, Colchester,
Crawley, Daventry, Derby, Doncaster, Dudley, East Ayrshire, East
Northamptonshire, East Staffordshire, Eastleigh, Enfield, Falkirk,
Gedling, Glasgow City, Gravesham, Guildford, Halton, Hammersmith and
Fulham, Harrogate, Harrow, Hart, Havant, Hinckley and Bosworth,
Hyndburn, Isle of Wight, Islington, Kensington and Chelsea, Knowsley,
Leicester, Lewes, Lewisham, Maldon, Manchester, Medway, Mendip, Merton,
Mid and East Antrim, Mid Sussex, Middlesbrough, Milton Keynes, Newark
and Sherwood, Newcastle-under-Lyme, Newport, North Ayrshire, North
Hertfordshire, North Kesteven, Northampton, Norwich, Nottingham,
Nuneaton and Bedworth, Oldham, Oxford, Portsmouth, Redbridge, Redditch,
Renfrewshire, Richmond upon Thames, Rochdale, Rochford, Salford,
Sandwell, Scarborough, Sheffield, Slough, Solihull, South Derbyshire,
South Kesteven, South Lanarkshire, South Tyneside, Southampton,
Spelthorne, St.~Helens, Staffordshire Moorlands, Stevenage, Stockport,
Stockton-on-Tees, Sunderland, Surrey Heath, Sutton, Tamworth, Telford
and Wrekin, Thurrock, Trafford, Vale of Glamorgan, Vale of White Horse,
Walsall, Warwick, Wealden, West Dunbartonshire, West Lothian, West
Suffolk, Wigan, Woking, Wolverhampton, York

\hypertarget{appendix2}{%
\section*{Appendix 2}\label{appendix2}}
\addcontentsline{toc}{section}{Appendix 2}

\begin{table}[!htbp] \centering 
  \caption{Descriptive statistics for the auxiliary regression explanatory variables\label{descr.aux}} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{0pt}}lccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Pctl(25)} & \multicolumn{1}{c}{Pctl(75)} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
pop, 2018 & 365 & 174,952.100 & 119,557.100 & 8,700.000 & 100,400.000 & 214,900.000 & 1,141,400.000 \\ 
job density, 2018 & 365 & 1.137 & 5.726 & 0.400 & 0.700 & 0.930 & 110.110 \\ 
distance to nearest met. area & 365 & 53.269 & 57.700 & 0.150 & 22.050 & 69.290 & 544.090 \\ 
distance to London & 365 & 201.558 & 173.634 & 0.150 & 76.180 & 278.880 & 1,003.950 \\ 
south of the UK & 365 & 0.463 & 0.499 & 0.000 & 0.000 & 1.000 & 1.000 \\ 
managerial jobs, 2020 & 363 & 12.009 & 4.013 & 3.600 & 9.000 & 14.300 & 27.900 \\ 
tech jobs, 2020 & 364 & 14.505 & 4.057 & 3.500 & 11.800 & 16.900 & 29.600 \\ 
skilled trade jobs, 2020 & 358 & 10.513 & 3.764 & 1.000 & 8.025 & 12.500 & 21.600 \\ 
professional jobs, 2020 & 364 & 21.223 & 6.902 & 4.400 & 16.775 & 24.850 & 71.600 \\ 
administrative jobs, 2020 & 359 & 9.965 & 2.738 & 3.200 & 8.100 & 11.400 & 21.300 \\ 
service jobs, 2020 & 362 & 9.261 & 2.827 & 2.800 & 7.300 & 11.400 & 17.800 \\ 
machine operation jobs, 2020 & 337 & 6.339 & 2.847 & 1.200 & 4.400 & 7.900 & 19.800 \\ 
earnings, 2019 & 360 & 592.184 & 81.129 & 437.600 & 534.625 & 633.875 & 893.200 \\ 
Virgin Media \%, 2020 & 381 & 0.150 & 0.140 & 0.000 & 0.018 & 0.238 & 0.753 \\ 
n. business est. per hab., 2019 & 365 & 0.057 & 0.164 & 0.023 & 0.038 & 0.056 & 3.174 \\ 
AM tests per hab., 2020 & 365 & 0.0005 & 0.0002 & 0.0001 & 0.0003 & 0.001 & 0.001 \\ 
furloughed per hab., 2020 & 362 & 0.115 & 0.013 & 0.077 & 0.107 & 0.122 & 0.178 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

\bibliographystyle{tfcad}
\bibliography{bibliography.bib}

\end{document}