### 2021-03-25 src/descriptive_Clusters.Rmd
leftovers -- sorry

---
title: "descriptive_Clusters"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r settings, include = FALSE}
library(tsbox)
library(ggplot2)
library(tidyverse)
library(geojsonio)
library(tsfeatures)
library(rprojroot)
library(openair)
library(reshape2)
library(lubridate)
library(rgdal)
library(maptools)
library(sp)
library(rgeos)
library(classInt)
library(RColorBrewer)
library(knitr)
library(rprojroot)
library(viridis)
library(hrbrthemes)

# This is the project path
path <- find_rstudio_root_file()
```

## Time variation plots (Figures 1 and 2)

```{r include = FALSE}

# load the data created by Data_Spatial.Rmd and ts_clusters.Rmd

LAD.cluster <- paste(path, "/data/temp/clusters_nodiff_9.csv", sep = "")

clusters <- read_csv(LAD.cluster)

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")

TSbb19_20sp <- read_csv(path.data)

#split out 2020 before join to clusters to create descriptives
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) 
TS2019 <- selectByDate(TSbb19_20sp, year = 2019)

#create time variation plots for speedtests for each year
Tests19 <- count(TS2019, date)
TS2019 <- left_join(TS2019, Tests19)
names(TS2019)[13] <- "test frequency 2019"
time.var.plot2019 <- timeVariation(TS2019, pollutant = "test frequency 2019", statistic = "mean",
                                   par.settings=list(fontsize=list(text=8)))
Tests20 <- count(TS2020, date)
TS2020 <- left_join(TS2020, Tests20)
names(TS2020)[13] <- "test frequency 2020"
time.var.plot2020 <- timeVariation(TS2020, pollutant = "test frequency 2020", statistic = "mean",
                                   par.settings=list(fontsize=list(text=8)))

#export time variation plots for paper
out.plot <- paste(path, "/paper/v2_taylor_francis/figures/time.var.plot2019.png", sep = "")
png(out.plot, res = 100, width = 615, height = 615) # 72dpi default res. 480x480. 
                                                    # ((100-72)/100)*480+480 = 615 
time.var.plot2019
dev.off() # if dev.off() doesn't work try dev.set(dev.next()) 2-3 times

out.plot <- paste(path, "/paper/v2_taylor_francis/figures/time.var.plot2020.png", sep = "")
png(out.plot, res = 100, width = 615, height = 615)
time.var.plot2020
dev.off()
```

```{r}
#join clusters to TS2020
names(clusters) [1] <- "LAD19NM"
TS2020 <- left_join(TS2020, clusters)
TS2019 <- left_join(TS2019, clusters)
#remove NAs? - otherwise cannot create unified dataframes for below stats

#add LAD population data
pop <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_31_1.data.csv?geography=1820327937...1820328318,1816133633...1816133848&date=latestMINUS1-latest&sex=7&age=0,22&measures=20100")

pop <- pop %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, AGE_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_NAME, AGE_NAME, .keep_all = T) %>% 
  spread(AGE_NAME, OBS_VALUE) %>% 
  rename(pop16_64 = `Aged 16 - 64`) %>%
  rename(pop = `All ages`) %>%
  rename(LAD19NM = GEOGRAPHY_NAME)

#join to clusters
pop$LAD19NM <- if_else(pop$LAD19NM == "Rhondda Cynon Taff", "Rhondda Cynon Taf", pop$LAD19NM)
clusters <- left_join(clusters, pop)

#summarise by LADs per cluster and Total tests / Mean speed per cluster
#Test Frequency cluster
LAD.cluster.tests <- clusters %>%
  group_by(cluster.tests) %>%
  summarise(n.LADs = n())
LAD.cluster.pop <- clusters %>%
  group_by(cluster.tests) %>%
  summarise(pop.LADs = sum(pop))
LAD.cluster.testsTotal <- TS2020 %>%
  group_by(cluster.tests) %>%
  summarise(n.cluster = n()) 
LAD.cluster.tests19 <- TS2019 %>%
  group_by(cluster.tests) %>%
  summarise(n.cluster19 = n()) 
LAD.cluster.tests <- cbind(LAD.cluster.tests, LAD.cluster.pop [2], LAD.cluster.testsTotal[,2], LAD.cluster.tests19[,2])
LAD.cluster.tests$TestsPop20 <- LAD.cluster.tests$n.cluster / LAD.cluster.tests$pop.LADs
LAD.cluster.tests$TestsPop19 <- LAD.cluster.tests$n.cluster19 / LAD.cluster.tests$pop.LADs
LAD.cluster.tests$n.20_19.diff <- LAD.cluster.tests$n.cluster - LAD.cluster.tests$n.cluster19
LAD.cluster.tests$perCent.change <- LAD.cluster.tests$n.20_19.diff/
  LAD.cluster.tests$n.cluster19
LAD.cluster.tests$perCent.change <- paste(round(100*LAD.cluster.tests$perCent.change, 2), "%", sep="")
#Upload Speed Clusters
LAD.cluster.up <- clusters %>%
  group_by(cluster.up) %>%
  summarise(up.LADs = n())
LAD.cluster.upPop <- clusters %>%
  group_by(cluster.up) %>%
  summarise(pop.LADs = sum(pop))
LAD.cluster.upMean <- summarise(group_by(TS2020, cluster.up),
  mean(speedup))
LAD.cluster.upSD <- summarise(group_by(TS2020, cluster.up),
  sd(speedup))
LAD.cluster.upAM <- summarise(group_by(TS2020[which(TS2020$hour == 9 | TS2020$hour == 10),], cluster.up), mean(speedup))
LAD.cluster.upPM <- summarise(group_by(TS2020[which(TS2020$hour == 19 | TS2020$hour == 20),], cluster.up), mean(speedup))
LAD.cluster.up <- cbind(LAD.cluster.up, LAD.cluster.upPop[,2], LAD.cluster.upMean[,2], LAD.cluster.upSD[,2], LAD.cluster.upAM[,2], LAD.cluster.upPM[,2])
names(LAD.cluster.up) [4:7] <- c("upMean", "upSD", "upAMmean",
                                 "upPMmean")
LAD.cluster.up$AM_slower <- (LAD.cluster.up$upPMmean - LAD.cluster.up$upAMmean) / LAD.cluster.up$upPMmean
LAD.cluster.up$SDtoMean <- LAD.cluster.up$upSD/LAD.cluster.up$upMean
LAD.cluster.up$AM_slower <- paste(round(100*LAD.cluster.up$AM_slower, 2), "%", sep="")
LAD.cluster.up$SDtoMean <- paste(round(100*LAD.cluster.up$SDtoMean, 2), "%", sep="")
#Download speed clusters  
LAD.cluster.down <- clusters %>%
  group_by(cluster.down) %>%
  summarise(down.LADs = n())
LAD.cluster.downPop <- clusters %>%
  group_by(cluster.down) %>%
  summarise(pop.LADs = sum(pop))
LAD.cluster.downMean <- summarise(group_by(TS2020, cluster.down),
  mean(speeddown))
LAD.cluster.downSD <- summarise(group_by(TS2020, cluster.down),
  sd(speeddown))
LAD.cluster.downAM <- summarise(group_by(TS2020[which(TS2020$hour == 9 | TS2020$hour == 10),], cluster.down), mean(speeddown))
LAD.cluster.downPM <- summarise(group_by(TS2020[which(TS2020$hour == 19 | TS2020$hour == 20),], cluster.down), mean(speeddown))
LAD.cluster.down <- cbind(LAD.cluster.down, LAD.cluster.downPop[,2], LAD.cluster.downMean[,2], LAD.cluster.downSD[,2], LAD.cluster.downAM[,2], LAD.cluster.downPM[,2])
names(LAD.cluster.down)[4:7] <- c("downMean", "downSD", "downAMmean",
                                  "downPMmean")
LAD.cluster.down$AM_slower <- (LAD.cluster.down$downPMmean - LAD.cluster.down$downAMmean) / LAD.cluster.down$downPMmean
LAD.cluster.down$SDtoMean <- LAD.cluster.down$downSD / LAD.cluster.down$downMean
LAD.cluster.down$AM_slower <- paste(round(100 * LAD.cluster.down$AM_slower, 2), "%", sep="")
LAD.cluster.down$SDtoMean <- paste(round(100*LAD.cluster.down$SDtoMean, 2), "%", sep="")
# commented to run the script
# LAD.cluster.diffUp <- clusters %>%
#   group_by(cluster.up.diff) %>%
#   summarise(diffUp.LADs = n())
# LAD.cluster.diffDown <- clusters %>%
#   group_by(cluster.down.diff) %>%
#   summarise(diffDown.LADs = n())

#write csvs to outputs
out.clustersLAD <- paste(path, "./data/temp/LAD.cluster.tests.csv", sep = "")
write_csv(LAD.cluster.tests, out.clustersLAD)
out.clustersLADup <- paste(path, "./data/temp/LAD.cluster.up.csv", sep = "")
write_csv(LAD.cluster.up, out.clustersLADup)
out.clustersLADdown <- paste(path, "./data/temp/LAD.cluster.down.csv", sep = "")
write_csv(LAD.cluster.down, out.clustersLADdown)

```

As people came to rely more on their broadband services, it is unsurprising that both the amount and rate of testing increased in 2020 during the lockdown compared to the same period in 2019. This increase was greatest in cluster 2, where 56% more tests were run compared to the same period in 2019. Cluster 2 also has the most local authorities, 148, which are home to more than 28.5 million people, so the sum of tests run in all these authorities is much higher than in all the other clusters, both in 2019 and 2020. Cluster 1 in comparison, not only has has the lowest  frequency of testing per capita in both 2019 and 2020, it also shows the lowest rate of increase in testing between the two time periods at just 33%.

Meanwhile, cluster 5 has the fewest local authorities, at 28, but the second highest total number of tests in both periods and the second largest increase in testing between 2019 and 2020, at 52%. It also has the highest number of tests per capita in both periods. A possible explanation is that cluster 5, with a population of just under 12 million contains some of the most populous local authorities in the country, including the cities of Birmingham, Manchester, Leeds, Sheffield, Glasgow and Edinburgh. However, this cluster also includes large rural counties, such as Cornwall and Northumberland. Are urban dwellers or rural households more likely to test their broadband speeds? Auxiliary regressions suggest that cluster 5 also had the highest earnings of the clusters, but was least likely to have a Virgin Media connection.

Yet the temporal profile in 2020 is very different to 2019, as can be seen in graphs [.] and [.]. The number of broadband speed tests run in local authorities in cluster 2 peaked each Monday to Thursday in the hour 19.00-19.59 in March through May 2019. On Fridays, the peak was slightly earlier. A similar, although less stark pattern could be seen in the other clusters, along with much smaller peaks around the beginning of the working day, at either 9.00-9.59 or 10.00-10.59. In comparison, the daily profile of broadband speed testing in 2020 is flatter, with less variation hour to hour. Although there is still an identifiable evening peak for almost all days of the week and all clusters, it is much less pronounced throughout, there is less of a gap between the number of tests in the morning peak and the evening peak, and there is less of a dip in testing in the middle of the day. All this suggests that more of the additional testing generated in 2020 compared to 2019 is occurring outside the traditional peak in the evening at streaming primetime.

The increase in daytime compared to evening testing is particularly noticeable for cluster 2, but the change is less pronounced for cluster 5. Indeed, in an auxiliary regression, the probability of additional tests per person run between 9.00 and 11.00 was significantly higher for those living in cluster 2 than in cluster 5, but both were higher than the other three clusters. ?? perhaps because of the diversity of urban and rural populations represented in cluster 5 ?? Cluster 5 mostly same as cluster 6 upload, few cluster 9, but same for cluster 2?
#re-base auxiliary regression so reference case is either 1 or 3 - we wanted to talk most about 5?
The graphs also show that Cluster 1, with the lowest increase in testing, falls from the middle position in terms of total number of tests to the second from bottom.

For upload speeds, 345 of 382 local authorities, or over 62 million people, fall into cluster 6 or cluster 9. Graph [.] shows that both of these clusters have relatively similar temporal profiles, with relatively stable speeds. However, the upload speeds at all times for cluster 9 are substantially higher than for cluster 6. This difference is reflected not only in the mean speeds for these clusters for the whole sample, but also the mean upload speeds in the morning peak from 9:00-10:59, as well as the evening peak period from 19:00-20:59. The quality of broadband service is thus much better in the 115 Local Authority areas in cluster 9, which are mostly in urban or suburban areas. These include 13 London Boroughs (of 32), 8 of the 10 local authorities of Greater Manchester, 5 of the 7 constituent authorities of the West Midlands Combined Authority, as well as cities like Glasgow, Leicester, Nottingham, Sheffield, and the Portsmouth and Southampton conurbation. There are also some notable medium-sized cities, including Aberdeen, Cardiff, Oxford, Milton Keynes, and York, and many suburban districts from the South East of England to South Tyneside. Auxiliary regressions indicate that the speed tests in these authorities are also more likely to have been run on services provided by Virgin Media, suggesting they are in the half of the country with the most lucrative ICT market, which originally attracted the cable TV provider (OfCom...).

Meanwhile, the 230 local authorities in cluster 6 still include major urban areas, such as Bristol, Liverpool and Leeds, and many suburban areas, but also include some of the most rural areas in the country. For example, although auxiliary regressions show that Cluster 6 local authorities are more likely to be in the South of the UK than Cluster 9, the cluster notably includes Southern rural districts from Cornwall to North Norfolk. Slower speeds could reflect the lower quality of service in rural areas compared to urban and suburban areas. The reliability of service in these areas is also in question. For although upload speeds in cluster 6 are only 4% slower in the morning than in the evening, compared to 12-19% slower for the almost 3 million people living in five of the smaller clusters, auxiliary regressions show that those living in cluster 6 had the highest probability of testing their broadband between 9:00 and 11:00 of any of the 10 upload speed clusters. Cluster 6 also experienced the joint highest ratio of standard deviation to mean across the time period under assessment. This suggests that although the time profile is relatively flat in graph [.], the experience is one of speeds that fluctuate from a lower mean, and therefore might more often impact on the activities users are doing.

In comparison, there is negligible difference between morning and evening upload speeds for cluster 9, at 1% slower in the morning. This suggests a high level of reliability of service, a conclusion supported by the finding in the auxiliary regressions that those in cluster 9 ran the fewest speed tests per person during the morning period of any cluster. Now this may be an indication of fewer people working from home, less contention, and less resultant frustration. Cluster 9 comprises many central urban areas and has the lowest number of established businesses per inhabitant, which could be interpreted as a dominance of large employers. However, the job density is lower than cluster 6, meaning there are not as many jobs per resident in these areas. As the cluster includes many suburban areas too, which may be largely residential, could the quality and reliability of internet service be reinforcing patterns of telecommuting by those in wealthier suburbs who can work from home? Earnings in cluster 9 are second highest of all the clusters, with only Cluster 2 (comprising just North East Lincolnshire and East Lothian, population 265k) earning more per person.

Meanwhile, Cluster 1, with 10 local authorities that are home to over 1 million people has the second slowest speeds in the morning compared to the evening 'peaks' and the second highest ratio of standard deviation to mean. Indeed, the time profile in graph [.] shows upload speeds in cluster 1 are on average lower at certain times of day during the study period than any other cluster. Yet the auxiliary regression suggests that there are not many tests being run during the am peak in cluster 1. This may be because there are fewer people working at home checking their broadband than in most other clusters. Indeed, the auxiliary regressions indicate that cluster 1 has the highest job density or proportion of jobs to working-age population, which is likely to due to the presence of Westminster, central London, cluster 1's most populous local authority. Westminster not only has more workplaces than residents, but it is reasonable to presume that many who would normally work in Westminster, but be able to work from home during lockdown are likely to live outside central London and not be subject to the fluctuating speeds there. Workplaces, meanwhile, some of which would still have been open, could be running programmes that cause the slowdown and variation, but would be more likely to have their own in-house diagnostics, rather than using a service like Speedchecker Ltd.

Indeed, the only cluster where upload speeds were slower in the evening than in the morning was cluster 5, made up of 2 local authorities with less than 200,000 people: Three Rivers, a suburban district north of London, and Fylde, a seaside suburb of Blackpool. However, these are likely to be outliers and may not have many tests from which the clusters are calcualted. **Take out smaller clusters!?** The auxiliary regressions show that compared to these two places, all the other clusters had a lower percentage of working people in managerial, professional and administrative jobs.???

This is an indication of the reliability **or maybe not - 5% slowdown from low base can be noticeable, and sd of cluster 6 bigger** of broadband services in places in both clusters, and suggests that upload speeds have generally been robust throughout the study period. The exceptions can be found in seven of the eight much smaller clusters including 35 local authority districts, where AM peak upload speeds are between 6% and 18% slower than PM peak upload speeds, although the mean speeds for each cluster are higher than cluster 6. The temporal profiles on Graph [.] for the three of those clusters with more than 5 local authorities also shows that speeds are far from dependable in these clusters. Indeed, in 25 local authorities with a combined population of almost 3 million, speeds are 13% or more slower in the morning than in the evening. Included in this latter group are central London borough of Westminster and the London Borough of Newham, rural authorities like Eden and West Devon, and small cities like Dundee and Carlisle.

Download speeds are the most commonly used measure of the quality and reliability of broadband services. They are most sensitive, which does result in substantial noise. This is clear from the way that download speeds vary the most, with the largest range and standard deviations in speed tending to be in a similar range to the mean speeds for each cluster, compared to upload speeds, where standard deviations are half to two thirds the level of the mean speeds. This temporal variation can be seen in Graph [.], which shows the 10 of 20 clusters with over five local authorities in them. The second largest of these 10 clusters, number 19 with 123 Local Authorities and a population of almost 18 million, along with cluster 1 with 16 local authorities and an additional 1.5million people have lower mean speeds than the other clusters. This is consistent throughout the temporal profile on the graph. However, cluster 19 also has the second lowest standard deviation at 88% of the mean speed, whilst the standard deviation of cluster 1 is 110% and thus greater than the mean speed.

What is of greater interest, however, is where download speeds are slower in the morning peak when broadband is more likely to be used for work purposes than in the evening peak. This difference is most stark in some of the smaller clusters, although since there are only 11 Local Authorities and a little less than 1.2 million people in these clusters altogether, the differences in speeds may be a reflection of quite localised issues. The largest two clusters of Local Authorities, which are home to over 39 million people, have speeds only 1-3% slower in the AM peak, and a few clusters have faster download speeds in the AM peak. However, there are 4 clusters with over 9 million people in 47 Local Authorities with download speeds 5-8% lower in the morning peak than in the evening peak, a noticeable reduction and one equivalent to the contention monitored in the evening peak by OfCom to measure service performance. Is the 5% or greater reduction in download speeds in the morning peak compared to the evening peak suffered in more places by more people than suffer a similar reduction in upload speeds a result of ISP management to support reliable services for telecommuters? There are only eight Local Authorities where both upload and download speeds are more than 5% slower in the AM peak than the PM peak.

## Summarise 2020 test frequency clusters by hour of each weekday and plot

```{r }
cluster.tests <- TS2020 %>%
  group_by(cluster.tests, weekday, hour) %>%
  summarise(n.tests = n())
names(cluster.tests)[1] <- "clusterID"
cluster.tests$clusterID <- as.factor(cluster.tests$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.tests <- merge(cluster.tests, help.dates, by = "weekday")
cluster.tests$hour <- paste(cluster.tests$hour, ":00", sep="")
cluster.tests$help.date.time <- with(cluster.tests, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for number tests
TestCluster <- ggplot(cluster.tests, aes(help.date.time, n.tests, 
                                         colour = clusterID))
TestCluster + geom_line() + labs(x = "Weekdays March-May 2020") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + geom_point()
#summarise test frequency clusters by hour of each weekday 2019 and plot
cluster.tests19 <- TS2019 %>%
  group_by(cluster.tests, weekday, hour) %>%
  summarise(n.tests = n())
names(cluster.tests19)[1] <- "clusterID"
cluster.tests19$clusterID <- as.factor(cluster.tests19$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2019, month = 3, day = 4),
  make_datetime(year = 2019, month = 3, day = 5),
  make_datetime(year = 2019, month = 3, day = 6),
  make_datetime(year = 2019, month = 3, day = 7),
  make_datetime(year = 2019, month = 3, day = 8)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.tests19 <- merge(cluster.tests19, help.dates, by = "weekday")
cluster.tests19$hour <- paste(cluster.tests19$hour, ":00", sep="")
cluster.tests19$help.date.time <- with(cluster.tests19, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for number tests
TestCluster19 <- ggplot(cluster.tests19, aes(help.date.time, n.tests, 
                                         colour = clusterID))
TestCluster19 + geom_line() + labs(x = "Weekdays March-May 2019") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + geom_point()
```

## Summarise 2020 upload speed clusters by hour of each weekday and plot

```{r}

# plot largest 6 clusters - all over 500k pop
cluster.up <- TS2020 %>%
  group_by(cluster.up, weekday, hour) %>%
  summarise(mean.up = mean(speedup))
names(cluster.up)[1] <- "clusterID"
names(LAD.cluster.up)[1] <- "clusterID"
cluster.up <- left_join(cluster.up, LAD.cluster.up)
cluster.up <- cluster.up[which(cluster.up$up.LADs > 4),]
cluster.up$clusterID <- as.factor(cluster.up$clusterID)

# add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.up <- merge(cluster.up, help.dates, by = "weekday")
cluster.up$hour <- paste(cluster.up$hour, ":00", sep="")
cluster.up$help.date.time <- with(cluster.up, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

# plot cluster statistic for mean upload speed
UpCluster <- ggplot(cluster.up, aes(help.date.time, mean.up, 
                                      colour = clusterID)) +
  theme_minimal() +
  geom_line() + 
  labs(x = "Weekdays March-May", y = "mean upload speed", col = "Cluster ID") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + 
  geom_point() +
  theme(legend.position = c(0.7, 0.08),
        legend.direction = "horizontal")

#export UpCluster plot for paper
out.plot <- paste(path, "/paper/v2_taylor_francis/figures/UpCluster.png", sep = "")
UpCluster
ggsave(out.plot)

# EXCLUDED from paper
# plot next 6 smaller clusters, but exclude 2 clusters with 1 LAD and under 100k pop
cluster.upS <- TS2020 %>%
  group_by(cluster.up, weekday, hour) %>%
  summarise(mean.up = mean(speedup))
names(cluster.upS)[1] <- "clusterID"
names(LAD.cluster.up)[1] <- "clusterID"
cluster.upS <- left_join(cluster.upS, LAD.cluster.up)
cluster.upS <- cluster.upS[which(cluster.upS$up.LADs < 10 & cluster.upS$up.LADs > 1),]
cluster.upS$clusterID <- as.factor(cluster.upS$clusterID)

# add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.upS <- merge(cluster.upS, help.dates, by = "weekday")
cluster.upS$hour <- paste(cluster.upS$hour, ":00", sep="")
cluster.upS$help.date.time <- with(cluster.upS, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

# plot cluster statistic for mean upload speed
UpClusterS <- ggplot(cluster.upS, aes(help.date.time, mean.up, 
                                      colour = clusterID)) +
  theme_minimal() +
  geom_line() + 
  labs(x = "Weekdays March-May", y = "mean upload speed", col = "Cluster ID") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + 
  geom_point() +
  theme(legend.position = c(0.8, 0.08),
        legend.direction = "horizontal")

# The figure with the small clusters is not included in the paper
#export UpClusterS plot for paper
#out.plot <- paste(path, "/paper/v2_taylor_francis/figures/UpClusterS.png", sep = "")
#png(out.plot)
#UpClusterS
#dev.off() 
```

## Summarise 2019 upload speed clusters by hour of each weekday and plot

This is a test to compare 2020 vs. 2019 line graphs.

```{r}

# plot largest 6 clusters - all over 500k pop
cluster.up <- TS2019 %>%
  group_by(cluster.up, weekday, hour) %>%
  summarise(mean.up = mean(speedup))
names(cluster.up)[1] <- "clusterID"
names(LAD.cluster.up)[1] <- "clusterID"
cluster.up <- left_join(cluster.up, LAD.cluster.up)
cluster.up <- cluster.up[which(cluster.up$up.LADs > 4),]
cluster.up$clusterID <- as.factor(cluster.up$clusterID)

# add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2019, month = 3, day = 4),
  make_datetime(year = 2019, month = 3, day = 5),
  make_datetime(year = 2019, month = 3, day = 6),
  make_datetime(year = 2019, month = 3, day = 7),
  make_datetime(year = 2019, month = 3, day = 8)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.up <- merge(cluster.up, help.dates, by = "weekday")
cluster.up$hour <- paste(cluster.up$hour, ":00", sep="")
cluster.up$help.date.time <- with(cluster.up, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))

# plot cluster statistic for mean upload speed
UpCluster2019 <- ggplot(cluster.up, aes(help.date.time, mean.up, 
                                      colour = clusterID)) +
  theme_minimal() +
  geom_line() + 
  labs(x = "Weekdays March-May", y = "mean upload speed", col = "Cluster ID") + 
  scale_x_datetime(breaks = as.POSIXct(c("2019-03-04 15:00:00", 
                    "2019-03-05 15:00:00", "2019-03-06 15:00:00",
                    "2019-03-07 15:00:00", "2019-03-08 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + 
  geom_point() +
  theme(legend.position = c(0.7, 0.08),
        legend.direction = "horizontal")

#export UpCluster plot for paper
out.plot <- paste(path, "/paper/v2_taylor_francis/figures/UpCluster2019.png", sep = "")
UpCluster2019
ggsave(out.plot)
```

```{r eval=F}
# Not used in the paper

#summarise download speed clusters by hour of each weekday and plot
cluster.down <- TS2020 %>%
  group_by(cluster.down, weekday, hour) %>%
  summarise(mean.down = mean(speeddown))
names(cluster.down)[1] <- "clusterID"
names(LAD.cluster.down)[1] <- "clusterID"
cluster.down <- left_join(cluster.down, LAD.cluster.down)
cluster.down <- cluster.down[which(cluster.down$down.LADs > 5),]
cluster.down$clusterID <- as.factor(cluster.down$clusterID)
#add fake dates for plotting
help.dates <- tibble(help.date.time = c(
  make_datetime(year = 2020, month = 3, day = 2),
  make_datetime(year = 2020, month = 3, day = 3),
  make_datetime(year = 2020, month = 3, day = 4),
  make_datetime(year = 2020, month = 3, day = 5),
  make_datetime(year = 2020, month = 3, day = 6)),
                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"))
cluster.down <- merge(cluster.down, help.dates, by = "weekday")
cluster.down$hour <- paste(cluster.down$hour, ":00", sep="")
cluster.down$help.date.time <- with(cluster.down, as.POSIXct(paste(help.date.time, hour), format="%Y-%m-%d %H:%M"))
#plot cluster statistic for mean download speeds
TestCluster <- ggplot(cluster.down, aes(help.date.time, mean.down, 
                                         colour = clusterID))
TestCluster + geom_line() + labs(x = "Weekdays March-May") + 
  scale_x_datetime(breaks = as.POSIXct(c("2020-03-02 15:00:00", 
                    "2020-03-03 15:00:00", "2020-03-04 15:00:00",
                    "2020-03-05 15:00:00", "2020-03-06 15:00:00"), 
                    labels = "%a"), labels = 
                     c("Mon", "Tues", "Weds", "Thurs", "Fri")) + geom_point()

```