### 2021-05-11 src/online_appendix.Rmd
word version for final submission

---
title: "Supplemental Material"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: menu
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/v2_taylor_francis/")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(plotly)
library(lubridate)
library(gridExtra)
library(openair)
library(rgdal)
library(maps)
library(rprojroot)
library(kableExtra)
#library(cowplot)
library(ggthemes)

path <- find_rstudio_root_file()
```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
# load the data created by Data_Spatial.Rmd

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")

TSbb19_20sp <- read_csv(path.data)

#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs

# save clusters file
path.clusters <- paste(path, "/data/temp/clusters_nodiff_9.csv", sep = "")
clusters.nodiff <- read_csv(path.clusters)
```

Column 
-----------------------------------------------------------------------

### LAD, Upload speed clusters, 2020

```{r, echo=FALSE, warning=FALSE, message = FALSE, error=FALSE, fig.height=10, fig.asp=T}
#out.width='100%'

# LAD fully generalised
la <- readOGR("https://opendata.arcgis.com/datasets/3a4fa2ce68f642e399b4de07643eeed3_0.geojson",
              verbose = FALSE)
# saved locally in /data/raw/Local_Authority_Districts_(December_2019)_Boundaries_UK_BUC.geojson

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))
#la@data$LAD19NM <- as.character(la@data$LAD19NM)
la@data$lad19nm <- as.character(la@data$lad19nm)

# upload metrics
for.tip.mean.up.am <- TS2020 %>% # raw data, individual tests
  filter(hour(date)==9 | hour(date)==10) %>% 
  group_by(LAD19NM) %>%
  summarise(mean.up.am = mean(speedup)) %>% 
  mutate(rank = dense_rank(desc(mean.up.am)))

for.tip.mean.up.pm <- TS2020 %>% # raw data, individual tests
  filter(hour(date)==19 | hour(date)==20) %>%
  group_by(LAD19NM) %>%
  summarise(mean.up.pm = mean(speedup))

# not used
# for.tip.mean.up <- TS2020 %>% # raw data, individual tests
#   group_by(LAD19NM) %>%
#   summarise(mean.up = mean(speedup)) 
# 
# # download metrics
# for.tip.mean.down.am <- TS2020 %>% # raw data, individual tests
#   filter(hour(date)==9 | hour(date)==10) %>% 
#   group_by(LAD19NM) %>%
#   summarise(mean.down.am = mean(speeddown)) 
# 
# for.tip.mean.up.pm <- TS2020 %>% # raw data, individual tests
#   filter(hour(date)==19 | hour(date)==20) %>% 
#   group_by(LAD19NM) %>%
#   summarise(mean.down.pm = mean(speeddown)) 
# 
# for.tip.mean.up <- TS2020 %>% # raw data, individual tests
#   group_by(LAD19NM) %>%
#   summarise(mean.down = mean(speeddown))

# fortify
la.f <- broom::tidy(la, region = "lad19nm")
la.f <- merge(la.f, clusters.nodiff, by.x = "id", by.y = "LAD")

# create the text to tooltip
la.f <- merge(la.f, for.tip.mean.up.am, by.x = "id", by.y = "LAD19NM", all.x=T)
la.f <- merge(la.f, for.tip.mean.up.pm, by.x = "id", by.y = "LAD19NM", all.x=T)

#sapply(la.f, function(x) sum(is.na(x)))
#No am data points for Isles of Scilly
la.f$tip <- paste0(la.f$id, "<br>",
                   "Mean upload AM speed: ", round(la.f$mean.up.am), " Kbps", "<br>",
                   "Mean upload PM speed: ", round(la.f$mean.up.pm), " Kbps")
la.f <- la.f[order(la.f$order),] # if i don't order merge.nuts.f loses order and the map has gaps

# using world.cities from `maps`
cities <- world.cities %>% filter(country.etc=="UK")

p <- ggplot(la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = as.factor(cluster.up), 
                   text = tip)) +# ,text =paste("test1", long, "test2", lat)
  #labs(title = "Upload clusters") +
  scale_fill_brewer(palette="Set1", 
                    direction = -1) +
  #geom_text_repel( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat, label=name), size=4) +
  geom_point( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat), color="black", size=0.5) +
  geom_text( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat, label=name), size=3) +

  guides(fill=guide_legend(title="Cluster")) +
  theme_map()
#   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
# panel.grid.minor = element_blank())

p %>%
  ggplotly(tooltip = "text") %>%
  style(hoverlabel = list(bgcolor = "white"), hoveron = "fill")
```

Column
-----------------------------------------------------------------------

### LAD mean AM and PM upload speeds

```{r, echo=FALSE, warning=FALSE, message = FALSE, error=FALSE,fig.height=10}
for.tip.mean.up.am <- merge(for.tip.mean.up.am, clusters.nodiff, 
                            by.x = "LAD19NM", by.y = "LAD", all.x = T)
for.tip.mean.up.am <- merge(for.tip.mean.up.am, for.tip.mean.up.pm, 
                            by.x = "LAD19NM", all.x = T)

for.tip.mean.up.am %>% 
  select(-cluster.down, -cluster.tests, -rank) %>% 
  arrange(-mean.up.am) %>% 
  relocate(cluster.up, .after=mean.up.pm) %>% 
  kable(digits = 0,
        format.args = list(big.mark = ","),
        col.names = c("LAD",
                      "Mean AM upload speed",
                      "Mean PM upload speed",
                      "Cluster")) %>% 
  kable_styling("striped", full_width = F) %>% 
  scroll_box() #width = "500px", height = "400px"
```

### Cluster description

- Cluster 1: Medium sized mostly rural cluster; slow mean upload speeds; high workday temporal variation.

- Cluster 2: Small rural cluster; highest mean upload speeds; slowdown more in evening.

- Cluster 3: Medium sized mixed cluster; fast mean upload speeds; medium temporal variation.

- Cluster 4: Rural reference LAD; medium mean upload speeds; high temporal variation.

- Cluster 5: Small suburban cluster; fast mean upload speeds; slowdown more in evening.

- Cluster 6: Large mixed cluster; slowest aggregate mean upload speeds; low workday temporal variation.

- Cluster 7: Small suburban cluster; fast mean upload speeds; medium temporal variation.

- Cluster 8: Small mainly surburban cluster; slow mean upload speeds; high workday temporal variation.

- Cluster 9: Large, more urban cluster; fast mean upload speeds; lowest workday temporal variation.