### 2021-05-11 paper/v2_taylor_francis_for_proof/tables/table_1.tex
typos and .tex files for tables

\begin{table}[!htbp] \centering 
  \caption{Upload speed cluster characteristics\label{up.cluster.descr}} 
  \label{} 
\footnotesize 
\begin{tabular}{@{\extracolsep{0pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Cluster & N. of LADs & LAD population & mean speed & SD speed & mean AM speed & mean PM speed \\ 
\hline \\[-1.8ex] 
1 & 9 & 903200 & 9564 & 6314 & 8798 & 10457 \\ 
2 & 2 & 162000 & 12085 & 6537 & 11882 & 10866 \\ 
3 & 12 & 1785800 & 11047 & 6079 & 10029 & 11634 \\ 
4 & 1 & 91100 & 9689 & 6122 & 7816 & 9689 \\ 
5 & 3 & 280000 & 10802 & 6116 & 11010 & 10084 \\ 
6 & 229 & 40552800 & 8761 & 5847 & 8555 & 8955 \\ 
7 & 5 & 682500 & 10326 & 6102 & 10045 & 11149 \\ 
8 & 6 & 510000 & 9769 & 6352 & 8989 & 10836 \\ 
9 & 115 & 21467800 & 10328 & 5915 & 10283 & 10333 \\ 
\hline \\[-1.8ex] 
\multicolumn{7}{l}{Note: All speed measures are upload speeds} \\ 
\end{tabular} 
\end{table}