### 2021-09-22 src/old/online_appendix_old.Rmd
tidy up

---
title: "Online Appendix"
author: ""
output: html_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/v2_taylor_francis/")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(plotly)
library(lubridate)
library(gridExtra)
library(openair)
library(rgdal)
library(maps)
library(rprojroot)
library(kableExtra)
#library(cowplot)
library(ggthemes)

path <- find_rstudio_root_file()
```

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
# load the data created by Data_Spatial.Rmd

path.data <- paste(path, "/data/temp/TSbb19_20sp.csv", sep = "")

TSbb19_20sp <- read_csv(path.data)

#split by year before create stats by each hour of each day of the week
TS2019 <- selectByDate(TSbb19_20sp, year = 2019) #272248 obs
TS2020 <- selectByDate(TSbb19_20sp, year = 2020) #343268 obs

# save clusters file
path.clusters <- paste(path, "/data/temp/clusters_nodiff_9.csv", sep = "")
clusters.nodiff <- read_csv(path.clusters)
```
## Local Authority Districts

The below map illustrates the different clusters of LAD based on the upload 
experienced internet speeds. Hovering over the LAD polygons, the map also presents
the mean AM upload internet speed and the relevant ranking in the country for each 
LAD.

**Is there a chance to briefly describe the clusters with 1-2 words?**

- Cluster 1:

- Cluster 2: 
...

<br>
```{r, echo=FALSE, warning=FALSE, message = FALSE, error=FALSE,fig.height=10}

# LAD fully generalised
la <- readOGR("https://opendata.arcgis.com/datasets/3a4fa2ce68f642e399b4de07643eeed3_0.geojson",
              verbose = FALSE)
# saved locally in /data/raw/Local_Authority_Districts_(December_2019)_Boundaries_UK_BUC.geojson

# spatial transformations
la <- spTransform(la, CRS("+init=epsg:4326"))
#la@data$LAD19NM <- as.character(la@data$LAD19NM)
la@data$lad19nm <- as.character(la@data$lad19nm)

# upload metrics
for.tip.mean.up.am <- TS2020 %>% # raw data, individual tests
  filter(hour(date)==9 | hour(date)==10) %>% 
  group_by(LAD19NM) %>%
  summarise(mean.up.am = mean(speedup)) %>% 
  mutate(rank = dense_rank(desc(mean.up.am)))

# not used
# for.tip.mean.up.pm <- TS2020 %>% # raw data, individual tests
#   filter(hour(date)==19 | hour(date)==20) %>% 
#   group_by(LAD19NM) %>%
#   summarise(mean.up.pm = mean(speedup)) 
# 
# for.tip.mean.up <- TS2020 %>% # raw data, individual tests
#   group_by(LAD19NM) %>%
#   summarise(mean.up = mean(speedup)) 
# 
# # download metrics
# for.tip.mean.down.am <- TS2020 %>% # raw data, individual tests
#   filter(hour(date)==9 | hour(date)==10) %>% 
#   group_by(LAD19NM) %>%
#   summarise(mean.down.am = mean(speeddown)) 
# 
# for.tip.mean.up.pm <- TS2020 %>% # raw data, individual tests
#   filter(hour(date)==19 | hour(date)==20) %>% 
#   group_by(LAD19NM) %>%
#   summarise(mean.down.pm = mean(speeddown)) 
# 
# for.tip.mean.up <- TS2020 %>% # raw data, individual tests
#   group_by(LAD19NM) %>%
#   summarise(mean.down = mean(speeddown))

# fortify
la.f <- broom::tidy(la, region = "lad19nm")
la.f <- merge(la.f, clusters.nodiff, by.x = "id", by.y = "LAD")

# create the text to tooltip
la.f <- merge(la.f, for.tip.mean.up.am, by.x = "id", by.y = "LAD19NM", all.x=T)
#sapply(la.f, function(x) sum(is.na(x)))
#No am data points for Isles of Scilly
la.f$tip <- paste0(la.f$id, "<br>",
                   "Upload am: ", round(la.f$mean.up.am), " Kbps", "<br>",
                   "Rank: ", la.f$rank)
la.f <- la.f[order(la.f$order),] # if i don't order merge.nuts.f loses order and the map has gaps

# using world.cities from `maps`
cities <- world.cities %>% filter(country.etc=="UK")

p <- ggplot(la.f, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = as.factor(cluster.up), 
                   text = tip)) +# ,text =paste("test1", long, "test2", lat)
  #labs(title = "Upload clusters") +
  scale_fill_brewer(palette="Set1", 
                    direction = -1) +
  #geom_text_repel( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat, label=name), size=4) +
  geom_point( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat), color="black", size=0.5) +
  geom_text( data=cities %>% arrange(pop) %>% tail(10), aes(x=long, y=lat, label=name), size=3) +

  guides(fill=guide_legend(title="Clusters")) +
  theme_map()
#   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
# panel.grid.minor = element_blank())

p %>%
  ggplotly(tooltip = "text") %>%
  style(hoverlabel = list(bgcolor = "white"), hoveron = "fill")
```

```{r, echo=FALSE, warning=FALSE, message = FALSE, error=FALSE,fig.height=10}
for.tip.mean.up.am %>% 
    arrange(rank) %>% 
  kable(digits = 0,
        format.args = list(big.mark = ",")) %>% 
  kable_styling("striped", full_width = F) %>% 
  scroll_box(width = "500px", height = "400px")

```